### 2021-04-06 Drafts and notes/Draft_01.tex

\documentclass[11pt,reqno]{amsart}
%============%============%============%============%

%============%============%============%============%
%\setlength{\columnseprule}{0.4pt}
%\setlength{\topmargin}{0cm}
%\setlength{\oddsidemargin}{.25cm}
%\setlength{\evensidemargin}{.25cm}
%\setlength{\textheight}{22.5cm}
%\setlength{\textwidth}{15.5cm}
\renewcommand{\baselinestretch}{1.05}
%============%============%============%============%
\usepackage[toc,page]{appendix}
%============%============%============%============%
%\usepackage{romannum}
\usepackage{xcolor}
\usepackage{placeins}
\usepackage{amsfonts,amsmath,amsthm}
\usepackage{amssymb,epsfig}
\usepackage{enumerate} 
\usepackage[notcite,notref]{showkeys}
\usepackage{fullpage}
%============%============%============%============%
\usepackage[utf8]{inputenc}
%\usepackage{mathpazo}

%\usepackage[libertine,cmintegrals,cmbrac,vvarbb]{newtxmath}
\usepackage{eucal}
\usepackage[euler-digits]{eulervm}
\usepackage[unicode=true]{hyperref}
\hypersetup{colorlinks = true}
%\hypersetup{hidelinks=true}
\hypersetup{
     colorlinks,
     linkcolor={black!10!red},
     linkbordercolor = {black!100!red},
%    <your other options...>,
     citecolor={blue}
}

%graphic
%\usepackage[text={425pt,650pt},centering]{geometry}

\usepackage{pdfsync}

\usepackage{geometry}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm,headheight=3.5cm}

\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{tikz}
\usepackage{caption}
\usepackage{color} %color
\definecolor{vert}{rgb}{0,0.6,0}

\usepackage{comment}
\numberwithin{figure}{section}
%\pagestyle{plain}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{ass}{Assumption}
\renewcommand{\theass}{}
\newtheorem{defn}{Definition}
\newtheorem{quest}{Question}
\newtheorem{com}{Comment}
\newtheorem{ex}{Example}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{remark}
\newtheorem{rem}{\bf{Remark}}
\numberwithin{equation}{section}

%\renewcommand{\thefootnote}{\fnsymbol{footnote}}

%Characters -- Shortcuts
\newcommand{\E}{\mathbb{E}}
\newcommand{\M}{\mathbb{M}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\bP}{\mathbb{P}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\bS}{\mathbb{S}}
\newcommand{\T}{\mathbb{T}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\bfS}{\mathbf{S}}
\newcommand{\cA}{\mathcal{A}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cL}{\mathcal{L}}
\newcommand{\cM}{\mathcal{M}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\cS}{\mathcal{S}}
\newcommand{\cT}{\mathcal{T}}
\newcommand{\cE}{\mathcal{E}}
\newcommand{\I}{\mathrm{I}}

%Functional spaces
\newcommand{\AC}{{\rm AC\,}}
\newcommand{\ACl}{{\rm AC}_{{\rm loc}}}
\newcommand{\BUC}{{\rm BUC\,}}
\newcommand{\USC}{{\rm USC\,}}
\newcommand{\LSC}{{\rm LSC\,}}
\newcommand{\Li}{L^{\infty}}
\newcommand{\Lip}{{\rm Lip\,}}
\newcommand{\W}{W^{1,\infty}}
\newcommand{\Wx}{W_x^{1,\infty}}

%Domains
\newcommand{\bO}{\partial\Omega}
\newcommand{\cO}{\overline\Omega}
\newcommand{\Q}{\mathbb{T}^{n}\times(0,\infty)}
\newcommand{\iQ}{\mathbb{T}^{n}\times\{0\}}
\newcommand{\cQ}{\mathbb{T}^{n}\times[0,\infty)}

%Greek alphabets -- Shortcuts
\newcommand{\al}{\alpha}
\newcommand{\gam}{\gamma}
\newcommand{\del}{\delta}
\newcommand{\ep}{\varepsilon}
\newcommand{\kap}{\kappa}
\newcommand{\lam}{ }
\newcommand{\sig}{\sigma}
\newcommand{\om}{\omega}
\newcommand{\Del}{\Delta}
\newcommand{\Gam}{\Gamma}
\newcommand{\Lam}{ }
\newcommand{\Om}{\Omega}
\newcommand{\Sig}{\Sigma}

%Overlines, Underlines -- Shortcuts
\newcommand{\ol}{\overline}
\newcommand{\ul}{\underline}
\newcommand{\pl}{\partial}
\newcommand{\supp}{{\rm supp}\,}
\newcommand{\inter}{{\rm int}\,}
\newcommand{\loc}{{\rm loc}\,}
\newcommand{\co}{{\rm co}\,}
\newcommand{\diam}{{\rm diam}\,}
\newcommand{\diag}{{\rm diag}\,}
\newcommand{\dist}{{\rm dist}\,}
\newcommand{\Div}{{\rm div}\,}
\newcommand{\sgn}{{\rm sgn}\,}
\newcommand{\tr}{{\rm tr}\,}
\newcommand{\Per}{{\rm Per}\,}

\newcommand{\rmC}{\mathrm{C}}
\newcommand{\rup}{\rightharpoonup}

\renewcommand{\subjclassname}{%
\textup{2010} Mathematics Subject Classification}

%Hyperlink in PDF file
%\usepackage[dvipdfm,
%  colorlinks=false,
%  bookmarks=true,
%  bookmarksnumbered=false,
%  bookmarkstype=toc]{hyperref}
%\makeatletter
%\def\@pdfm@dest#1{%
%  \Hy@SaveLastskip
%  \@pdfm@mark{dest (#1) [@thispage /\@pdfview\space @xpos @ypos null]}%
%  \Hy@RestoreLastskip
%}

%BibLatex
%\usepackage[
%backend=biber,
%style=alphabetic,
%sorting=ynt
%]{biblatex}
%\addbibresource{rate.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{import}
\usepackage{xifthen}
\usepackage{pdfpages}
\usepackage{transparent}
\newcommand{\incfig}[1]{%
    \def\svgwidth{\columnwidth}
    \import{./figs/}{#1.pdf_tex}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title[Rate of convergence]
{\textsc{Second-order state-constraint Hamilton-Jacobi equations}}
\thanks{The authors are supported in part by NSF grant DMS-1664424.}
\begin{abstract}
We investigate qualitatively the convergence of large, or state-constraint solution to nonlinear elliptic equation as the viscosity vanishes.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Yuxi Han}
\address[Y. Han]
{
Department of Mathematics, 
University of Wisconsin Madison, 480 Lincoln  Drive, Madison, WI 53706, USA}
\email{yuxi.han@wisc.edu}
\author{Son N. T. Tu}
\address[S. N.T. Tu]
{
Department of Mathematics, 
University of Wisconsin Madison, 480 Lincoln  Drive, Madison, WI 53706, USA}
\email{thaison@math.wisc.edu}
\date{\today}
\keywords{first-order Hamilton--Jacobi equations; state-constraint problems; optimal control theory; rate of convergence; viscosity solutions.}
\subjclass[2010]{
35B40, %Asymptotic behavior of solutions, 
35D40, %Viscosity solutions
49J20, %Optimal control problems involving partial differential equations
49L25, %Viscosity solutions
70H20 %Hamilton-Jacobi equations
}
\maketitle
\setcounter{tocdepth}{1}
\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
\subsection{Motivation} Let $\Omega$ be an open, bounded and connected with $\mathrm{C}^2$ boundary domain of $\mathbb{R}^n$. %Let us consider the following Hamiltonian $ H(x,\varrho) = |\varrho|^p-f(x)$ for $(x,p)\in \overline{\Omega}\times \mathbb{R}^n$,
Let $f\in \mathrm{C}(\overline{\Omega})\cap W^{1,\infty}(\Omega)$. Let $u^\varepsilon\in \mathrm{C}^2(\Omega)$ (see \cite{Lasry1989}) be the solution to 
\begin{equation}\label{eq:PDEeps}
    \begin{cases}
      u^\varepsilon(x) + |Du^\varepsilon(x)|^p - f(x) - \varepsilon \Delta u^\varepsilon(x) = 0 \qquad
    \text{in}\;\Omega, \vspace{0cm}\\
    \displaystyle  \lim_{\mathrm{dist}(x,\partial \Omega)\to 0} u^\varepsilon(x) = +\infty.
    \end{cases} \tag{PDE$_\varepsilon$}
\end{equation}
When $1<p\leq 2$, equation \eqref{eq:PDEeps} describes the value function associated with a minimization of a stochastic optimal control problem with state-constraint. We are interested in studying the asymptotic behavior of $\{u^\varepsilon\}_{\varepsilon>0}$ as $\varepsilon\rightarrow 0^+$. Heuristically, the state-constraint second-order problem converges to the state-constraint first-order, which is associated with deterministic optimal control, which is described in the framework of viscosity solution as follows:
\begin{equation}\label{eq:PDE0}
    \begin{cases}
       u(x) + |Du(x)|^p - f(x) \leq 0\;\qquad\text{in}\;\Omega,\\
       u(x) + |Du(x)|^p - f(x) \geq 0\;\qquad\text{on}\;\overline{\Omega}.
    \end{cases} \tag{PDE$_0$}
\end{equation}
Equation \eqref{eq:PDE0} admits a unique viscosity solution in the space $\rmC(\overline{\Omega})$, which is also the maximal viscosity subsolution among all viscosity subsolution $v\in\rmC(\overline{\Omega})$. The problem is interesting since in the limit we no longer have blowing up behavior. In this paper, we are interested in the rate of convergence of $u^\varepsilon \to u$ as $\varepsilon\to 0$. A boundary layer is expected to describe the behavior of this convergence near the boundary.
\subsection{Assumptions} We will always assume $\Omega$ is an open, bounded and connected with $\mathrm{C}^2$ boundary in $\mathbb{R}^n$ and $H:\mathbb{R}^n\times \mathbb{R}^n$ is a continuous Hamiltonian. 
\begin{itemize}
    \item[$\mathrm{(A1)}$] $H(x,\varrho) = |\varrho|^p - f(x)$ where $1<p\leq 2$ and $f\in \mathrm{C}(\overline{\Omega})\cap W^{1,\infty}(\Omega)$.
\end{itemize}
    The following assumptions on $\Omega$ will also be used.
\begin{itemize}    
    \item[(A2)] There exists a universal pair of positive numbers $(r,h)$ and $\eta\in \mathrm{BUC}(\overline{\Omega};\mathbb{R}^n)$ such that $B(x+t\eta(x), rt)\subset\Omega$ for all $t\in (0,h]$.
    \item[(A3)] $\Omega$ is a bounded and star-shaped (with respect to the origin) open subset of $\mathbb{R}^n$ such that
    \begin{equation*}
        \mathrm{dist}(x,\overline{\Omega}) \geq \kappa r \qquad\text{for all}\; x\in (1+r) \partial\Omega, \;\text{for all}\;r>0,
    \end{equation*}
for some $\kappa > 0$.
\end{itemize}
Assumption $\mathrm{(A1)}$ follows \cite{Lasry1989}, where the specific structure of the Hamiltonian enables more explicit estimates on the solution of \eqref{eq:PDEeps}. Assumptions $\mathrm{(A2)}$ or $\mathrm{(A3)}$ guarantees a comparison principle holds for the state-constraint problem \eqref{eq:PDE0}. We note that $\mathrm{(A1)}$ was first studied by M. Soner in \cite{Soner1986} (and $\mathrm{(A1)}$ holds for any smooth bounded domain), while $\mathrm{(A2)}$ was posed and studied by I. Capuzzo-Dolcetta and P. L. Lions in \cite{Capuzzo-Dolcetta1990}. 
%We list the assumptions on a general Hamiltonian as follow.
%\begin{itemize}
%    \item[$\mathrm{(H1)}$] There exists $C_1 > 0$ such that $H(x,p) \geq -C_1$ for all $(x,p)\in \overline{\Omega}\times\mathbb{R}^n$.
%    \item[$\mathrm{(H2)}$] There exists $C_2>0$ such that $|H(x,0)|\,\leq \,C_2$ for all $(x,p)\in \overline{\Omega}\times \mathbb{R}^n$.
%    \item[$\mathrm{(H3)}$] For each $R>0$ there exists a modulus $\omega_{R}[0,\infty)\to [0,\infty)$ such that $\omega_R(0^+) = 0$ and 
%    \begin{equation*}
%        \begin{cases}
%        |H(x,p) - H(y,p)| \leq \omega_R(|x-y|),\\
%        |H(x,p) - H(x,q)| \leq \omega_R(|p-q|),
%        \end{cases} \qquad\text{for all}\;x,y\in \overline{\Omega}, p,q \in \mathbb{R}^n\;\text{with}\;|p|,|q|\leq R.
%    \end{equation*}
%    \item[$\mathrm{(H4)}$] $H(x,p)\rightarrow \infty$ as $|p|\to \infty$ uniformly in $x\in \overline{\Omega}$.
%\end{itemize}

\section{Preliminaries}\label{sec:prelim} 
\subsection{Setting and simplifications} Let $\Omega$ be an open, bounded and connected subset of $\mathbb{R}^n$ with boundary $\partial\Omega$ is of class $C^2$. For small $\delta>0$, we denote $\Omega_\delta = \{x\in \Omega: \mathrm{dist}(x,\Omega) > \delta\}$ and $\Omega^\delta = \{x\in \mathbb{R}^n: \mathrm{dist}(x,\overline{\Omega}) < \delta\}$. 
\begin{figure}[ht]
    \centering
    %\incfig{Domains}
    \def\svgwidth{0.47\columnwidth}
    \import{./figs/}{Domains.pdf_tex}
    \caption{The domain $\Omega$ variations $\Omega_\delta, \Omega^\delta$.}
    \label{fig:Domains}
\end{figure}

\begin{defn} Let us define
\begin{equation}\label{def:delta_0}
    \delta_{0,\Omega} = \sup \big\{ \delta > 0: x\mapsto\mathrm{dist}(x,\partial\Omega)\;\text{is}\;\rmC^2\;\text{in}\;\Omega\backslash\overline{\Omega}_{\delta} \big\}.
\end{equation}
The distance function is of class $\mathrm{C}^2$ in the region where $0<\mathrm{dist}(x,\partial\Omega) < \delta_0$. We will simply write $\delta_0$ if the underlying domain is understood.
\end{defn}

We refer the readers to \cite{gilbarg_elliptic_2001} for the regularity of the distance function defined in the above definition. We then extend $\mathrm{dist}(x,\partial\Omega)$ to a function $d(x)\in \mathrm{C}^2(\mathbb{R}^n)$ such that 
\begin{equation}\label{e:distance_def}
    \begin{cases}
    d(x)\geq 0\;\text{for}\;x\in\Omega\;\text{with}\;d(x) = +\mathrm{dist}(x,\partial\Omega)\;\text{for}\;x\in \Omega\backslash \Omega_{\delta_0},\\
    d(x)\leq 0\;\text{for}\;x\notin \Omega\;\text{with}\;d(x) = -\mathrm{dist}(x,\partial\Omega)\;\text{for}\;x\in \Omega^{\delta_0}\backslash \Omega.
    \end{cases}
\end{equation}
We recall that $|D d(x)| = 1$ in the viscosity sense (and thus in classical sense) in $\Omega^{\delta_0}\backslash \Omega_{\delta_0}$. Let 
\begin{equation*}
   K_0:= \max_{x\in \overline{\Omega}}|d(x)|, \qquad K_1 := \max_{x\in \overline{\Omega}} |\Delta d(x)| \qquad\text{and}\qquad K_2 := \max_{x\in \overline{\Omega}} |D d(x)|.
\end{equation*}
\noindent We denote by $\mathcal{L}^\varepsilon:\rmC^2(\Omega)\to \rmC^2(\Omega)$ the elliptic operator
\begin{equation*}
    \mathcal{L}^\varepsilon[u](x) :=   u(x) + |Du(x)| - f(x) - \varepsilon \Delta u(x), \qquad x\in \Omega.
\end{equation*}

\subsection{Local gradient estimate} 
We recall the local gradient bound following \cite[Appendix]{Lasry1989}. The technique being used is the classical Bernstein's method. Let $\Omega$ be an open, bounded subset of $\R^n$, $\varepsilon \in (0,1)$. We consider $p>1$ in this section.
%We consider the Hamiltonian $H(x,\rho) = |\rho|^p - f(x)$ where $1<p < \infty$.

\begin{thm}\label{thm:grad_1} Let $f\in \rmC(\overline{\Omega})\cap W^{1,\infty}(\Omega)$ and $u\in \mathrm{C}^2(\Omega)$ be a solution to $\mathcal{L}^\varepsilon[u] = 0$ in $\Omega$. Assume that $  u(x)-f(x)\geq -L$ in $\Omega$, then for $\delta>0$ there exists $C_\delta = C(L,p,\delta, \Vert D f\Vert_{L^\infty(\Omega)})$ such that 
\begin{equation*}
    \sup_{x\in \overline{\Omega}_\delta} \Big(|u(x)|+|Du(x)|\Big) \leq C_\delta.
\end{equation*}
\end{thm}
\noindent We provide a proof for this Theorem in Appendix for the reader's convenience.

\subsection{Well-posedness of large solution for subquaratic case} In this section, with the specific form of the Hamiltonian $H(x,\varrho) = |\varrho|^p - f(x)$ where $1<p\leq 2$ and $f\in \mathrm{C}(\overline{\Omega})\cap W^{1,\infty}(\Omega)$, we show the existence and uniqueness of solutions to \eqref{eq:PDEeps}. 
We note that the assumption of $f$ can be relaxed to $f\in L^\infty(\Omega)$, but for the clarity of the proof we will assume $f\in \mathrm{C}(\overline{\Omega})\cap W^{1,\infty}(\Omega)$.

\begin{thm}\label{thm:wellposed1<p<2} If $H(x,\varrho) = |\varrho|^p - f(x)$ where $1<p< 2$ and $f\in \mathrm{C}(\overline{\Omega})\cap W^{1,\infty}(\Omega)$ then there exists a unique solution $u^\varepsilon\in \mathrm{C}^2(\Omega)$ of \eqref{eq:PDEeps} such that
\begin{equation*}
    %\frac{C_\varepsilon-\eta}{d(x)^\alpha}-\frac{M_\eta}{ } \leq u(x)\leq \frac{C_\varepsilon+\eta}{d(x)^\alpha}+\frac{M_\eta}{ } 
    \lim_{d(x)\to 0} u^\varepsilon(x) \,d(x)^\alpha = C_\alpha \varepsilon^{\alpha+1}
\end{equation*}
where
\begin{equation*}
    \displaystyle\alpha = \frac{2-p}{p-1} \qquad\text{and}\qquad C_\alpha = \left(\frac{1}{\alpha}(\alpha+1)^{\alpha+1}\right).
\end{equation*}
\end{thm}
\noindent This is Theorem I.1 in \cite{Lasry1989}. We provide a proof in Appendix with the explicit constant $C_\alpha\varepsilon^{\alpha+1}$ in terms of $\varepsilon$ for later use, it is useful to note that $\alpha+1 = \frac{1}{p-1}$. We also state the following useful results.

\subsection{Convergence result} We show qualitatively the convergence of \eqref{eq:PDEeps} to \eqref{eq:PDE0}. We first state the following Lemma (\cite{Capuzzo-Dolcetta1990}), which characterizes the state-constraint of the first-order equation
\begin{equation}\label{S_0}
\begin{cases}
       u(x) + |Du(x)| - f(x) = 0\;\qquad\text{in}\;\Omega,\\
       u(x) + |Du(x)| - f(x) \geq 0\;\qquad\text{on}\;\partial\Omega. 
\end{cases}\tag{$S_0$}   
\end{equation}

\begin{lem}\label{lem:max} Let $u\in \rmC(\overline{\Omega})$ be a viscosity subsolution of \eqref{S_0} such that, for all viscosity subsolution $v\in \rmC(\overline{\Omega})$ of \eqref{S_0} one has $v\leq u$ in $\overline{\Omega}$, then $u$ is a viscosity supersolution of \eqref{S_0} on $\overline{\Omega}$.
\end{lem}
\noindent We give a proof of Lemma \ref{lem:max} in Appendix for the sake of completeness.

\begin{lem}\label{lem:lower-bound} Let $u^\varepsilon\in \mathrm{C}^2(\Omega)$ be a solution to \eqref{eq:PDEeps}. We have $\{  u^\varepsilon\}_{\varepsilon>0}$ is uniformly bounded from below by a constant independent of $\varepsilon$. Precisely, we have $  u^\varepsilon \geq \min_\Omega f$ and $  u\geq \min_\Omega f$.
%\begin{equation}\label{e:lower_bound_u_eps}
%    \inf_{x\in \Omega}   u^\varepsilon(x) \geq  \inf_{\Omega} f
%\end{equation}
%and also
%\begin{equation}\label{e:lower_bound_u}
%    \inf_{x\in \Omega}   u(x) \geq  \inf_{\Omega} f
%\end{equation}
\end{lem}
\begin{proof} For $m\in \mathbb{N}$, let $u^{\varepsilon,m}\in \mathrm{C}^2(\Omega)\cap \rmC(\overline{\Omega})$ solves the Drichlet problem
\begin{equation}\label{e:uepsm}
    \begin{cases}
      u(x) + |Du(x)|^p - f(x) - \varepsilon \Delta u(x) = 0 &\qquad
    \text{in}\;\Omega, \vspace{0cm}\\
    \;\qquad\quad\qquad\qquad\qquad\qquad u(x) = m &\qquad
    \text{on}\;\partial\Omega.
    \end{cases} \tag{PDE$_{\varepsilon,m}$}
\end{equation}
We have $u^{\varepsilon}_m(x) \to u^\varepsilon(x)$ as $m\to \infty$ in $\Omega$. Let $\varphi(x) \equiv  ^{-1} \inf_{\Omega} f$ for $x\in \overline{\Omega}$ then $\varphi(x)$ is a classical subsolution of \eqref{e:uepsm} in $\Omega$ with
\begin{equation*}
    \varphi(x) =   \inf_\Omega f \leq m = u^\varepsilon_m(x) \qquad\text{for}\;x\in \partial\Omega.
\end{equation*}
By comparison principle for the uniformly elliptic equation \eqref{e:uepsm} we have
\begin{equation*}
     ^{-1}\inf_\Omega f \leq u^{\varepsilon}_m(x) \qquad\text{for all}\;x\in \Omega.
\end{equation*}
As $m\to \infty$ we obtain $  u^\varepsilon \geq \min_\Omega f$. The inequality $  u\geq \min_{\Omega}f$ follows from the comparison principle of \eqref{eq:PDE0} between supersolution $u$ on $\overline{\Omega}$ and subsolution $\varphi(x)$ in $\Omega$.
\end{proof}

We present a simple proof of the convergence $u^\varepsilon \to u$ using Lemma \ref{lem:max} for the reader's convenience, see also \cite[Theorem VII.3]{Capuzzo-Dolcetta1990}.
\begin{thm}[Vanishing viscosity]\label{thm:qual} Assume $\mathrm{(A1)}$. Let $u^\varepsilon$ be the solution to \eqref{eq:PDEeps}, then there exists $u \in \mathrm{C}(\overline{\Omega})$ such that $u^\varepsilon \rightarrow u$ locally uniformly in $\Omega$ as $\varepsilon\rightarrow 0$, $u$ solves \eqref{eq:PDE0} in $\Omega$.
\end{thm}

\begin{proof}[Proof of Theorem \ref{thm:qual}] By a priori estimate (Theorem \ref{thm:grad_1}) we have
\begin{equation}\label{e:priorie_eps}
    |u^\varepsilon(x)| + |Du^\varepsilon(x)| \leq C_\delta \qquad\text{for}\;x\in \overline{\Omega}_\delta
\end{equation}
we deduce from the Arzel\'a--Ascoli theorem that there exists a subsequence $\varepsilon_j\to 0$ and a function $u\in \rmC(\Omega)$ such that $u^{\varepsilon_j}\to u$ locally uniformly in $\Omega$. By the stability of viscosity solution we easily deduce that 
\begin{equation}\label{eq:u0int}
     u(x) + |Du(x)| - f(x) = 0 \qquad\text{in}\;\Omega.
\end{equation}
From Lemma \ref{lem:lower-bound} we have $  u^\varepsilon(x)\geq \min_{\Omega} f$ and  $  u(x)\geq \min_{\Omega} f$ for all $x\in \Omega$, together with \eqref{eq:u0int} we obtain $|\xi|\leq \max_\Omega f - \min_\Omega f$ for all $\xi\in D^+u(x)$ and $x\in \Omega$. In other words there is a constant $C_0$ such that
\begin{equation*}
    |u(x) - u(y)| \leq C_0|x-y| \qquad\text{for all}\;x,y\in \Omega.
\end{equation*}
Thus we can extend $u$ uniquely to $u\in \rmC(\overline{\Omega})$. We will use Lemma \ref{lem:max} to show that $u$ is a supersolution of \eqref{S_0} on $\overline{\Omega}$. It suffices to show that $u\geq w$ on $\overline{\Omega}$ where $w\in \rmC(\overline{\Omega})$ is the unique solution to \eqref{eq:PDE0}.

For $\delta>0$, let $\Omega_\delta = \{x\in \Omega: \mathrm{dist}(x,\partial \Omega) > \delta\}$ and $u_\delta\in\rmC(\overline{\Omega}_\delta)$ be the state-constraint viscosity solution to the problem $  u(x) + |Du(x)| - f(x) = 0$ in $\Omega_\delta$, i.e., $u_\delta$ is the unique viscosity solution to
\begin{equation}\label{e:v_v}
    \begin{cases}
      u_\delta(x) + |Du_\delta(x)|-f(x) \leq 0 &\qquad\text{in}\;\Omega_\delta,\\
      u_\delta(x) + |Du_\delta(x)| - f(x) \geq 0 &\qquad\text{on}\;\overline{\Omega}_\delta.
    \end{cases}
\end{equation}
As $u_\delta\rightarrow w$ locally uniformly as $\delta\rightarrow 0^+$ (see \cite{kim_state-constraint_2020}) and $w$ is bounded, therefore $\{u_\delta\}_{\delta>0}$ is uniformly bounded. Let $v^\varepsilon_\delta\in \rmC^2(\Omega_\delta)\cap \rmC(\overline{\Omega}_\delta)$ be the unique solution to the Dirichlet problem
\begin{equation}\label{eq:vv_eps}
\begin{cases}
      v_\delta^\varepsilon(x) + |Dv_\delta^\varepsilon(x)|^p - f(x) = \varepsilon \Delta v_\delta^\varepsilon(x) &\qquad\text{in}\;\Omega_\delta,\\
    \;\;\;\,\quad\qquad\qquad\qquad\qquad v_\delta^\varepsilon = v_\delta &\qquad \text{on}\;\partial\Omega_\delta.
\end{cases}
\end{equation}
It is well-known that $v^\varepsilon_\delta\to v_\delta$ uniformly on $\overline{\Omega}_\delta$ as $\varepsilon\to 0$.

For all $\delta$ small enough $v_\delta\leq u^\varepsilon$ on $\partial \Omega_\delta$, thus by maximum principle $v^\varepsilon_\delta \leq u^\varepsilon$ on $\overline{\Omega}_\delta$. Let $\varepsilon\to 0$ we have $v_\delta \leq u$ on $\overline{\Omega}_\delta$.
Let $\delta\rightarrow 0$ we obtain $w\leq u$ in $\Omega$, which implies $w\leq u$ on $\overline{\Omega}$ since both $w,u$ belong to $\rmC(\overline{\Omega})$.
\end{proof}

\section{Rate of convergence via the doubling variables method}
In this section, we focus on the rate of convergence for the case $f\in \mathrm{C}^1(\overline{\Omega})$. By adding a constant to the solution $u^\varepsilon$ of \eqref{eq:PDEeps}, without loss of generality we can assume the \emph{nonnegativity}
\begin{equation}\label{e:minf}
    \min_{\overline{\Omega}} f(x) = 0.
\end{equation}
As a consequence, $u^\varepsilon(x),u(x)\geq 0$ for $x\in \Omega$ by Lemma \ref{lem:lower-bound}.
\subsection{Lower bound}
We start with the lower bound of $u^\varepsilon - u$.
\begin{thm} Let $\Omega$ be an open, bounded and connected subset of $\R^n$ with $\mathrm{C}^2$ boundary. Assume \eqref{e:minf} and $H(x,\xi) = |\xi|^p - f(x)$ where $p\in\left(1,2\right]$. Let $u^\varepsilon$ be the unique solution to \eqref{eq:PDEeps} and $u$ be the unique solution to \eqref{eq:PDE0}, then there exists a constant $C$ independent of $\varepsilon$ such that
\begin{equation}\label{e:lower1}
    -C\sqrt{\varepsilon} \leq u^\varepsilon(x) - u(x) \qquad\text{for all}\;x\in \Omega.
\end{equation} 
\end{thm}
\begin{proof} The proof relies on a well-known rate of convergence for vanishing viscosity of viscous Hamilton--Jacobi equation with Dirichlet boundary condition (\textcolor{blue}{citation}). As $u\in \mathrm{C}(\overline{\Omega})$, let $g(x) = u(x)$ for $x\in \partial\Omega$ then $g\in \mathrm{C}(\partial\Omega)$. Let $v^\varepsilon\in \mathrm{C}^2(\Omega)\cap \mathrm{C}(\overline{\Omega})$ be the unique viscosity solution to
\begin{equation*}
    \begin{cases}
          v^\varepsilon(x) + |Dv^\varepsilon(x)|^p - f(x) - \varepsilon \Delta v^\varepsilon(x) = 0 &\qquad\text{in}\;\Omega,\\
        \qquad\qquad\qquad\qquad\qquad \qquad\quad v^\varepsilon(x) = g(x) &\qquad\text{on}\;\partial\Omega.
    \end{cases}
\end{equation*}
It is well-known that $v^\varepsilon \to u$ uniformly where $v$ is the viscosity solution to
\begin{equation*}
\begin{cases}
       v(x) + |Dv(x)| - f(x) = 0 &\qquad\text{in}\;\Omega,\\
     \qquad\qquad\qquad\quad\; v(x) = g(x)&\qquad\text{on}\;\partial\Omega. 
\end{cases}
\end{equation*}
By comparison principle, it is clear that $v = u$, the solution of \eqref{eq:PDE0}. Furthermore, there exists a positive constant $C$ such that 
\begin{equation}\label{e:cp1}
     |v^\varepsilon(x)  - u(x)| \leq C\sqrt{\varepsilon} \qquad\text{for}\;x\in \overline{\Omega}.
\end{equation}
By comparison principle for \eqref{eq:PDEeps} we have
\begin{equation}\label{e:cp2}
    v^\varepsilon(x)\leq u^\varepsilon(x) \qquad\text{for}\;x\in \Omega.
\end{equation}
From \eqref{e:cp1} and \eqref{e:cp2} we obtain the lower bound \eqref{e:lower1}.
\end{proof}

%We state some observations  regarding the property of the solution $u$ to \eqref{eq:PDE0}.

\begin{lem}\label{lem:f=0} If $f\geq 0$ in $\Omega$ then $u(x) = 0$ whenever $f(x) = 0$. In particular $f \equiv 0$ implies $u \equiv 0$.
\end{lem}
\begin{proof} We have $u\geq 0$ in $\overline{\Omega}$. Using the optimal control formula (\cite{Bardi1997,Hung2019}), we have
\begin{equation}\label{e:foru}
    u(x) = \inf \left\lbrace \int_0^\infty e^{-  s}\Big(c|\dot{\eta}(s)|^{q} +f(\eta(s))\Big)ds: \eta\in \mathrm{AC}([0,\infty);\overline{\Omega}), \eta(0) = x\right\rbrace
\end{equation}
where 
\begin{equation*}
    c = \left(pq^{\frac{1}{p-1}}\right)^{-1} \qquad\text{and}\qquad \frac{1}{p} + \frac{1}{q} = 1.
\end{equation*}
Assume that $f(x) = 0$, we can choose $\eta(s) = x$ for all $s\in [0,\infty)$ as an admissible path in \eqref{e:foru} to obtain that
\begin{equation*}
    u(x) \leq  \int_0^\infty e^{-  s}\Big(c|\dot{\eta}(s)|^{q} +f(\eta(s))\Big)ds = f(x) =  0.
\end{equation*}
\end{proof}
\noindent For simplicity, let us assume $  = 1$ from now on. We state the following lemma as a crucial estimate we will be using. It is a refined construction of the supersolution for \eqref{eq:PDEeps}.
\begin{lem}\label{lem:super_refined} There exist positive constants $\nu = \nu(\alpha,\delta_0)$ and $C_\nu$ such that if $0\leq \delta \leq \frac{1}{2}\delta_0$ then
\begin{equation*}
    w(x) = \frac{\nu C_\alpha \varepsilon^{\alpha+1}}{(d(x)-\delta)^\alpha} + \max f + C_\nu \varepsilon^{\alpha+2}, \qquad x\in \Omega_\delta 
\end{equation*}
is a supersolution of \eqref{eq:PDEeps} in $\Omega_\delta$. 
\end{lem}

\begin{proof}
We have $|\nabla d(x)| = 1$ for all $x\in \Omega_\delta\backslash \Omega_{\delta_0}$. Recall from Theorem \ref{thm:wellposed1<p<2} that $C_\alpha^p \alpha^p = C_\alpha \alpha (\alpha+1)$ and $p(\alpha+1) = \alpha+2$, we compute 
\begin{equation*}
    |\nabla w(x)|^p = \nu^p\frac{(C_\alpha\alpha)^p\varepsilon^{p(\alpha+1)}}{(d(x)-\delta)^{p(\alpha+1)}} = \nu^p\frac{C_\alpha \alpha(\alpha+1)\varepsilon^{\alpha+2}}{(d(x)-\delta)^{\alpha+2}}
\end{equation*}
and
\begin{equation*}
    \varepsilon\Delta w(x) = \nu\frac{C_\alpha\alpha(\alpha+1)\varepsilon^{\alpha+2}}{(d(x)-\delta)^{\alpha+2}} - \nu\frac{C_\alpha\alpha\varepsilon^{\alpha+2}\Delta d(x)}{(d(x)-\delta)^{\alpha+1}}.
\end{equation*}
We have
\begin{equation*}
    \begin{split}
        \mathcal{L}\left[w_\delta\right] \geq \frac{\nu C_\alpha\varepsilon^{\alpha+1}}{(d(x)-\delta)^\alpha} + \frac{C_\alpha \alpha (\alpha+1)\varepsilon^{\alpha+2}}{(d(x)-\delta)^{\alpha+2}}\left[\nu^p-\nu +\nu\frac{(d(x)-\delta)\Delta d(x)}{\alpha+1}\right].
    \end{split}
\end{equation*}
Recall that $K_2= \Vert \Delta d\Vert_{L^\infty}$, we observe that in the region where $\delta< d(x)\leq \delta_0$ then
\begin{equation*}
    \left|\frac{(d(x)-\delta)\Delta d(x)}{\alpha+1}\right| \leq \frac{\delta_0\Vert \Delta d\Vert_{L^\infty}}{\alpha+1} \leq \frac{K_2\delta_0}{\alpha+1}\leq K_2\delta_0.
\end{equation*}
Therefore
\begin{equation}\label{e:choose_nu}
    \begin{split}
        \nu^p-\nu +\nu\frac{(d(x)-\delta)\Delta d(x)}{(\alpha+1)} \geq \nu^p - \nu - \nu K_2\delta_0 = \nu\Big(\nu^{p-1} - (1+K_2\delta_0)\Big).
    \end{split}
\end{equation}
We will choose $\nu$ as follows. For $\gamma>1$ we have
\begin{equation*}
    \Big||x+y|^\gamma - |x|^\gamma\Big|\leq \gamma\Big(|x|+|y|\Big)^{\gamma-1}|y|
\end{equation*}
which implies that
\begin{equation*}
     0 \leq (1+K_2\delta_0)^{\alpha+1} - 1 \leq\underbrace{(\alpha+1)\left(1+K_2\delta_0\right)^\alpha K_2}_{C_2}\delta_0.
\end{equation*}
which implies that $(1+K_2\delta_0)^{\alpha+1} \leq 1 + C_2\delta_0$, thus
\begin{equation}\label{e:cru1}
    (1+K_2\delta_0) \leq (1+C_2\delta_0)^{\frac{1}{\alpha+1}} = (1+C_2\delta_0)^{p-1}
\end{equation}
since $\alpha+1 = \frac{1}{p-1}$. Choose $\nu = 1+C_2\delta_0$ in \eqref{e:choose_nu} we obtain $\mathcal{L}[w]\geq 0$ in $\{x\in \Omega_\delta: \delta <d(x)\leq \delta_0\}$. Recall that $K_0 = \Vert d\Vert_{L^\infty}$ and $K_1 = \Vert \nabla d\Vert_{L^\infty}$. In the region where $d(x)\geq \delta_0$ then $d(x)-\delta \geq \frac{1}{2}\delta_0$, we have 
\begin{align*}
    \mathcal{L}[w] &= \frac{\nu C_\alpha\alpha\varepsilon^{\alpha+1}}{(d(x)-\delta)^\alpha} + \max_{\Omega} f - f(x)\\
    &+  \nu^p\frac{C_\alpha\alpha(\alpha+1)\varepsilon^{\alpha+1}}{(d(x)-\delta)^{\alpha+2}}|\nabla d(x)|^p - \nu \frac{C_\alpha \alpha(\alpha+1)\varepsilon^{\alpha+2}}{(d(x)-\delta)^{\alpha+2}}|\nabla d(x)|^2 
    + \nu \frac{C_\alpha \alpha \varepsilon^{\alpha+2}\Delta d(x)}{(d(x)-\delta)^{\alpha+1}} + C_\nu \varepsilon^{\alpha+2}\\
    &\geq \frac{C_\alpha \alpha(\alpha+1)\varepsilon^{\alpha+2}}{(d(x)-\delta)^{\alpha+2}}\left(\nu^p|\nabla d(x)|^p - \nu |\nabla d(x)|^2 + \nu \frac{\Delta d(x)(d(x)-\delta)}{\alpha+1}\right) + C_\nu \varepsilon^{\alpha+2}\\
    &\geq (C_\nu - C_3)\varepsilon^{\alpha+2}
\end{align*}
where 
\begin{equation*}
    C_3 = \left(\frac{2}{\delta_0}\right)^{\alpha+2} C_\alpha\alpha(\alpha+1) \left(\nu^pK_1^p + \nu K_1^2 + \nu \frac{K_0K_2}{\alpha+1}\right).
\end{equation*}
Therefore we can choose $C_\nu = C_3$ to obtain $\mathcal{L}[w]\geq 0$ in this region $\{x\in \Omega_\delta:d(x)\geq \delta_0\}$.
\end{proof}

\subsection{Zero data}
\begin{thm}\label{thm:rate_doubling0} Let $\Omega$ be an open, bounded and connected subset of $\R^n$ with $\mathrm{C}^2$ boundary. Assume that $H(x,\xi) = |\xi|^p - f(x)$ where $p\in\left(1,2\right)$ and $f\equiv 0$ in $\Omega$. Let $u^\varepsilon$ be the unique solution to \eqref{eq:PDEeps} and $u \equiv 0 $ be the unique solution to \eqref{eq:PDE0}. Then there exists a constant $C$ independent of $\varepsilon$ such that
\begin{equation*}
 -C \varepsilon^{\frac{1}{2}}\leq  u^\varepsilon(x) - u(x) \leq C\varepsilon \quad  \text{for} \quad x\in \Omega_{\varepsilon}.
\end{equation*}
\end{thm}

\begin{proof} From Lemma \ref{lem:super_refined} and comparison principle, we have the estimate
\begin{equation*}
    u^\varepsilon(x)\leq \frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x)^\alpha} + C_\nu\varepsilon^{\alpha+2}, \qquad x\in \Omega.
\end{equation*}
Note that in this case $u^\varepsilon \to u \equiv 0$ as $\varepsilon\to 0$. We consider
\begin{equation*}
    \Phi(x,y) := u^\varepsilon(x) - u(y) - \frac{C_0|x-y|^2}{\sigma} - (1+\delta^\alpha)\frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x)^\alpha}, \qquad (x,y)\in \overline{\Omega}\times\overline{\Omega}.
\end{equation*}
It has a maximum at $(x_\sigma,y_\sigma)\in \Omega\times\overline{\Omega}$. From $\Phi(x_\sigma,y_\sigma)\geq \Phi(x_\sigma,x_\sigma)$ we have
\begin{equation*}
    \frac{C_0|x_\sigma-y_\sigma|^2}{\sigma}\leq u(x_\sigma) - u(y_\sigma)\leq C_0|x_\sigma - y_\sigma| \qquad\Longrightarrow\qquad |x_\sigma - y_\sigma|\leq \sigma.
\end{equation*}
From $\Phi(x_\sigma,y_\sigma)\geq \Phi(x_0,x_0)$ for a fixed point $x_0 \in \Omega$ that has the largest distance away from the boundary, we have
\begin{equation*}
    \delta^\alpha\frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x_{\sigma})^\alpha} \leq u^\varepsilon(x_{\sigma}) - \frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x_{\sigma})^\alpha} + C\varepsilon^{\alpha+1} \leq C_{\nu}\varepsilon^{\alpha+2}+ C\varepsilon^{\alpha+1}.
\end{equation*}
Thus
\begin{equation*}
    d(x_\sigma) \geq C\delta.
\end{equation*}
Now the subsolution test of $x\mapsto \Phi(x,y_\sigma)$ gives us
\begin{equation*}
\begin{split}
    u^\varepsilon(x) + &\left|\frac{2C_0(x_\sigma-y_\sigma)}{\sigma}-(1+\delta^\alpha)\frac{\nu C_\alpha\alpha\varepsilon^{\alpha+1}\nabla d(x_\sigma)}{d(x_\sigma)^{\alpha+1}}\right|^p \\
    &- \varepsilon \left(\frac{2nC_0}{\sigma} - (1+\delta^\alpha)\frac{\nu C_\alpha\alpha(\alpha + 1)\varepsilon^{\alpha+1}|\nabla d(x_\sigma)|^2}{d(x_\sigma)^{\alpha+2}} + (1+\delta^\alpha)\frac{\nu C_\alpha \alpha \varepsilon^{\alpha+1}\Delta d(x_\sigma)}{d(x_\sigma)^{\alpha+1}}\right)\leq 0.
\end{split}    
\end{equation*}
Similarly, the supersolution test of $y\mapsto \Phi(x_\sigma,y)$ gives us
\begin{equation*}
    u(y_\sigma) + \left|\frac{2C_0(x_\sigma-y_\sigma)}{\sigma}\right|^p \geq 0.
\end{equation*}
Therefore
\begin{equation}
\label{doubling}
\begin{split}
    u^\varepsilon(x_\sigma) - u(y_\sigma) \leq & p\left|2C_0+(1+\delta^\alpha)\nu C_\alpha \alpha\left( \frac{\varepsilon}{\delta}\right)^{\alpha+1}\right|^{p-1}(1+\delta^\alpha) \nu C_\alpha \alpha \left(\frac{\varepsilon}{\delta}\right)^{\alpha+1} \\
    &+2nC_0\frac{\varepsilon}{\sigma} + (1+\delta^\alpha)\nu C_\alpha \alpha(\alpha+1)\left(\frac{\varepsilon}{\delta}\right)^{\alpha+2} + (1+\delta^\alpha)\nu C_\alpha \alpha C \left(\frac{\varepsilon}{\delta}\right)^{\alpha+1}\varepsilon\\
    &\leq C\left(\frac{\varepsilon}{\sigma} + \left(\frac{\varepsilon}{\delta}\right)^{\alpha+1} + \left(\frac{\varepsilon}{\delta}\right)^{\alpha+2} \right).
\end{split}
\end{equation}
Max happens when we choose $\sigma = 1$, and hence
\begin{equation*}
    u^\varepsilon(x_\sigma) - u(y_\sigma) \leq C\left( \frac{\varepsilon}{1} + \left(\frac{\varepsilon}{\delta}\right)^{\alpha+1} + \left(\frac{\varepsilon}{\delta}\right)^{\alpha+2}\right).
\end{equation*}
Choosing $\delta = \varepsilon^\frac{\alpha}{\alpha+1}$ we obtain 
\begin{equation*}
    u^\varepsilon(x_\sigma) - u(y_\sigma) \leq C(\varepsilon + \varepsilon + \varepsilon ^p ) \leq C\varepsilon.
\end{equation*}
As a consequence, 
\begin{equation}
\label{zerodata}
   -C \varepsilon^{\frac{1}{2}}\leq  u^\varepsilon(x) - u(x) \leq C\varepsilon + \left(1+\varepsilon^{\frac{\alpha^2}{\alpha+1}}\right)\frac{\nu C_\alpha  \varepsilon^{\alpha+1}}{d(x)^\alpha} 
\end{equation}
for every $x\in \Omega$.
\end{proof}

\begin{rem} \textcolor{blue}{We believe that in case $f\equiv  0$, the lower bound in Theorem \ref{thm:rate_doubling0} can be made to $-C\varepsilon \leq u^\varepsilon - u$ instead of $-C\sqrt{\varepsilon}$.}
\end{rem}

\subsection{Compactly supported data}
\begin{thm}\label{thm:rate_doubling0} Let $\Omega$ be an open, bounded and connected subset of $\R^n$ with $\mathrm{C}^2$ boundary. Assume that $H(x,\xi) = |\xi|^p - f(x)$ where $p\in\left(1,2\right)$ and $f$ is compactly supported in $\Omega$. Let $u^\varepsilon$ be the unique solution to \eqref{eq:PDEeps} and $u$ be the unique solution to \eqref{eq:PDE0}. Then there exists a constant $C$ depends on the support of $f$ such that
\begin{equation*}
|u^\varepsilon(x) - u(x)| \leq C\sqrt{\varepsilon} \quad  \text{for} \quad x\in \Omega_{\varepsilon}.
\end{equation*}
Furthermore, in the region where $f$ vanishes, the rate of convergence can be improved to $\mathcal{O}(\varepsilon)$ in $\Omega_\varepsilon$.
\end{thm}

\begin{proof} Without loss of generality, we assume $f$ is compactly supported in $\Omega_{\kappa}$ where $0<\kappa < \frac{1}{2}\delta_0$. Let $g_\kappa = u^\varepsilon$ on $\partial\Omega_{\kappa}$, then the solution $u^\varepsilon$ of \eqref{eq:PDEeps} also solves
\begin{equation*}
    \left\{
  \begin{aligned}
  u^\varepsilon(x) + |Du^\varepsilon(x)|^p-\varepsilon \Delta u^\varepsilon(x) &=0 \;\qquad \text{in } U ,\\
  u^\varepsilon(x) &= +\infty \quad \text{on } \partial \Omega,\\
  u^\varepsilon(x) &= g_\kappa \;\quad \text{on } \partial \Omega_{\kappa},
    \end{aligned}
\right.
\end{equation*}
in the annulus $U_\kappa= \Omega \setminus \overline{\Omega}_{\kappa} = \{x\in \Omega: 0< d(x) < \kappa\}$. From Theorem \ref{thm:wellposed1<p<2}, let $\tilde{u}^\varepsilon\in \mathrm{C}^2(U_\kappa)$ be the solution to the following problem in the annulus $U_\kappa= \Omega \setminus \overline{\Omega}_{\kappa}$
\begin{equation*}
    \left\{
        \begin{aligned}
            \tilde{u}^\varepsilon(x) + |D\tilde{u}^\varepsilon(x)|^p-\varepsilon \Delta \tilde{u}^\varepsilon(x) &=0 \;\qquad \text{in } U_\kappa ,\\
            \tilde{u}^\varepsilon(x) &= +\infty \quad \text{on } \partial U = \partial \Omega\cup \partial \Omega_{\kappa}.
        \end{aligned}
    \right.
\end{equation*}
Here the boundary condition is understood as $\tilde{u}^\varepsilon(x)\to \infty$ as $d_\kappa(x)\to 0$ where $d_\kappa(\cdot)$ is the distance function from the boundary of the annulus $U_\kappa$, i.e.,
\begin{equation*}
    d_\kappa(x) = \min \big\lbrace \mathrm{dist}(x,\partial \Omega_\kappa),\mathrm{dist}(x)  \big\rbrace \leq d(x) \qquad\text{for}\;x\in U_\kappa.
\end{equation*}
Since $f = 0$ in $\overline{U}$, by Lemma \ref{lem:f=0} $u=0$ in $\overline{U}$, thus it is also the unique state-constraint solution to
\begin{equation*}
    \left\{
        \begin{aligned}
            u(x)+ |Du(x)|^p &=0 \quad \text{in } U_\kappa ,\\
            u(x)+ |Du(x)|^p &\geq 0 \quad \text{on } \partial U_\kappa = \partial \Omega \cup\partial \Omega_{\kappa}.
        \end{aligned}
    \right.
\end{equation*}
From \eqref{zerodata} and the fact that $u \equiv 0$ in $U$, there exists a constant $C$ (depends on $\kappa$) such that
\begin{equation} \label{annulus}
   \tilde{u}^\varepsilon(x) \leq C\varepsilon + \left(1+\varepsilon^{\frac{\alpha^2}{\alpha+1}}\right)\frac{\nu C_\alpha  \varepsilon^{\alpha+1}}{d_\kappa(x)^\alpha} \qquad\text{for}\;x\in U_\kappa.
\end{equation}
%Since $f \equiv 0 $ in $U$, $u (x) = 0$ in $U$, which gives us
%\begin{equation}\label{annulus}
%   -C \varepsilon^{\frac{1}{2}}\leq \tilde{u}^\varepsilon(x)  \leq C\varepsilon + \left(1+\varepsilon^{\frac{\alpha^2}{\alpha+1}}\right)\frac{\nu C_\alpha  \varepsilon^{\alpha+1}}{\tilde{d}(x)^\alpha}. 
%\end{equation}
By comparison principle in $U_\kappa$, $u^\varepsilon \geq 0$ and \eqref{annulus} we have
\begin{equation}\label{annulus2}
    0\leq u^\varepsilon(x) \leq \tilde{u}^\varepsilon(x) \leq C\varepsilon + \left(1+\varepsilon^{\frac{\alpha^2}{\alpha+1}}\right)\frac{\nu C_\alpha  \varepsilon^{\alpha+1}}{d_\kappa(x)^\alpha}  \qquad\text{for}\;x\in U_\kappa.
\end{equation}
In other words, in the region where $f = 0$ the have an upper bound for $u^\varepsilon - u$. We proceed with the doubling variable method. Let $\sigma,\delta\in (0,1)$, we consider the auxiliary functional 
\begin{equation*}
    \Phi(x,y)= u^\varepsilon(x) - u(y) -\frac{C_0|x-y|^2}{\sigma} - (1+\delta^\alpha) \frac{C_\alpha \varepsilon^{\alpha +1}}{d(x)^\alpha}, \qquad (x,y)\in \overline{\Omega}\times \overline{\Omega}.
\end{equation*}
Since $\displaystyle d(x)^\alpha u^\varepsilon(x) \to C_\alpha \varepsilon^{\alpha+1}$ as $d(x) \to 0$, we deduce that
\begin{equation*}
    \max_{(x,y) \in \overline{\Omega} \times \overline{\Omega}} \Phi(x,y) = \Phi(x_\sigma, y_\sigma) \qquad\text{for some}\;(x_\sigma,y_\sigma) \in \Omega \times \overline{\Omega}.
\end{equation*}
Similar to the zero data case, from $\Phi(x_\sigma, y_\sigma) \geq \Phi(x_\sigma, x_\sigma)$, we can deduce that
\begin{equation*}
    \left| x_\sigma - y_\sigma \right| \leq \sigma.  
\end{equation*}
If $ d(x_\sigma) \geq \frac{1}{2}\kappa$, then by a similar  computation of subsolution test as in \eqref{doubling} we get 
\begin{align*}
     u^\varepsilon(x_\sigma) - u(y_\sigma) \leq & p\left(2C_0+(1+\delta^\alpha)\nu C_\alpha \alpha\left( \frac{2\varepsilon}{\kappa}\right)^{\alpha+1}\right)^{p-1}(1+\delta^\alpha) \nu C_\alpha \alpha \left(\frac{2\varepsilon}{\kappa}\right)^{\alpha+1} + C\sigma\nonumber\\
    &+\frac{2nC_0\varepsilon}{\sigma} + (1+\delta^\alpha)\nu C_\alpha \alpha(\alpha+1)\left(\frac{2\varepsilon}{\kappa}\right)^{\alpha+2} + (1+\delta^\alpha)\nu C_\alpha \alpha C \left(\frac{2\varepsilon}{\kappa}\right)^{\alpha+1}\varepsilon\nonumber\\
    &\leq C\left(\sigma + \frac{\varepsilon}{\sigma} + \left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1} + \left(\frac{\varepsilon}{\kappa}\right)^{\alpha+2} \right).
\end{align*}
Max happens when $\sigma = \sqrt{\varepsilon}$, for example, which gives us  
\begin{equation}\label{e:final1}
    \Phi(x_\sigma,y_\sigma) \leq u^\varepsilon(x_\sigma) - u(y_\sigma) \leq C\sqrt{\varepsilon}.
\end{equation}
If $d(x_\sigma) < \frac{1}{2}\kappa$ then $x_\sigma \in U_\kappa$ and furthermore $\mathrm{dist}(x_\sigma,\partial \Omega_\kappa) > \frac{1}{2}\kappa$, therefore $d_\kappa(x_\sigma) = d(x_\sigma)$. Indeed, for any $y\in\partial\Omega$ and $z\in \partial\Omega_k$, we have
\begin{equation*}
    |x_\sigma - z| + |x_\sigma - y| \geq |y-z|.
\end{equation*}
Taking the infimum over all $y\in \partial\Omega$ we deduce that
\begin{equation*}
    |x_\sigma - z| + d(x_\sigma) \geq \inf_{y\in \partial \Omega}|y-z| = d(z) = \kappa
\end{equation*}
since $z\in \partial\Omega_k = \{x\Omega: d(x) = \kappa\}$. Thus we have $|x_\sigma - z| \geq \kappa - d(x_\sigma) > \frac{1}{2}\kappa$ for all $z\in \partial\Omega_k$. We have further that
\begin{equation*}
    \mathrm{dist}(x_\sigma,\partial\Omega_\kappa) = \kappa - d(x_\sigma).
\end{equation*}
Using \eqref{annulus2} and the fact that $u \geq 0$ we have
\begin{align}
    \Phi(x_\sigma, y_\sigma) &\leq u^\varepsilon (x_\sigma)-\left(1+\delta^\alpha\right)\frac{\nu C_\alpha  \varepsilon^{\alpha+1}}{d(x_\sigma)^\alpha}\nonumber\\
    &= u^\varepsilon (x_\sigma)-\left(1+\varepsilon^{\frac{\alpha^2}{\alpha+1}}\right)\frac{\nu C_\alpha  \varepsilon^{\alpha+1}}{d_\kappa(x_\sigma)^\alpha}
    \leq C \varepsilon\label{e:final2}
\end{align}
where we choose $\delta = \varepsilon^{\frac{\alpha}{\alpha+1}}$. Since $\Phi(x,x) \leq \Phi(x_\sigma,y_\sigma)$ for all $x\in \Omega$, we obtain from \eqref{e:final1} and \eqref{e:final2} that
\begin{equation*}
    u^\varepsilon(x)-u(x)-\left(1+\varepsilon^{\frac{\alpha^2}{\alpha+1}}\right)\frac{\nu C_\alpha  \varepsilon^{\alpha+1}}{d(x)^\alpha} \leq C\sqrt{\varepsilon}.
\end{equation*}
Therefore,
\begin{equation}
   -C \sqrt{\varepsilon}\leq  u^\varepsilon(x) - u(x) \leq C\sqrt{\varepsilon} + \left(1+\varepsilon^{\frac{\alpha^2}{\alpha+1}}\right)\frac{\nu C_\alpha  \varepsilon^{\alpha+1}}{d(x)^\alpha} 
\end{equation}
for every $x\in \Omega$. The final conclusions follow.
\end{proof}

\subsection{General bounded data}

We start with the following Lemma, which will be used later.
\begin{lem} Let $0<\kappa < \delta_0$ and $U_\kappa = \big\{x\in \Omega: 0<\mathrm{dist}(x,\partial\Omega) < \kappa\big\} = \Omega\backslash \overline{\Omega}_\kappa$. There holds
\begin{equation*}
    \mathrm{dist}(x,\partial\Omega_\kappa) = \kappa - \mathrm{dist}(x,\partial\Omega) \qquad\text{for all}\;x\in U_\kappa.
\end{equation*}
\end{lem}

\begin{proof} By definition of $\delta_0$ we have $d(x) = \mathrm{dist}(x,\partial\Omega)$ is twice continuously differentiable in the region $U_{\delta_0} = \Omega\backslash \overline{\Omega}_{\delta_0}$. The proof follows from \textcolor{blue}{\cite{} fill in Trudinger}. 
\end{proof}

\begin{lem} Let $u^\varepsilon\in \mathrm{C}^2(U_\kappa)$ be the unique solution to 
\begin{equation*}
    \left\{
        \begin{aligned}
            \tilde{u}^\varepsilon(x) + |D\tilde{u}^\varepsilon(x)|^p-\varepsilon \Delta \tilde{u}^\varepsilon(x) &=0 \;\qquad \text{in } U_\kappa ,\\
            \tilde{u}^\varepsilon(x) &= +\infty \quad \text{on } \partial U_\kappa = \partial \Omega\cup \partial \Omega_{\kappa},
        \end{aligned}
    \right.
\end{equation*}
and $u \equiv 0 $ be the unique solution to 
\begin{equation*}
    \left\{
        \begin{aligned}
            u(x)+ |Du(x)|^p &=0 \quad \text{in } U_\kappa ,\\
            u(x)+ |Du(x)|^p &\geq 0 \quad \text{on } \partial U_\kappa = \partial \Omega \cup\partial \Omega_{\kappa}.
        \end{aligned}
    \right.
\end{equation*}
Then there exists a constant $C$ independent of $\varepsilon$ such that
\begin{equation*}
  |u^\varepsilon(x) - u(x)| \leq C\varepsilon \quad  \text{for} \quad x\in \Omega_{\varepsilon}
\end{equation*}
where $C$ is independent of $\kappa$.
\end{lem}
\begin{rem} Unlike Theorem \ref{thm:rate_doubling0} where we made use of Lemma \ref{lem:super_refined}, the constant in the rate of convergence blows up as $\delta_{0,U_\kappa} \to 0$.
\end{rem}
%\nocite{*}

\clearpage

\section{Rate of convergence via nonlinear adjoint method}

\begin{appendices}
\section{Proofs of some preliminary results}
\begin{proof}[Proof of Theorem \ref{thm:grad_1}] Let $\theta\in (0,1)$ be chosen later, $\varphi\in \mathrm{C}_c^\infty(\Omega)$, $0\leq \varphi\leq 1$, $\mathrm{supp}\;\varphi\subset \Omega$ and $\varphi = 1$ on $\Omega_\delta$ such that
\begin{equation}\label{e:ass_power}
    |\Delta \varphi(x)| \leq C|\varphi(x)|^\theta \qquad\text{and}\qquad |D \varphi(x)|^2 \leq C\varphi^{1+\theta}
\end{equation}
for $x\in \Omega$ and $C = C(\delta,\theta)$.
Let $w(x) = |Du(x)|^2$ for $x\in \Omega$. The goal is to derive an equation for $\varphi w$. The equation for $w$ is given by
\begin{equation*}
    -\varepsilon \Delta w + p|D u|^{p-2}D u \cdot D w + 2  w - D f\cdot D u + \varepsilon |D^2u|^2 = 0 \qquad\text{in}\;\Omega.
\end{equation*}
Using that, we can derive an equation for $(\varphi w)$ as follows.
\begin{align*}
    &-\varepsilon \Delta (\varphi w) + p|D u|^{p-2}D u \cdot D (\varphi w) + 2  (\varphi w) + \varepsilon \varphi|D^2u|^2 + 2\varepsilon \frac{D \varphi}{\varphi}\cdot D (\varphi w) \\
    &\qquad\qquad\qquad\qquad = \varphi(D f\cdot D u) + p|D u|^{p-2}(D u\cdot D \varphi)w -\varepsilon w \Delta \varphi + 2\varepsilon \frac{|D \varphi|^2}{\varphi}w
    \qquad\text{in}\;\mathrm{supp}\;\varphi.
\end{align*}
Assume that $\varphi w$ achieves its maximum over $\overline{\Omega}$ at $x_0\in \Omega$, then we can assume that $x_0\in \mathrm{supp}\;\varphi$, since otherwise the maximum of $\varphi w$ over $\overline{\Omega}$ is zero. By the classical maximum principle 
\begin{equation*}
    -\varepsilon \Delta(\varphi w)(x_0)\geq 0 \qquad\text{and}\qquad |D(\varphi w)(x_0)| = 0.
\end{equation*}
Using that in the equation of $\varphi w$ above, we obtain that
\begin{equation*}
    \varepsilon \varphi|D^2u|^2 \leq  \varphi (Df\cdot Du)+ p|Du|^{p-1} |D\varphi|w + \varepsilon w |\Delta\varphi|  + 2\varepsilon  w\frac{|D\varphi|^2}{\varphi},
\end{equation*}
where all terms are evaluated at $x_0$. Using \eqref{e:ass_power} we have
\begin{equation}\label{e:est_for_D^2u}
    \varepsilon \varphi|D^2u|^2 \leq  \varphi |Df|w^{\frac{1}{2}}+ Cp w^{\frac{p-1}{2}+1} \varphi^{\frac{1+\theta}{2}} + C\varepsilon w \varphi^{\theta} + 2C\varepsilon  w\varphi^\theta.
\end{equation}
By Cauchy-Schwartz inquality, we have $n|D^2u|^2\geq (\Delta u)^2$, thus if $n\varepsilon < 1$ then
\begin{equation}\label{e:est_for_D^2u_2}
    \varepsilon |D^2u|^2 \geq \frac{(\varepsilon \Delta u)^2}{n\varepsilon} \geq (\varepsilon \Delta u)^2 = \left(  u + |Du|^p - f\right)^2 \geq |Du|^{2p} - 2C_0|Du|^p \geq \frac{|Du|^{2p}}{2} - 2K_0.
\end{equation}
Using \eqref{e:est_for_D^2u_2} in \eqref{e:est_for_D^2u} we obtain that
\begin{equation*}
    \varphi\left(\frac{1}{2}w^p - 2K_0\right) \leq \varphi |Df|w^{\frac{1}{2}}+ Cp w^{\frac{p-1}{2}+1} \varphi^{\frac{1+\theta}{2}} + 3C\varepsilon w \varphi^{\theta}.
\end{equation*}
Multiply both sides by $\varphi^{p-1}$ we deduce that
\begin{align*}
    (\varphi w)^p \leq 4K_0\varphi^{p-1} + 2\Vert Df\Vert_{L^\infty}\varphi^p w^{\frac{1}{2}} + 2Cp \varphi^{\frac{2p+\theta - 1}{2}}w^{\frac{p+1}{2}} + 6C\varepsilon \varphi^{p+\theta - 1}w.
\end{align*}
Choose $2p+\theta -1 \geq p+1$, i.e., $p+\theta\geq 2$. This is always possible with the requirement $\theta \in (0,1)$, as $1<p <\infty$. We deduce that 
\begin{equation}\label{e:est_for_D^2u_3}
    (\varphi w)^p \leq C\left(1+ (\varphi w)^\frac{1}{2} + (\varphi w)^\frac{p+1}{2} +(\varphi w)\right)
\end{equation}
As a polynomial in $z = (\varphi w)(x_0)$, this implies that $(\varphi w)(x_0)\leq C$ where $C$ depends on coefficients of the right hand side of \eqref{e:est_for_D^2u_3}, which implies our desired gradient bound since $\overline{\Omega}_\delta\subset \mathrm{supp}\;\varphi$.
\end{proof}

\begin{proof} [Proof of Theorem \ref{thm:wellposed1<p<2}] To find a candidate for subsolution and supersolution to \eqref{eq:PDEeps}, we use the \emph{ansatz}
\begin{equation}\label{eq:ansatz}
    u(x) = C_\varepsilon d(x)^{-\alpha}, \qquad x\in \Omega
\end{equation}
as we expect it blows up near the boundary like some thing proportionate to inverse of the distance function, due to the structure of $H(x,\varrho) = |\varrho|^p - f(x)$. Let us plug \eqref{eq:ansatz} into \eqref{eq:PDEeps} we obtain that 
\begin{align*}
      u(x) &\approx C_\varepsilon d(x)^{-\alpha},\\
    \frac{\partial u}{\partial x_i}(x) &\approx -\alpha C_\varepsilon d(x)^{-(\alpha+1)} \frac{\partial d}{\partial x_i}(x) \qquad\qquad\qquad \Longrightarrow\qquad |Du (x)|^p \approx \alpha^p C_\varepsilon^p d(x)^{-p(\alpha+1)}|D d(x)|^p,\\
    %\frac{\partial^2 u}{\partial x_i^2}(x) &\approx \alpha(\alpha+1)C_0 d(x)^{-(\alpha+2)}\left|\frac{\partial d}{\partial x_i}(x)\right|^2 -\alpha C_0 d(x)^{-(\alpha+1)}\frac{\partial ^2 d}{\partial x_i^2}(x),\\
    \varepsilon\Delta u(x) &\approx \frac{\varepsilon C_\varepsilon\alpha(\alpha+1)}{d(x)^{\alpha+2}}|D d(x)|^2 - \frac{\varepsilon C_\varepsilon\alpha}{d(x)^{\alpha+1}}\Delta d(x).
\end{align*}
As $|D d(x)| = 1$ for $x$ near $\partial\Omega$, we see that as $x\to \partial \Omega$, the highest explosive order terms are
\begin{equation*}
        -\varepsilon C_\varepsilon \alpha(\alpha+1)d^{-(\alpha+2)} + C_\varepsilon^p \alpha^p d^{-(\alpha+1)p}.
\end{equation*}
Setting them to zero, as $\alpha+1 = \frac{1}{p-1}$ we deduce that 
\begin{equation}\label{e:relation}
    \displaystyle\alpha = \frac{2-p}{p-1} \qquad\text{and}\qquad C_\varepsilon = \left(\frac{1}{\alpha}(\alpha+1)^\frac{1}{p-1}\right) \varepsilon^{\frac{1}{p-1}} = \frac{1}{\alpha}(\alpha+1)^{\alpha+1}\varepsilon^{\alpha+1}.
\end{equation}
Recall that $K_1 = \max_{\overline{\Omega}}|\Delta d(x)|$, let us denote
\begin{equation*}
    G = \left[\left(K_1\delta_0+(\alpha+1)\right)^{\alpha+1} - (\alpha+1)^{\alpha+1}\right]\alpha.
\end{equation*}

\begin{lem}\label{lem:subsln} For each $\delta > 0$ such that $0<2\delta < \delta_0$, if $\eta\geq G\varepsilon^{\alpha+1}$ then
\begin{equation*}
    \overline{w}_{\eta,\delta}(x) = \frac{C_\varepsilon+\eta}{(d(x)-\delta)^\alpha} +\frac{M_\varepsilon}{ },\qquad x\in \Omega_\delta, 
\end{equation*}
is a supersolution of \eqref{eq:PDEeps} in $\Omega_\delta$ where
\begin{equation}\label{e:M_eps}
    M_\varepsilon = \max_{\overline{\Omega}} f + 2\left(\frac{2}{\delta_0}\right)^{\alpha+1}K_1(\alpha+1)^{\alpha+1}\varepsilon^{\alpha+2} + \left(\frac{2(\alpha+1)\varepsilon}{\delta_0}\right)^{\alpha+2}\big(2^pK_2^p+2K_2^2\big).
\end{equation}
\end{lem}

\begin{proof} Recall that $p(\alpha+1) = \alpha+2$, we compute
\begin{align*}
    \mathcal{L}^\varepsilon\left[\overline{w}_{\eta,\delta}\right](x) = \frac{ (C_\varepsilon + \eta)}{(d(x)-\delta)^\alpha} &+ M_\varepsilon + \frac{(C_\varepsilon+\eta)^p \alpha^p}{(d(x)-\delta)^{\alpha+2}}|D d(x)|^p - f(x) \\
    &- \frac{\varepsilon(C_\varepsilon+\eta)\alpha(\alpha+1)}{(d(x)-\delta)^{\alpha+2}}|D d(x)|^2 + \frac{\varepsilon(C_\varepsilon+\eta)\alpha}{(d(x)-\delta)^{\alpha+1}}\Delta d(x).
\end{align*}
Regarding the relation \eqref{e:relation}, we define
\begin{equation*}
    1<\nu = \left(\frac{C_\varepsilon+\eta}{C_\varepsilon}\right) \leq 2.
\end{equation*}
There are two cases to be considered.
\begin{itemize}
    \item In the region where $d(x)\geq \delta_0$ then $d(x)-\delta \geq \frac{1}{2}\delta_0$, therefore 
    \begin{align*}
        \mathcal{L}^\varepsilon\left[\overline{w}_{\eta,\delta}\right](x) = \frac{ (C_\varepsilon + \eta)}{(d(x)-\delta)^\alpha} + M_\varepsilon &+ \frac{\alpha^pC_\varepsilon^p}{(d(x)-\delta)^{\alpha+2}}\left(\nu^{p}|Dd(x)|^p - \nu|Dd(x)|^2\right)\\
        &+\frac{\nu\varepsilon \alpha C_\varepsilon}{(d(x)-\delta)^{\alpha+1}}\Delta d(x) - f(x).
    \end{align*}
Recall that $K_2 = \max_{x\in\overline{\Omega}}|Dd(x)|$ and $1<|\nu|\leq 2$, we have
\begin{align*}
    \frac{\alpha^pC_\varepsilon^p}{(d(x)-\delta)^{\alpha+2}}\left|\nu^{p}|Dd(x)|^p - \nu|Dd(x)|^2\right| &\leq \left(\frac{2(\alpha+1)\varepsilon}{\delta_0}\right)^{\alpha+2}\big(2^pK_2^p+2K_2^2\big).
\end{align*}
%As $|Dd(x)|\leq 1$, $\nu^p|Dd(x)|^p - \nu|Dd(x)|^2 \geq 0$ since $1<p<2$ and $\nu>1$. Finally
\begin{equation*}
    \frac{\nu\varepsilon \alpha C_\varepsilon}{(d(x)-\delta)^{\alpha+1}}|\Delta d(x)| \leq 2\left(\frac{2}{\delta_0}\right)^{\alpha+1}K_1 \alpha\varepsilon C_\varepsilon \leq 2\left(\frac{2}{\delta_0}\right)^{\alpha+1}K_1 (\alpha+1)^{\alpha+1}\varepsilon^{\alpha+2}.
\end{equation*}
 From the definition of $M_\varepsilon$ in \eqref{e:M_eps} we deduce that $\mathcal{L}^\varepsilon\left[\overline{w}_{\eta,\delta}\right]\geq 0$ in this region.
    \item In the region where $\delta < d(x)\leq \delta_0$, we have $|Dd(x)|=1$, thus
    \begin{align*}
         \mathcal{L}^\varepsilon\left[\overline{w}_{\eta,\delta}\right](x) &= \frac{ (C_\varepsilon + \eta)}{(d(x)-\delta)^\alpha} + M - f(x)\\
         & + \frac{\alpha^p C_\varepsilon^p\nu }{(d(x)-\delta)^{\alpha+2}}\left(\nu^{p-1} - 1 - \frac{ (d(x)-\delta)\Delta d(x)}{(\alpha+1)}\right).
    \end{align*}
    We note that as $0<d(x)\leq \delta_0$, we have
    \begin{equation*}
        \left|\frac{(d(x)-\delta)\Delta  d(x)}{(\alpha+1)}\right| \leq {K_1\delta_0}{\alpha+1}.
    \end{equation*}
    Recall that $\nu = \left(1+\frac{\eta}{C_\varepsilon}\right)$, we want to choose $\eta$ such that
    \begin{equation*}
        \left(1+\frac{\eta}{C_\varepsilon}\right)^{p-1} - 1 \geq \frac{K_1\delta_0}{\alpha+1} \qquad\Longrightarrow\qquad \eta \geq C_\varepsilon\left[\left(1+\frac{K_1\delta_0}{\alpha+1}\right)^{\alpha+1}-1\right] = G\varepsilon^{\alpha+1}
    \end{equation*}
    where we use $\frac{1}{p-1} = \alpha+1$ again. This will ensure that $\mathcal{L}^\varepsilon\left[\overline{w}_{\eta,\delta}\right](x) \geq 0$ in this region.
\end{itemize}
\end{proof}

\begin{lem}\label{lem:supersln} For each $\eta\geq G\varepsilon^{\alpha+1}$ there exists $M_\eta > 0$ such that 
\begin{equation*}
    \underline{w}_{\eta, \delta}(x) = \frac{C_\varepsilon-\eta}{(d(x)+\delta)^\alpha} -\frac{M_\varepsilon}{ }, \qquad x\in \Omega^\delta 
\end{equation*}
is a subsolution of \eqref{eq:PDEeps} in $\Omega^\delta$.
\end{lem}

\begin{proof} The proof for this Lemma should be the same with Lemma \ref{lem:subsln}.
\end{proof}

\noindent We divide the rest of the proof into 3 steps. We first construct a minimal solution, then a maximal solution to \eqref{eq:PDEeps}, and finally show that they are equal to conclude the existence and uniqueness of solution to \eqref{eq:PDEeps}.

\begin{prop}\label{pro:minimalsol} There exists a minimal solution $\underline{u}\in \mathrm{C}^2(\Omega)$ of \eqref{eq:PDEeps} such that $v\geq \underline{u}$ for any other solution $v\in \mathrm{C}^2(\Omega)$ solving \eqref{eq:PDEeps}.
\end{prop}
\begin{proof} Let $w_{\eta,\delta}\in \mathrm{C}^2(\Omega)$ solves
    \begin{equation}\label{e:w_def}
    \begin{cases}
        \mathcal{L}\left[w_{\eta,\delta}\right] = 0 &\qquad\text{in}\;\Omega,\\
        \qquad w_{\eta,\delta} = \underline{w}_{\eta,\delta} &\qquad\text{on}\;\partial\Omega.
    \end{cases}
    \end{equation}
    \begin{itemize}
        \item Fix $\eta>0$ (and $\eta\geq G\varepsilon^{\alpha+1}$ in view of Lemmas \ref{lem:subsln} and \ref{lem:supersln}), as $\delta\to 0$ the value of $\underline{w}_{\eta,\delta}$ blows up on the boundary, therefore by the standard comparison principle for second-order elliptic equation with Dirichlet boundary we have $\delta_1 \leq  \delta_2$ implies $w_{\eta,\delta_1}\geq  w_{\eta,\delta_2}$ on $\overline{\Omega}$. 
        \item For $\delta'>0$, since $\underline{w}_{\eta,\delta'}$ is a subsolution in $\overline{\Omega}$ with finite boundary, we obtain that
            \begin{equation}\label{e:cp_delta1}
                0<\delta \leq \delta'\qquad\Longrightarrow\qquad \underline{w}_{\eta,\delta'} \leq w_{\eta_,\delta'}\leq w_{\eta,\delta} \qquad\text{on}\;\overline{\Omega}.
            \end{equation}
        \item Similarly, since $\overline{w}_{\eta,\delta'}$ is a supersolution on $\Omega_{\delta'}$ with infinity value on the boundary $\partial\Omega_{\delta'}$, by comparison principle
            \begin{equation}\label{e:cp_delta2}
                w_{\eta,\delta} \leq \overline{w}_{\eta, \delta'} \qquad\text{in}\;\Omega_{\delta'} \qquad\Longrightarrow\qquad w_{\eta,\delta} \leq \overline{w}_{\eta,0} \qquad\text{in}\;\Omega.
            \end{equation}
    \end{itemize}
    \noindent From \eqref{e:cp_delta1} and \eqref{e:cp_delta2} we have
    \begin{equation}\label{e:cp_delta3}
        0<\delta \leq \delta'\qquad\Longrightarrow\qquad \underline{w}_{\eta,\delta'} \leq w_{\eta_,\delta'}\leq w_{\eta,\delta} \leq \overline{w}_{\eta,0} \qquad\text{in}\;\Omega.
    \end{equation}
    Thus $\{w_{\eta,\delta}\}_{\delta>0}$ is locally bounded in $L^{\infty}_{\mathrm{loc}}(\Omega)$ ($\{w_{\eta,\delta}\}_{\delta>0}$ is uniformly bounded from below), thus using the local gradient estimate for $w_{\eta,\delta}$ solving \eqref{e:w_def} we deduce that $\{w_{\eta,\delta}\}_{\delta>0}$ is locally bounded in $W^{1,\infty}_{\mathrm{loc}}(\Omega)$. Plug it back into the defining equation \eqref{e:w_def} we deduce that $\{w_{\eta,\delta}\}_{\delta>0}$ is locally bounded in $W^{2,r}_{\mathrm{loc}}(\Omega)$ for all $r<\infty$.

    \noindent Local boundedness of $\{u_{\eta,\delta}\}_{\delta>0}$ in $W^{2,r}_{\mathrm{loc}}(\Omega)$ implies weak$^*$ compactness, that is there exists a function $u\in W^{2,r}(\Omega)$ such that (via subsequence and monotonicity)
    \begin{equation*}
        w_{\eta,\delta} \rup u \qquad\text{weakly in}\;W^{2,r}_{\mathrm{loc}}(\Omega),\qquad \text{and}\qquad
        w_{\eta,\delta} \to u \qquad\text{strongly in}\;W^{1,r}_{\mathrm{loc}}(\Omega).
    \end{equation*}
    In particular, $w_{\eta,\delta}\to u$ in $\mathrm{C}^1_{\mathrm{loc}}(\Omega)$ thanks to Sobolev compact embedding. Let us rewrite the equation $\mathcal{L}^\varepsilon\left[u_{\eta,\delta}\right] = 0$ as $\varepsilon\Delta w_{\eta,\delta}(x) = F[w_{\eta,\delta}](x)$ in $U$ for $U\subset\subset \Omega$ where
    \begin{equation*}
        F[w_{\eta,\delta}](x) =    w_{\eta,\delta}(x) + H(x,Dw_{\eta,\delta}(x)).
    \end{equation*}
    Since $u_{\eta,\delta}\to u$ in $\mathrm{C}^1(U)$ as $\delta\to 0$, we have $F[w_{\eta,\delta}](x) \to F(x)$ uniformly in $U$ where 
    \begin{equation*}
        F(x) =   u(x) + H(x,Du(x)).
    \end{equation*}
    In the limit as $\delta\to 0$ we obtain that $u\in L^2(U)$ is a weak solution of $\varepsilon\Delta u = F$ in $U$ where $F$ is continuous, thus $u\in \mathrm{C}^2(\Omega)$ and by stability $u$ solves $\mathcal{L}^\varepsilon[u] = 0$ in $\Omega$. From \eqref{e:cp_delta3} we also have
    \begin{equation*}
        \underline{w}_{\eta,0} \leq u \leq \overline{w}_{\eta,0} \qquad\text{in}\;\Omega.
    \end{equation*}
    It is clear that $u(x)\to \infty$ as $\mathrm{dist}(x,\partial\Omega)\to 0$ with the precise rate like \eqref{eq:ansatz}. We note that at this point, by construction, $u$ may depend on $\eta$. We show that $u$ is the maximal solution of $\mathcal{L}^\varepsilon[u] = 0$ in $\Omega$ such that $u = +\infty$ on $\partial\Omega$, thus consequently showing that $u$ is independent of $\eta$.

    \noindent Let $v\in W^{2,r}(\Omega)$ for all $r<\infty$ and $v$ solves \eqref{eq:PDEeps}. Fix $\delta>0$ then since $v(x)\to \infty$ as $x\to \partial\Omega$ while $w_{\eta,\delta}$ remains bounded on $\partial \Omega$, comparison principle yields
    \begin{equation*}
        v\geq w_{\eta,\delta} \qquad\text{in}\Omega.
    \end{equation*}
    Let $\delta\to 0$ we deduce that $v\geq u$ on $\Omega$. This concludes that $u$ is the minimal solution in $W^{2,r}(\Omega)(\forall\,r<\infty)$ and thus $u$ is independent of $\eta>0$. 
\end{proof}

\begin{prop} There exists a maximal solution $\overline{u}\in \mathrm{C}^2(\Omega)$ of \eqref{eq:PDEeps} such that $v\leq \overline{u}$ for any other solution $v\in \mathrm{C}^2(\Omega)$ solving \eqref{eq:PDEeps}.
\end{prop}

\begin{proof} For each $\delta>0$, let us denote $u_\delta\in \mathrm{C}^2(\Omega_\delta)$ be the minimal solution to $\mathcal{L}[u_\delta] = 0$ in $\Omega$ and $u_\delta = +\infty$ on $\partial\Omega_\delta$ according to Proposition \ref{pro:minimalsol}. By comparison principle, for every $\eta>0$ there holds
\begin{equation*}
    \underline{w}_{\eta,\delta} \leq u_\delta \leq \overline{w}_{\eta,\delta} \qquad\text{in}\;\Omega_\delta,
\end{equation*}
and
\begin{equation*}
    0<\delta<\delta' \qquad \Longrightarrow\qquad u_\delta \leq u_\delta' \qquad\text{in}\;\Omega_{\delta'}.
\end{equation*}
We have the local boundedness of $\{u_\delta\}_{\delta>0}$ in $W^{2,r}(\Omega)$ as well, thus together with monotonicity there exists $u\in W^{2,r}(\Omega)$ for all $r<\infty$ such that $u_\delta\to u$ strongly in $\rmC^1_{\mathrm{loc}}(\Omega)$, therefore using the equation $\mathcal{L}[u_\delta] = 0$ in $\Omega_\delta$ and the regularity of Laplace equation we deduce that $u\in \mathrm{C}^2(\Omega)$ solving \eqref{eq:PDEeps} and 
\begin{equation*}
    \underline{w}_{\eta,0} \leq u\leq \overline{w}_{\eta,0} \qquad\text{in}\;\Omega
\end{equation*}
for all $\eta>0$. From the construction, as $u_\delta$ is independent of $\eta$, it is clear that $u$ is also independent of $\eta$. We now show that $u$ is the maximal solution of \eqref{eq:PDEeps}. Let $v\in\rmC^2(\Omega)$ solves \eqref{eq:PDEeps}, then clearly $v\leq u_\delta$ on $\Omega_\delta$ and therefore as $\delta \to 0$ we have $v\leq u$.
\end{proof}
\noindent In conclusion we have found a minimal solution $\underline{u}$ and a maximal solution $\overline{u}$ in $\rmC^2(\Omega)$ such that
\begin{equation}\label{e:chain}
    \underline{w}_{\eta,0} \leq \underline{u}\leq \overline{u}\leq \overline{w}_{\eta,0} \qquad\text{in}\;\Omega
\end{equation}
for any $\eta>0$. This extra parameter $\eta$ now enables us to show that $\overline{u} = \underline{u}$ in $\Omega$. The key ingredient here is the convexity of $\varrho\mapsto H(x,\varrho)$.

\begin{prop}[Uniqueness] We have $\overline{u}\equiv \underline{u}$ in $\Omega$ and therefore solution to \eqref{eq:PDEeps} in $\mathrm{C}^2(\Omega)$ is unique.
\end{prop}

\begin{proof} Let $\theta\in (0,1)$. We define
\begin{equation*}
    w_\theta = \theta \overline{u} + (1-\theta) ^{-1}\left(\inf_{\Omega} f\right).
\end{equation*}
If $w_\theta$ is a subsolution to \eqref{eq:PDEeps}, then we can use comparison principle to obtain that
\begin{equation*}
    w_\theta = \theta \overline{u} + (1-\theta) ^{-1}\left(\inf_{\Omega} f\right) \leq \underline{u} \qquad\text{in}\;\Omega.
\end{equation*}
If that is the case, let $\theta\to 1$ we conclude that $\overline{u} \leq \underline{u}$. There is a problem here, as they are both explosive solutions, to use comparison principle we need to show that $w_\theta \leq \underline{u}$ in a neighborhood of $\partial\Omega$. From \eqref{e:chain} we see that
\begin{equation*}
    1\leq \frac{\overline{u}(x)}{\underline{u}(x)} \leq \frac{\overline{w}_{\eta,0}(x)}{\underline{w}_{\eta,0}(x)} = \frac{(C_0+\eta)+ ^{-1}C_\eta d(x)^\alpha}{(C_0-\eta)- ^{-1}C_\eta d(x)^\alpha} \qquad\text{for}\;x\in \Omega
\end{equation*}
and therefore
\begin{equation*}
   1\leq  \lim_{d(x)\to 0} \left(\frac{\overline{u}(x)}{\underline{u}(x)}\right) \leq \frac{C_0+\eta}{C_0-\eta} \qquad\Longrightarrow\qquad  \lim_{d(x)\to 0} \left(\frac{\overline{u}(x)}{\underline{u}(x)}\right) = 1
\end{equation*}
since $\eta>0$ is chosen arbitrary. This means for any $\varsigma\in(0,1)$ we can find $\delta(\varsigma)>0$ small such that on $\Omega\backslash \Omega_\delta$ one has
\begin{equation*}
\frac{\overline{u}(x)}{\underline{u}(x)}\leq (1+\varsigma)     \qquad\Longrightarrow\qquad \left(\frac{1}{1+\varsigma}\right)\overline{u}(x) \leq \underline{u}(x) \qquad\text{in}\; \Omega\backslash \Omega_\delta
\end{equation*}
For given $\theta\in (0,1)$, we can always choose $\varsigma$ small enough such that $(1+\varsigma)^{-1} > \theta$ which renders $\underline{u}(x) \geq \theta \overline{u}(x) + (1-\theta) ^{-1}\left(\inf_\Omega f\right)$ for $0< d(x) < \delta' < \delta$.
\end{proof}
\noindent This finishes the well-posedness of \eqref{eq:PDEeps} for $1<p<2$.
\end{proof}

\begin{proof}[Proof of Lemma \ref{lem:max}] The proof is a variation of Perron's method (see \cite{Capuzzo-Dolcetta1990}). Let $\varphi\in \rmC(\overline{\Omega})$ and $x_0\in \overline{\Omega}$ such that $u(x_0) = \varphi(x_0)$ and $u-\varphi$ has a global strict minimum over $\overline{\Omega}$ at $x_0$ and that 
\begin{equation}\label{eq:max_a1}
      \varphi(x_0) + H(x_0,D\varphi(x_0)) < 0.
\end{equation}
Let $\varphi^\varepsilon(x) = \varphi(x) - |x-x_0|^2 + \varepsilon$ for $x\in \overline{\Omega}$. Let $\delta > 0$, we see that for $x\in \partial B(x_0,\delta)\cap \overline{\Omega}$ then
\begin{equation*}
    \varphi^\varepsilon(x) = \varphi(x) - \delta^2 +\varepsilon \leq \varphi(x) - \varepsilon
\end{equation*}
if $2\varepsilon \leq \delta^2$. We observe that
\begin{equation*}
    \begin{split}
    \varphi^\varepsilon(x) - \varphi(x_0)  &= \varphi(x)-\varphi(x_0) + \varepsilon - |x-x_0|^2 \\
    D\phi^\varepsilon(x) - D\phi(x_0) &= D\varphi(x) - D\varphi(x_0) - 2(x-x_0)
    \end{split}
\end{equation*}
for $x\in B(x,\delta)\cap \overline{\Omega}$. We deduce from \eqref{eq:max_a1}, the continuity of $H(x,p)$ near $(x_0,D\varphi(x_0))$ and the fact that $\varphi\in \rmC^1(\overline{\Omega})$ that if $\delta$ is small enough and $0<2\varepsilon < \delta^2$ then
\begin{equation}\label{eq:max_a2}
      \varphi^\varepsilon(x)+H(x,D\varphi^\varepsilon(x)) < 0 \qquad\text{for}\;x\in B(x_0,\delta)\cap \overline{\Omega}.
\end{equation}
We have found $\phi^\varepsilon\in \mathrm{C}^1(\overline{\Omega})$ such that $\varphi^\varepsilon(x_0)>u(x_0)$, $\varphi^\varepsilon<u$ on $\partial B(x_0,\delta)\cap \overline{\Omega}$ and \eqref{eq:max_a2}. Let
\begin{equation*}
    \tilde{u}(x) = \begin{cases}
    \max \big\lbrace u(x),\phi^\varepsilon(x) \big\rbrace &x\in B(x_0,\delta)\cap \overline{\Omega},\\
    u(x)&x\notin B(x_0,\delta)\cap \overline{\Omega},\\
    \end{cases}
\end{equation*}
We see that $\tilde{u}\in \rmC(\overline{\Omega})$ is a subsolution of \eqref{S_0} in $\Omega$ with $\tilde{u}(x_0) > u(x_0)$, which is a contradiction, thus $u$ is a supersolution of \eqref{S_0} on $\overline{\Omega}$.
\end{proof}

\section{Differentiability with respect to the parameter}
\noindent For the vanishing viscosity problem with the Dirichlet boundary condition,
\begin{equation}
\label{dir}
\left\{
  \begin{aligned}
    H(x, Du^\epsilon(x)) &= \epsilon \Delta u^\epsilon \quad \, \text{in } U, \\
              u^\epsilon &= 0 \quad \qquad \text{on } \partial U,
  \end{aligned}
\right.
\end{equation}
where  $H(x,p)$ is $C^\infty(\overline{U}\times \mathbb{R}^n)$, $\displaystyle \frac{H(x,p)}{|p|} \to \infty$ uniformly in $x$ as $|p| \to \infty$ and $\displaystyle \sup_{x\in U}|D_xH(x,p)|\leq C(1+|p|)$, we want to show the smooth dependence of $u^\epsilon$ on $\epsilon$.
Formally, if we differentiate \eqref{dir} with respect to $\epsilon$, we get
\begin{equation}
\label{dir_dif}
\left\{
  \begin{aligned}
    D_pH(x, Du^\epsilon(x))\cdot Du^\epsilon_\epsilon &= \epsilon \Delta u^\epsilon_\epsilon +\Delta u^\epsilon \quad \, \text{in } U, \\
              u^\epsilon_\epsilon &= 0 \quad \qquad  \qquad  \text{on } \partial U.
  \end{aligned}
\right.
\end{equation}
By Schaefer's fixed point theorem and the maximal principle, $u^\epsilon_\epsilon$ is the unique solution in $C^{2,\alpha}(\overline{U})$ of \eqref{dir_dif}.

\noindent The main idea is, we look at the difference quotients $\displaystyle \frac{u^{\epsilon+h}-u^\epsilon}{h}$and prove that as $h \to 0^+$, they converge to a limiting function $w^{\ast}$ in the uniform norm  such that $w^{\ast}$ solves \eqref{dir_dif}. Since \eqref{dir_dif} has a unique solution, we have $$u^\epsilon_\epsilon=\lim_{h \to 0^+}\frac{u^{\epsilon+h}-u^\epsilon}{h}.$$

\subsection{Solution $u^\epsilon \in C^{2,\alpha}(\overline{U})$ exists} We use Schaufer's fixed point theorem as follows.

\begin{thm} Suppose $X$ is a Banach space. Let $A:X \to X$ be continuous and compact. Assume the set $\{u\in X : u=  A[u]$ for some $0 \leq    \leq 1\}$ is bounded. Then $A$ has a fixed point $u =A[u]$.
\end{thm}
\noindent Fix $0<\alpha<1$. Let $X=C^{1,\alpha}(\overline{U})$. Given $u \in X=C^{1,\alpha}(\overline{U})$, we look at the linear PDE
\begin{equation}
\label{fix}
\left\{
  \begin{aligned}
   \epsilon \Delta v &= H(x, Du) \quad \, \text{in } U, \\
              v &= 0 \qquad \qquad \text{on } \partial U.
  \end{aligned}
\right.
\end{equation}
\noindent
Estimate the Holder norm of RHS
$$\|H(x, Du)\|_{C^{0, \alpha}(\overline{U})}:=\sup_{x\in\overline{U}} |H(x,Du(x))| + \sup_{x, y \in \overline{U}}\frac{|H(x, Du(x))-H(y, Du(y))|}{|x-y|^\alpha}.$$
Since $Du$ is bounded and H is smooth,
\begin{equation}
    \sup_{x\in\overline{U}} |H(x,Du(x))| \leq C.
\end{equation}

\begin{equation}
    \begin{aligned}
  &\sup_{x, y \in \overline{U}}\frac{|H(x, Du(x))-H(y, Du(y))|}{|x-y|^\alpha}\\
  \leq &  \sup_{x, y \in \overline{U}}\frac{|H(x, Du(x))-H(y, Du(x))|}{|x-y|^\alpha} + \sup_{x, y \in \overline{U}}\frac{|H(y, Du(x))-H(y, Du(y))|}{|x-y|^\alpha} \\
  =&    \sup_{x, y \in \overline{U}}\frac{|\int_0^1D_xH(y+\theta (x-y), Du(x))d\theta \cdot (x-y)|}{|x-y|^\alpha}\\& +
     \sup_{x, y \in \overline{U}}\frac{|\int_0^1D_pH(y, Du(y)+\theta (Du(x)-Du(y)))d\theta \cdot (Du(x)-Du(y))|}{|x-y|^\alpha}\\
  \leq &  C\sup_{x, y \in \overline{U}} (1+|Du(x)|)|x-y|^{1-\alpha} +  C \sup_{x, y \in \overline{U}}\frac{| (Du(x)-Du(y))|}{|x-y|^\alpha}\\
 \leq & C(1 + \|u\|_{C^{1,\alpha}(\overline{U})})
    \end{aligned}
\end{equation}
since $Du$ is bounded on $\overline{U}$, $D_pH \in C^\infty (\overline{U} \times \mathbb{R}^n)$ and
$\displaystyle \sup_{x\in U}|D_xH(x,p)|\leq C(1+|p|)$. Therefore,
\begin{equation}
 \|H(x, Du)\|_{C^{0, \alpha}(\overline{U})} \leq C(1 + \|u\|_{C^{1,\alpha}(\overline{U})}).
\end{equation}
By Schauder estimates, there exists a unique solution $v \in C^{2,\alpha }(\overline{U})$ such that
\begin{equation}
\label{schauder}
    \|v\|_{C^{2, \alpha}(\overline{U})} \leq C  \|H(x, Du)\|_{C^{0, \alpha}(\overline{U})}. 
\end{equation}
\noindent Define operator $A$ on $X := C^{1, \alpha} (\overline{U}) $ by $A[u]=v$. So 
\begin{equation}
     \|A[u]\|_{C^{2, \alpha}(\overline{U})} \leq C(1 + \|u\|_{C^{1,\alpha}(\overline{U})}),
\end{equation}
\noindent
 and thus $A$ is continuous and compact. (i.e., if $\{u_k\}_{k=1}^\infty$ is bounded in $X=C^{1, \alpha}(\overline{U})$, then $\{A[u_k]\}_{k=1}\infty$ is bounded in $C^{2,\alpha} (\overline{U})$, thus precompact in $C^{1, \alpha}(\overline{U})$. Lemma 6.36 in Gilbarg and Trudinger. )

\noindent Next we try to bound $\{u\in X : u=  A[u]$ for some $0 \leq    \leq 1\}$. If $u=  A[u]$, the PDE becomes

\begin{equation}
\label{lamb}
\left\{
  \begin{aligned}
   \epsilon \Delta u &=   H(x, Du) \quad \, \text{in } U, \\
              u &= 0 \qquad \qquad \text{on } \partial U.
  \end{aligned}
\right.
\end{equation}

\noindent Calderon-Zygmund estimates tell us that if we have
\begin{equation}
\label{cald}
\left\{
  \begin{aligned}
  - \Delta v &=\Tilde{f}   \qquad \, \text{in } U, \\
              v &= 0 \qquad \text{on } \partial U.
  \end{aligned}
\right.
\end{equation}
and $\Tilde{f} \in {L^p(U)}$ for some $p\in (1, \infty)$, then $v \in W^{2,p}(U)$ and $$\|v\|_{w^{2,p}(U)} \leq C\|\Tilde{f}\|_{L^p(U)}.$$

\noindent Apply to \eqref{lamb} and we get

\begin{equation}
    \|u\|_{w^{2,p}(U)} \leq C\|H(x, Du)\|_{L^p(U)}
\end{equation}

(We want RHS to be bounded by some constant so that later we can choose $p$ larger than $n$ to  conclude $ \|u\|_{C^{1, \alpha}(U)} \leq   C\|u\|_{w^{2,p}(U)} \leq C$ by Morrey's estimate.) By a priori estimate, if we assume the solution of \eqref{dir} $u$ exists, then $\| Du\|_{L^\infty} \leq C_0$ where $C_0$ is independent of $\epsilon$. We can modify $H$ to get a new $\Tilde{H}$ so that it is smooth, $\Tilde{H} = H$ for $|p|<C_1$ and $H(x,p)=C_1+1$ for $|p|>C_1+1$ for some constant $C_1$ ($C_1$ is definitely larger than $C^0$). Moreover, the same prior estimate is correct for $\Tilde{H}$. Namely, if the solution $\Tilde{u}$ to \eqref{dir} with $H$ replaced by $\Tilde{H}$ exists, then $\| D\Tilde{u}\|_{L^\infty} \leq C_0$. (The way we choose this $C_1$ is that we work back from the beginning of the proof of a priori estimate for $u$, on both the boundary and interior of $\Omega$, and modify $H$ so that all the proofs go through and the same a priori estimate holds for the equation with $\Tilde{H}$.) Now with $\Tilde{H}$, we go through the same argument of Schaufer's fixed point theorem from the very beginning, and \eqref{cald} reads
\begin{equation}
       \|\Tilde{u}\|_{w^{2,p}(U)} \leq C\|\Tilde{H}(x, D\Tilde{u})\|_{L^p(U)} \leq C(1+\|D\Tilde{u}\|_{L^\infty}) \leq C.
\end{equation}
Choose $p=2n$ and $\displaystyle \alpha =\frac{1}{2}$. We have $\{\Tilde{u}\in X : \Tilde{u}=  A[\Tilde{u}]$ for some $0 \leq    \leq 1\}$ is bounded in $X= C^{1, \frac{1}{2}}(
\overline{U})$.
Thus Schaefer's fixed point theorem implies the equation \eqref{dir} with $H$ replaced by $\Tilde{H}$ has a solution $\Tilde{u} \in C^{2,\alpha} (\overline{\Omega})$. Since $\| D\Tilde{u}\|_{L^\infty} \leq C_0$, $\Tilde{u}$ also solves the original equation \eqref{dir}.

\subsection{Uniqueness}
Let $u$ and $v$ be two solutions to \eqref{dir} and $w := u-v$. Then we have
\begin{equation}
\begin{aligned}
    &-\epsilon \Delta (u-v) = H(x, Dv)-H(x, Du)\\
    \Rightarrow &-\epsilon \Delta w   =\int_0^1 D_pH(x, tDv+(1-t)Du) \cdot (Dv-Du)dt\\
    \Rightarrow &-\epsilon \Delta w  +\int_0^1 D_pH(x, tDv+(1-t)Du)dt \cdot Dw = 0.
\end{aligned}
\end{equation}
By the strong maximum principle, $w \equiv 0$.

\subsection{Smooth dependence on $\epsilon$}
Fix $\epsilon >0$. Let  $$w^h(x):=\frac{u^{\epsilon+h}(x)-u^\epsilon(x)}{h} \in C^{2,\alpha}(\overline{U}).$$

A little computation shows that $w^h$ solves

\begin{equation}
\label{dir_quo}
\left\{
  \begin{aligned}
   \epsilon \Delta w^h(x) + \frac{\epsilon}{\epsilon + h}\Delta u^\epsilon &= \frac{\epsilon}{\epsilon +h} \int_0^1 D_pH(x, Du^\epsilon+\theta (Du^{\epsilon+h}-Du^\epsilon)) d\theta \cdot Dw^h \quad \, \text{in } U, \\
              w^h &= 0 \quad \qquad \text{on } \partial U.
  \end{aligned}
\right.
\end{equation}
From the existence proof, we know $\|u^\epsilon\|_{C^{2,\alpha}(\overline{U})} \leq C$ uniformly in $\epsilon$. So $\|Du^{\epsilon+h}-Du^\epsilon\|_{C^{0,\alpha}(\overline{U})}$and $\|\Delta u\|_{C^{0,\alpha}(\overline{U})}
$ is uniformly bounded in $h$.

By Schauder estimates, $\{w^h\}_{h>0} \subset C^{2, \alpha}(\overline{U})$ are bounded, hence is precompact in $ C^{2, \beta}(\overline{U})$ for any $\beta < \alpha$. Therefore, there exits a subsequence $\{w^{h_j}\}_{j=1}^\infty$ such that $w^{h_j} \to w^\ast$ for some $w^\ast \in  C^{2, \beta}(\overline{U}) $ and $w^\ast$ solves \eqref{dir_dif}. This implies $w^h \to w^\ast$ in $C^{2, \beta}(\overline{U})$.
\end{appendices}

\section*{Acknowledgement}
The authors would like to express their appreciation to Hung V. Tran for his invaluable guidance.

\bibliography{zzzzlibrary}{}
%\bibliographystyle{ieeetr}
\bibliographystyle{acm}

\end{document}