### 2021-07-15 Drafts and notes/Draft_07.tex

\documentclass[12pt,reqno]{amsart}
%============%============%============%============%

%============%============%============%============%
%\setlength{\columnseprule}{0.4pt}
%\setlength{\topmargin}{0cm}
%\setlength{\oddsidemargin}{.25cm}
%\setlength{\evensidemargin}{.25cm}
%\setlength{\textheight}{22.5cm}
%\setlength{\textwidth}{15.5cm}
\renewcommand{\baselinestretch}{1.05}
%============%============%============%============%
\usepackage[toc,page]{appendix}
\newcommand{\stoptocwriting}{%
  \addtocontents{toc}{\protect\setcounter{tocdepth}{-5}}}
\newcommand{\resumetocwriting}{%
  \addtocontents{toc}{\protect\setcounter{tocdepth}{\arabic{tocdepth}}}}

%\setcounter{secnumdepth}{4}
%\setcounter{tocdepth}{3}
%============%============%============%============%
%\usepackage{romannum}
\usepackage{xcolor}
\usepackage{placeins}
\usepackage{amsfonts,amsmath,amsthm}
\usepackage{amssymb,epsfig}
\usepackage{enumerate} 
\usepackage[notcite,notref]{showkeys}
\usepackage{fullpage}
%============%============%============%============%
\usepackage[utf8]{inputenc}
\usepackage{mathpazo}

%\usepackage[libertine,cmintegrals,cmbrac,vvarbb]{newtxmath}
\usepackage{eucal}
%\usepackage[utopia]{mathdesign}
\usepackage[euler-digits]{eulervm}
\usepackage[unicode=true]{hyperref}
\hypersetup{colorlinks = true}
%\hypersetup{hidelinks=true}
\hypersetup{
     colorlinks,
     linkcolor={black!10!red},
     linkbordercolor = {black!100!red},
%    <your other options...>,
     citecolor={blue}
}

%graphic
%\usepackage[text={425pt,650pt},centering]{geometry}

\usepackage{pdfsync}

\usepackage{geometry}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm,headheight=3.5cm}

\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{tikz}
\usepackage{caption}
\usepackage{color} %color
\definecolor{vert}{rgb}{0,0.6,0}

\usepackage{comment}
\numberwithin{figure}{section}
%\pagestyle{plain}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{ass}{Assumption}
\renewcommand{\theass}{}
\newtheorem{defn}{Definition}
\newtheorem{quest}{Question}
\newtheorem{com}{Comment}
\newtheorem{ex}{Example}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{remark}
\newtheorem{rem}{\bf{Remark}}
\numberwithin{equation}{section}

%\renewcommand{\thefootnote}{\fnsymbol{footnote}}

%Characters -- Shortcuts
\newcommand{\E}{\mathbb{E}}
\newcommand{\M}{\mathbb{M}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\bP}{\mathbb{P}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\bS}{\mathbb{S}}
\newcommand{\T}{\mathbb{T}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\bfS}{\mathbf{S}}
\newcommand{\cA}{\mathcal{A}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cL}{\mathcal{L}}
\newcommand{\cM}{\mathcal{M}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\cS}{\mathcal{S}}
\newcommand{\cT}{\mathcal{T}}
\newcommand{\cE}{\mathcal{E}}
\newcommand{\I}{\mathrm{I}}

%Functional spaces
\newcommand{\AC}{{\rm AC\,}}
\newcommand{\ACl}{{\rm AC}_{{\rm loc}}}
\newcommand{\BUC}{{\rm BUC\,}}
\newcommand{\USC}{{\rm USC\,}}
\newcommand{\LSC}{{\rm LSC\,}}
\newcommand{\Li}{L^{\infty}}
\newcommand{\Lip}{{\rm Lip\,}}
\newcommand{\W}{W^{1,\infty}}
\newcommand{\Wx}{W_x^{1,\infty}}

%Domains
\newcommand{\bO}{\partial\Omega}
\newcommand{\cO}{\overline\Omega}
\newcommand{\Q}{\mathbb{T}^{n}\times(0,\infty)}
\newcommand{\iQ}{\mathbb{T}^{n}\times\{0\}}
\newcommand{\cQ}{\mathbb{T}^{n}\times[0,\infty)}

%Greek alphabets -- Shortcuts
\newcommand{\al}{\alpha}
\newcommand{\gam}{\gamma}
\newcommand{\del}{\delta}
\newcommand{\ep}{\varepsilon}
\newcommand{\kap}{\kappa}
\newcommand{\lam}{ }
\newcommand{\sig}{\sigma}
\newcommand{\om}{\omega}
\newcommand{\Del}{\Delta}
\newcommand{\Gam}{\Gamma}
\newcommand{\Lam}{ }
\newcommand{\Om}{\Omega}
\newcommand{\Sig}{\Sigma}

%Overlines, Underlines -- Shortcuts
\newcommand{\ol}{\overline}
\newcommand{\ul}{\underline}
\newcommand{\pl}{\partial}
\newcommand{\supp}{{\rm supp}\,}
\newcommand{\inter}{{\rm int}\,}
\newcommand{\loc}{{\rm loc}\,}
\newcommand{\co}{{\rm co}\,}
\newcommand{\diam}{{\rm diam}\,}
\newcommand{\diag}{{\rm diag}\,}
\newcommand{\dist}{{\rm dist}\,}
\newcommand{\Div}{{\rm div}\,}
\newcommand{\sgn}{{\rm sgn}\,}
\newcommand{\tr}{{\rm tr}\,}
\newcommand{\Per}{{\rm Per}\,}

\newcommand{\rmC}{\mathrm{C}}
\newcommand{\rup}{\rightharpoonup}

\renewcommand{\subjclassname}{%
\textup{2010} Mathematics Subject Classification}

%Hyperlink in PDF file
%\usepackage[dvipdfm,
%  colorlinks=false,
%  bookmarks=true,
%  bookmarksnumbered=false,
%  bookmarkstype=toc]{hyperref}
%\makeatletter
%\def\@pdfm@dest#1{%
%  \Hy@SaveLastskip
%  \@pdfm@mark{dest (#1) [@thispage /\@pdfview\space @xpos @ypos null]}%
%  \Hy@RestoreLastskip
%}

%BibLatex
%\usepackage[
%backend=biber,
%style=alphabetic,
%sorting=ynt
%]{biblatex}
%\addbibresource{rate.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{import}
\usepackage{xifthen}
\usepackage{pdfpages}
\usepackage{transparent}
\newcommand{\incfig}[1]{%
    \def\svgwidth{\columnwidth}
    \import{./figs/}{#1.pdf_tex}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title[Rate of convergence]
{\textsc{Remarks on the vanishing viscosity process of state-constraint Hamilton--Jacobi equations}}
\thanks{The authors are supported in part by NSF grant DMS-1664424 and NSF CAREER grant DMS-1843320. The work of Son N. T. Tu is supported in part by the GSSC Fellowship, University of Wisconsin--Madison.}
\begin{abstract}
We investigate the convergence rate in the vanishing viscosity process of the solutions to the state-constraint Hamilton-Jacobi equations. We give two different proofs of the fact that, for Lipschitz data that vanish on the boundary, the rate of convergence is $\mathcal{O}(\sqrt{\varepsilon})$ in the interior. Moreover, the one-sided rate can be improved to $\mathcal{O}(\varepsilon)$ for compactly supported data. Our approach relies on deep understanding of the blow-up behavior near the boundary and semiconcavity of the solutions.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Yuxi Han}
\address[Y. Han]
{
Department of Mathematics, 
University of Wisconsin Madison, 480 Lincoln  Drive, Madison, WI 53706, USA}
\email{yuxi.han@wisc.edu}
\author{Son N. T. Tu}
\address[S. N.T. Tu]
{
Department of Mathematics, 
University of Wisconsin Madison, 480 Lincoln  Drive, Madison, WI 53706, USA}
\email{thaison@math.wisc.edu}
\date{\today}
\keywords{first-order Hamilton--Jacobi equations; second-order Hamilton--Jacobi equations; state-constraint problems; optimal control theory; rate of convergence; viscosity solutions; semiconcavity; boundary layer.}
\subjclass[2010]{
35B40, %Asymptotic behavior of solutions, 
35D40, %Viscosity solutions
49J20, %Optimal control problems involving partial differential equations
49L25, %Viscosity solutions
70H20 %Hamilton-Jacobi equations
}
\maketitle
\setcounter{tocdepth}{1}
%\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
\subsection{Settings} Let $\Omega$ be an open, bounded and connected domain in $\mathbb{R}^n$ with $\mathrm{C}^2$ boundary, $f\in \mathrm{C}(\overline{\Omega})\cap W^{1,\infty}(\Omega)$. For $\varepsilon>0$, let $u^\varepsilon\in \mathrm{C}^2(\Omega)$ (see \cite{Lasry1989}) be the solution to
 \begin{equation}\label{eq:PDEepsa}
    \begin{cases}
      u^\varepsilon(x) + H(Du^\varepsilon(x)) - f(x) - \varepsilon \Delta u^\varepsilon(x) = 0 \qquad
    \text{in}\;\Omega, \vspace{0cm}\\
    \displaystyle  \lim_{\mathrm{dist}(x,\partial \Omega)\to 0} u^\varepsilon(x) = +\infty
    \end{cases} 
\end{equation}
 where $H:\R^n\to\R^n$ is a given continuous Hamiltonian. The solution that blows up uniformly on the boundary is also called a \emph{large solution}. A typical subquadratic Hamiltonian that has been considered in the literature is $H(\xi) = |\xi|^p$ for $\xi\in \R^n$ where $1<p\leq 2$. We focus on this Hamiltonian in our paper, and the equation of interest becomes
\begin{equation}\label{eq:PDEeps}
    \begin{cases}
      u^\varepsilon(x) + |Du^\varepsilon(x)|^p - f(x) - \varepsilon \Delta u^\varepsilon(x) = 0 \qquad
    \text{in}\;\Omega, \vspace{0cm}\\
    \displaystyle  \lim_{\mathrm{dist}(x,\partial \Omega)\to 0} u^\varepsilon(x) = +\infty.
    \end{cases} \tag{PDE$_\varepsilon$}
\end{equation}
When $1<p\leq 2$, equation \eqref{eq:PDEeps} describes the value function associated with a minimization problem in stochastic optimal control with state constraint (\cite{fabbri_stochastic_2017, Lasry1989}. We are interested in studying the asymptotic behavior of $\{u^\varepsilon\}_{\varepsilon>0}$ as $\varepsilon\rightarrow 0^+$. Heuristically, the solution of the second-order state-constraint problem converges to that of a first-order state-constraint problem associated with the deterministic optimal control, namely,
\begin{equation}\label{eq:PDE0}
    \begin{cases}
       u(x) + |Du(x)|^p - f(x) \leq 0\;\qquad\text{in}\;\Omega,\\
       u(x) + |Du(x)|^p - f(x) \geq 0\;\qquad\text{on}\;\overline{\Omega},
    \end{cases} \tag{PDE$_0$}
\end{equation}
as is described in the framework of the theory of viscosity solutions. Equation \eqref{eq:PDE0} admits a unique viscosity solution in the space $\rmC(\overline{\Omega})$, which is also the maximal viscosity subsolution among all the viscosity subsolutions in $\rmC(\overline{\Omega})$(see \cite{Capuzzo-Dolcetta1990,Soner1986}). The problem is interesting since in the limit we no longer have blowing up behavior. In this paper, we investigate the rate of convergence of $u^\varepsilon \to u$ as $\varepsilon\to 0^+$. What is intriguing and delicate here is the blow-up behavior of $u^\varepsilon$ in a narrow strip near $\partial \Omega$ as $\varepsilon\to 0^+$. This is often called the boundary layer theory in the literature.

Note that a comparison principle holds for \eqref{eq:PDE0} since we always assume $\Omega$ is an open, bounded and connected domain in $\mathbb{R}^n$ with $\mathrm{C}^2$ boundary (\cite{Capuzzo-Dolcetta1990,Soner1986}). Equation \eqref{eq:PDEeps} follows the setting of \cite{Lasry1989}, where the specific structure of the Hamiltonian $H(x,\xi) = |\xi|^p - f(x)$ enables more explicit estimates for the solution of \eqref{eq:PDEeps}.

%\subsection{Assumptions} We will always assume $\Omega$ is an open, bounded and connected domain in $\mathbb{R}^n$ with $\mathrm{C}^2$ boundary. Essentially the domain is assumed so that a comparison principle holds for \eqref{eq:PDE0}.

%\begin{itemize}    
%    \item[(A1)] There exists a universal pair of positive numbers $(r,h)$ and $\eta\in \mathrm{BUC}(\overline{\Omega};\mathbb{R}^n)$ such that $B(x+t\eta(x), rt)\subset\Omega$ for all $t\in (0,h]$,
%    \end{itemize}
%    or
%    \begin{itemize}
%    \item[(A2)] $\Omega$ is a bounded star-shaped (with respect to the origin) open subset of $\mathbb{R}^n$ such that for some $\kappa > 0$,
%    \begin{equation*}
%        \mathrm{dist}(x,\overline{\Omega}) \geq \kappa r \qquad\text{for all}\; x\in (1+r) \partial\Omega, \;\text{for all}\;r>0.
%    \end{equation*}
%\end{itemize}

 %Assumption $\mathrm{(A1)}$ or $\mathrm{(A2)}$ guarantees that a comparison principle holds for the state-constraint problem \eqref{eq:PDE0}. Assumption $\mathrm{(A1)}$, which holds for any smooth bounded domain, was first studied in \cite{Soner1986}. Assumption $\mathrm{(A2)}$ was posed and studied in \cite{Capuzzo-Dolcetta1990} and the references therein. 
%We list the assumptions on a general Hamiltonian as follow.
%\begin{itemize}
%    \item[$\mathrm{(H1)}$] There exists $C_1 > 0$ such that $H(x,p) \geq -C_1$ for all $(x,p)\in \overline{\Omega}\times\mathbb{R}^n$.
%    \item[$\mathrm{(H2)}$] There exists $C_2>0$ such that $|H(x,0)|\,\leq \,C_2$ for all $(x,p)\in \overline{\Omega}\times \mathbb{R}^n$.
%    \item[$\mathrm{(H3)}$] For each $R>0$ there exists a modulus $\omega_{R}[0,\infty)\to [0,\infty)$ such that $\omega_R(0^+) = 0$ and 
%    \begin{equation*}
%        \begin{cases}
%        |H(x,p) - H(y,p)| \leq \omega_R(|x-y|),\\
%        |H(x,p) - H(x,q)| \leq \omega_R(|p-q|),
%        \end{cases} \qquad\text{for all}\;x,y\in \overline{\Omega}, p,q \in \mathbb{R}^n\;\text{with}\;|p|,|q|\leq R.
%    \end{equation*}
%    \item[$\mathrm{(H4)}$] $H(x,p)\rightarrow \infty$ as $|p|\to \infty$ uniformly in $x\in \overline{\Omega}$.
%\end{itemize}

\subsection{Relevant literature} There is a
vast amount of work in the literature on viscosity solutions with state constraints and large solutions. We would like to first mention that the problem \eqref{eq:PDE0} with general Hamiltonian is a huge subject of research interest, started with the pioneer work \cite{Soner1986} (see also \cite{ishii_new_1996,ishii_class_2002}).  Some of the recent work related to the asymptotic behavior of solutions of \eqref{eq:PDE0} can be found in \cite{ishii_vanishing_2017,kim_state-constraint_2020,mitake_asymptotic_2008,tu2021vanishing}. The problem \eqref{eq:PDEeps} is first studied in \cite{Lasry1989} and subsequently many works have been done in understanding deeper the properties of solutions (see \cite{ Porretta_a,alessio_asymptotic_2006,Bandle_1994, marcus_existence_2003} and the references therein). The time-dependent version of \eqref{eq:PDEepsa} is also studied by many works, for instance, \cite{barles_generalized_2004,barles_large_2010,leonori_local_2011,moll_large_2012} and the references therein. %For deterministic and stochastic optimal control problems with state constraints, we refer the read to [\textcolor{blue}{NEED CITATION (Ishii, \cite{}}] and the references therein.

In terms of rate of convergence, that is, the convergence rate of $u^\varepsilon\to u$ as $\varepsilon\to 0^+$, to the best of our knowledge, such a question has not been studied in the literature. For the case where \eqref{eq:PDEeps} is equipped with the Dirichlet boundary condition, a rate $\mathcal{O}(\sqrt{\varepsilon})$ is well known with multiple proofs (see \cite{Bardi1997,crandall1984, tran_hamilton-jacobi_2021}). We believe part of the reason why the problem is intriguing is that the blow-up behavior of the solutions near the boundary of $\Omega$ is complicated and deserves further investigation.

\subsection{Main results} The main results of the paper are the following theorems.

\begin{thm}\label{main_thm1} Let $\Omega$ be an open, bounded and connected subset of $\R^n$ with $\mathrm{C}^2$ boundary. Assume that $1 < p\leq 2$ and $f$ is Lipschitz with $f = 0$ on $\partial\Omega$. Let $u^\varepsilon$ be the unique solution to \eqref{eq:PDEeps} and $u$ be the unique solution to \eqref{eq:PDE0}. Then there exists a constant $C$ independent of $\varepsilon\in (0,1)$ such that for $x\in \Omega$,
\begin{align*}
    &-C\sqrt{\varepsilon}\leq  u^\varepsilon(x) - u(x)\leq C\sqrt{\varepsilon} + C\varepsilon \left(\frac{\varepsilon}{d(x)}\right)^{\frac{2-p}{p-1}} &\qquad\text{if}\; p < 2,\\
    &-C\sqrt{\varepsilon}\leq  u^\varepsilon(x) - u(x)\leq C\sqrt{\varepsilon} + \varepsilon|\log(d(x))| &\qquad\text{if}\; p = 2.
\end{align*}
\end{thm}

\begin{rem} To the best of our knowledge, this theorem is new in the literature. We don't know if the order of $\varepsilon$ on the right hand side of the inequalities above is optimal or not near the boundary and this deserves further investigation. The condition $f = 0$ on $\partial\Omega$ is a little bit restrictive but can be explained as follows. One can see that the solution to \eqref{eq:PDEeps} is continuous with respect to data $f$ in the weak$^*$ topology of $L^\infty(\Omega)$ (see \cite{Lasry1989}). If $f = 0$ on $\partial\Omega$, we are able to approximate $f$ \emph{uniformly} in $L^\infty(\Omega)$ by a sequence of compactly supported functions, where a convergence rate of $u^\varepsilon-u$ is easier to obtain. Indeed, it is natural to consider the doubling variable method with
\begin{equation}\label{heur1}
    \psi^\varepsilon(x) =  u^\varepsilon(x) - \frac{C_\alpha\varepsilon^{\alpha+1}}{d(x)^{\alpha}}
\end{equation}
and $u(x)$, where $d(x)$ is a $\mathrm{C}^2$ extension of the distance function $\mathrm{dist}(x,\partial\Omega)$, $\alpha = (2-p)/(p-1)$ and $C_\alpha\varepsilon^{\alpha+1}d(x)^{-\alpha}$ is the leading order term in the asymptotic expansion of $u^\varepsilon(x)$ as $d(x) \to 0^{+}$. If we take the derivative of \eqref{heur1} formally, it becomes
\begin{equation}\label{heur2}
    D\psi^\varepsilon(x) =  Du^\varepsilon(x) + C_\alpha\alpha \left(\frac{\varepsilon}{d(x)}\right)^{\alpha+1}Dd(x).
\end{equation}
We will see that $D\psi^\varepsilon(x)$ is uniformly bounded if $d(x)\geq \varepsilon$ (Lemma \ref{lem:boundDu^eps}). Indeed, 
\begin{equation*}
    -C_\alpha \alpha \left(\frac{\varepsilon}{d(x)}\right)^{\alpha+1}Dd(x)    
\end{equation*}
is more or less the leading order term in the asymptotic expansion of $Du^\varepsilon$ near $\partial\Omega$. Heuristically, this means that the boundary layer is $\mathcal{O}(\varepsilon)$ from the boundary.

However, to get a useful estimate by the doubling variable, at the maximum point $x_0$ of $\psi^\varepsilon(x) - u(x)$, we need to have $d(x_0)\geq \varepsilon^{\gamma}$ for $\gamma<1$ so that the latter term in \eqref{heur2} vanishes as $\varepsilon\to 0^+$. In the other case where $d(x_0) < \varepsilon^{\gamma}$, we introduce a new localization idea, that is, we construct a blow-up solution in the ball of radius $\varepsilon^\gamma$ from the boundary. Finally, a technical (and common for the doubling variable) computation leads to $\gamma = 1/2$.

%Essentially we avoid the boundary layer here.

%In the complicated case that $x_0$ is in the boundary layer, we introduce a new localization idea, that is, we construct a blow-up solution in the ball of radius $\varepsilon$.
%\textcolor{red}{(Actually, I am thinking should it be the ball of radius $\sqrt{\varepsilon}$ from the boundary? Since we choose $\kappa=\sqrt{\varepsilon}$.)}
\end{rem}

As a different approach, the convexity of $|\xi|^p$ and the semiconcavity of the solution to \eqref{eq:PDE0} give us a better one-sided $\mathcal{O}(\varepsilon)$ estimate for compactly supported $f$. Such an one-sided $\mathcal{O}(\varepsilon)$ rate is well known for the case of the Dirichlet boundary problem (see \cite{Bardi1997,tran_hamilton-jacobi_2021}). Moreover, this approach provides us with a new proof of the $\mathcal{O}(\sqrt{\varepsilon})$ rate in Theorem \ref{main_thm1} without using the doubling variable method.

%As a different approach, if we utilize the convexity of $|\xi|^p$ and the semiconcavity of the solution to \eqref{eq:PDE0} then we have a proof of the $\mathcal{O}(\sqrt{\varepsilon})$ rate without using the doubling variable method. More importantly, this approach gives a better one-sided $\mathcal{O}(\varepsilon)$ estimate for compactly supported $f$. Such an one-sided $\mathcal{O}(\varepsilon)$ rate is well known for the case of Dirichlet boundary problem (see \cite{Bardi1997,tran_hamilton-jacobi_2021}).

\begin{thm}[One-sided $\mathcal{O}(\varepsilon)$ rate for compactly supported data]\label{thm:rate_doubling2} Assume that $f\in \mathrm{C}^2_c(\Omega)$ such that $\mathrm{supp}(f)\subset\Omega_\kappa$ where  $\Omega_\kappa := \{x\in \Omega: \mathrm{dist}(x,\Omega) > \kappa\}$ for some fixed $\kappa \in (0, \delta_{0})$. Let $u^\varepsilon$ be the unique solution to \eqref{eq:PDEeps} and $u$ be the unique solution to \eqref{eq:PDE0}. Then there exists a constant $C$ independent of $\varepsilon$ and $\kappa$ such that
\begin{equation*}
\begin{aligned}
    u^\varepsilon(x) - u(x) &\leq\frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x)^\alpha} + C\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1} \left(1+ \left(\frac{\varepsilon}{\kappa}\right) +\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1} +\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+2}\right) + nc\varepsilon,  &\quad p <2, \\
    u^\varepsilon(x) - u(x) & \leq \nu \varepsilon \log\left( \frac{1}{d(x)}\right)+C\left(\frac{\varepsilon}{\kappa}\right) \left(1+\frac{\varepsilon}{\kappa}\right)+ nc \varepsilon, &\quad p=2.
\end{aligned}
\end{equation*}
where $c = \max \big\lbrace  D_{\xi\xi}f(x): |\xi|=1, x\in \overline{\Omega} \big\rbrace\geq 0$.
\end{thm}

\begin{rem} As mentioned, Theorem \ref{thm:rate_doubling2} implies Theorem \ref{main_thm1}. While the second approach looks more powerful, we need to know more about the gradient bound $Du^\varepsilon$ (Lemma \ref{lem:boundDu^eps}). On the other hand, the first approach using doubling variable is relatively simple and does not require any explicit asymptotic behavior of $Du^\varepsilon$ except the fact that it is locally bounded. At this moment, we do not know how to generalize to the more general case where $f = 0$ on $\partial\Omega$ in Theorem \ref{thm:rate_doubling2}.
\end{rem}

\subsection*{Organization of the paper} Section 2 contains some preliminary results. The proof of Theorem \ref{main_thm1} is given in Section 3. Then in Section 4, we give the proof of Theorem \ref{thm:rate_doubling2} and how Theorem \ref{main_thm1} follows from Theorem \ref{thm:rate_doubling2}. Finally, the proofs of some useful lemmas are presented in Appendix.

\section{Preliminaries}\label{sec:prelim} 
Let $\Omega$ be an open, bounded and connected subset of $\mathbb{R}^n$ with boundary $\partial\Omega$ of class $C^2$. For small $\delta>0$, we denote $\Omega_\delta = \{x\in \Omega: \mathrm{dist}(x,\Omega) > \delta\}$ and $\Omega^\delta = \{x\in \mathbb{R}^n: \mathrm{dist}(x,\overline{\Omega}) < \delta\}$. 
\begin{figure}[ht]
    \centering
    %\incfig{Domains}
    \def\svgwidth{0.35\columnwidth}
    \import{./figs/}{Domains.pdf_tex}
    \caption{The domain $\Omega$ and its variations $\Omega_\delta, \Omega^\delta$.}
    \label{fig:Domains}
\end{figure}

\begin{defn} Define
\begin{equation}\label{def:delta_0}
    \delta_{0,\Omega} =\frac{1}{2}\sup \big\{ \delta > 0: x\mapsto\mathrm{dist}(x,\partial\Omega)\;\text{is}\;\rmC^2\;\text{in}\;\Omega\backslash\overline{\Omega}_{\delta} \big\}.
\end{equation}
%The distance function is of class $\mathrm{C}^2$ in the region where $0<\mathrm{dist}(x,\partial\Omega) < \delta_0$. 
We will simply write $\delta_0$ instead of $\delta_{0,\Omega}$ when the underlying domain is understood.
\end{defn}

We refer the reader to \cite{gilbarg_elliptic_2001} for the regularity of the distance function defined in the definition above. We then extend $\mathrm{dist}(x,\partial\Omega)$ to a function $d(x)\in \mathrm{C}^2(\mathbb{R}^n)$ such that 
\begin{equation}\label{e:distance_def}
    \begin{cases}
    d(x)\geq 0\;\text{for}\;x\in\Omega\;\text{with}\;d(x) = +\mathrm{dist}(x,\partial\Omega)\;\text{for}\;x\in \Omega\backslash \Omega_{\delta_0},\\
    d(x)\leq 0\;\text{for}\;x\notin \Omega\;\text{with}\;d(x) = -\mathrm{dist}(x,\partial\Omega)\;\text{for}\;x\in \Omega^{\delta_0}\backslash \Omega.
    \end{cases}
\end{equation}
Note that $d(x) = \mathrm{dist}(x,\partial\Omega)$ and $|D d(x)| = 1$ in the classical sense in $\Omega^{\delta_0}\backslash \Omega_{\delta_0}$. Let 
\begin{equation}\label{boundond}
   K_0:= \max_{x\in \overline{\Omega}}|d(x)|, \qquad K_1 := \max_{x\in \overline{\Omega}} |D d(x)| \qquad\text{and}\qquad K_2 := \max_{x\in \overline{\Omega}} |\Delta d(x)|.
\end{equation}
\noindent Denote by $\mathcal{L}^\varepsilon:\rmC^2(\Omega)\to \rmC(\Omega)$ the operator
\begin{equation*}
    \mathcal{L}^\varepsilon[u](x) :=   u(x) + |Du(x)|^p - f(x) - \varepsilon \Delta u(x), \qquad x\in \Omega.
\end{equation*}

\subsection{Local gradient estimate} For $\varepsilon \in (0,1)$ and $p>1$, we state a priori estimate for $\rmC^2$ solutions to \eqref{eq:PDEeps} (\cite[Appendix]{Lasry1989}). Since we are working with nice solutions, the proof is relatively simple using the classical Bernstein method (\cite{bernstein_sur_1910}). We provide a proof in Appendix for the reader's convenience.

\begin{thm}\label{thm:grad_1} Let $f\in \rmC(\overline{\Omega})\cap W^{1,\infty}(\Omega)$ and $u^\varepsilon \in \mathrm{C}^2(\Omega)$ be a solution to $\mathcal{L}^\varepsilon[u^\varepsilon] = 0$ in $\Omega$. Let $m:= \max_{\overline{\Omega}}f(x)$. Then for $\delta>0$, there exists $C_\delta = C(m,p,\delta, \Vert D f\Vert_{L^\infty(\Omega)})$ such that 
\begin{equation*}
    \sup_{x\in \overline{\Omega}_\delta} \Big(|u^\varepsilon(x)|+|Du^\varepsilon(x)|\Big) \leq C_\delta
\end{equation*}
for $\varepsilon$ small enough.
\end{thm}

%\textcolor{orange}{(I think here we want to add for $\varepsilon$ small enough?)}

\subsection{Well-posedness of \eqref{eq:PDEeps}} In this section, we recall the existence and the uniqueness of solutions to \eqref{eq:PDEeps} for $1<p\leq 2$ and $f\in \mathrm{C}(\overline{\Omega})\cap W^{1,\infty}(\Omega)$. In fact, the assumption of $f$ can be relaxed to $f\in L^\infty(\Omega)$ (\cite{Lasry1989}).

\begin{thm}\label{thm:wellposed1<p<2} Let $f\in \mathrm{C}(\overline{\Omega})\cap W^{1,\infty}(\Omega)$. There exists a unique solution $u^\varepsilon\in \mathrm{C}^2(\Omega)$ of \eqref{eq:PDEeps} such that:
\begin{itemize}
    \item[(i)] If $1<p< 2$, then 
\begin{equation}\label{rate_p<2}
    %\frac{C_\varepsilon-\eta}{d(x)^\alpha}-\frac{M_\eta}{ } \leq u(x)\leq \frac{C_\varepsilon+\eta}{d(x)^\alpha}+\frac{M_\eta}{ } 
    \lim_{d(x)\to 0}\left( u^\varepsilon(x) \,d(x)^\alpha \right)= C_\alpha \varepsilon^{\alpha+1}
\end{equation}
where $\alpha = (p-1)^{-1}(2-p)$ and $C_\alpha = \alpha^{-1}(\alpha+1)^{\alpha+1}$.
%\begin{equation*}
%    \displaystyle\alpha = \frac{2-p}{p-1} \qquad\text{and}\qquad C_\alpha = \frac{1}{\alpha}(\alpha+1)^{\alpha+1}.
%\end{equation*}
\item[(ii)] If $p=2$, then
\begin{equation}\label{rate_p=2}
    \lim_{d(x)\to 0} \left(-\frac{u^\varepsilon(x)}{\log(d(x))}\right) = \varepsilon.
\end{equation}
\end{itemize}
Furthermore, $u^\varepsilon$ is the maximal subsolution among all the subsolutions $v\in W^{2,r}_{\mathrm{loc}}(\Omega)$ for all $r\in [1,\infty)$ of \eqref{eq:PDEeps}.
\end{thm}
\noindent This is Theorem I.1 in \cite{Lasry1989} with an explicit dependence on $\varepsilon$. A proof of this theorem is carried out explicitly in Appendix for later use. Also, it is useful to note that $\alpha+1 = (p-1)^{-1}$. More results on the behavior of the gradient of $u^\varepsilon$ can be found in \cite{alessio_asymptotic_2006} and Lemma \ref{lem:boundDu^eps}, where we show $|Du^\varepsilon|\leq C + C\left(\frac{\varepsilon}{d(x)}\right)^{\alpha+1}$. We believe Lemma \ref{lem:boundDu^eps} is new in the literature.

\subsection{Convergence results} We first state the following Lemma (\cite{Capuzzo-Dolcetta1990}), which characterizes the solution to the first-order state-constraint  equation \eqref{eq:PDE0}.
\begin{lem}\label{lem:max} Let $u\in \rmC(\overline{\Omega})$ be a viscosity subsolution of \eqref{eq:PDE0} such that, for any viscosity subsolution $v\in \rmC(\overline{\Omega})$ of \eqref{eq:PDE0}, one has $v\leq u$ on $\overline{\Omega}$. Then $u$ is a viscosity supersolution of \eqref{eq:PDE0} on $\overline{\Omega}$.
\end{lem}
\noindent Again, a proof of Lemma \ref{lem:max} is given in Appendix for the reader's convenience.

\begin{lem}\label{lem:lower-bound} Assume $1<p\leq 2$. Let $u^\varepsilon\in \mathrm{C}^2(\Omega)$ be the solution to \eqref{eq:PDEeps}. We have $\{  u^\varepsilon\}_{\varepsilon>0}$ is uniformly bounded from below by a constant independent of $\varepsilon$. More precisely, $  u^\varepsilon \geq \min_\Omega f$ and $  u\geq \min_\Omega f$.
%\begin{equation}\label{e:lower_bound_u_eps}
%    \inf_{x\in \Omega}   u^\varepsilon(x) \geq  \inf_{\Omega} f
%\end{equation}
%and also
%\begin{equation}\label{e:lower_bound_u}
%    \inf_{x\in \Omega}   u(x) \geq  \inf_{\Omega} f
%\end{equation}
\end{lem}
\begin{proof} For $m\in \mathbb{N}$, let $u^{\varepsilon,m}\in \mathrm{C}^2(\Omega)\cap \rmC(\overline{\Omega})$ solve the Drichlet problem
\begin{equation}\label{e:uepsm}
    \begin{cases}
      u(x) + |Du(x)|^p - f(x) - \varepsilon \Delta u(x) = 0 &\qquad
    \text{in}\;\Omega, \vspace{0cm}\\
    \quad \qquad\quad\qquad\qquad\qquad\qquad u(x) = m &\qquad
    \text{on}\;\partial\Omega.
    \end{cases} \tag{PDE$_{\varepsilon,m}$}
\end{equation}
We have $u^{\varepsilon}_m(x) \to u^\varepsilon(x)$ in $\Omega$ as $m\to \infty$. Let $\varphi(x) \equiv  \inf_{\Omega} f$ for $x\in \overline{\Omega}$. Then $\varphi(x)$ is a classical subsolution of \eqref{e:uepsm} in $\Omega$ with
\begin{equation*}
    \varphi(x) =   \inf_\Omega f \leq m = u^\varepsilon_m(x) \qquad \text{on } \partial \Omega,
\end{equation*}
for $m$ large enough.
By the comparison principle of the uniformly elliptic equation \eqref{e:uepsm},
\begin{equation*}
     \inf_\Omega f \leq u^{\varepsilon}_m(x) \qquad\text{for all}\;x\in \Omega.
\end{equation*}
As $m\to \infty$, we obtain $  u^\varepsilon \geq \min_\Omega f$. The inequality $  u\geq \min_{\Omega}f$ follows from the comparison principle of \eqref{eq:PDE0} applied to the supersolution $u$ on $\overline{\Omega}$ and the subsolution $\varphi$ in $\Omega$.
\end{proof}

We present here a simple proof of the convergence $u^\varepsilon \to u$ using Lemma \ref{lem:max}. See also \cite[Theorem VII.3]{Capuzzo-Dolcetta1990}.
\begin{thm}[Vanishing viscosity]\label{thm:qual}. Let $u^\varepsilon$ be the solution to \eqref{eq:PDEeps}. Then there exists $u \in \mathrm{C}(\overline{\Omega})$ such that $u^\varepsilon \rightarrow u$ locally uniformly in $\Omega$ as $\varepsilon\rightarrow 0$ and $u$ solves \eqref{eq:PDE0}.
\end{thm}

\begin{proof}[Proof of Theorem \ref{thm:qual}] By a priori estimate (Theorem \ref{thm:grad_1}), 
\begin{equation}\label{e:priorie_eps}
    |u^\varepsilon(x)| + |Du^\varepsilon(x)| \leq C_\delta \qquad\text{for}\;x\in \overline{\Omega}_\delta.
\end{equation}
By the Arzel\`a--Ascoli theorem, there exists a subsequence $\varepsilon_j\to 0$ and a function $u\in \rmC(\Omega)$ such that $u^{\varepsilon_j}\to u$ locally uniformly in $\Omega$. By the stability of viscosity solutions, we easily deduce that 
\begin{equation}\label{eq:u0int}
     u(x) + |Du(x)|^p - f(x) = 0 \qquad\text{in}\;\Omega.
\end{equation}
From Lemma \ref{lem:lower-bound}, $  u^\varepsilon(x)\geq \min_{\Omega} f$ and  $  u(x)\geq \min_{\Omega} f$ for all $x\in \Omega$. Together with \eqref{eq:u0int}, we obtain $|\xi|\leq \max_\Omega f - \min_\Omega f$ for all $\xi\in D^+u(x)$ and $x\in \Omega$. This implies there exits a constant $C_0$ such that
\begin{equation}\label{e:C0}
    |u(x) - u(y)| \leq C_0|x-y| \qquad\text{for all}\;x,y\in \Omega.
\end{equation}
Thus, we can extend $u$ uniquely to $u\in \rmC(\overline{\Omega})$. We use Lemma \ref{lem:max} to show that $u$ is a supersolution of \eqref{eq:PDE0} on $\overline{\Omega}$.

It suffices to show that $u\geq w$ on $\overline{\Omega}$, where $w\in \rmC(\overline{\Omega})$ is the unique solution to \eqref{eq:PDE0}. For $\delta>0$, let $u_\delta\in\rmC(\overline{\Omega}_\delta)$ be the  unique viscosity solution to
\begin{equation}\label{e:v_v}
    \begin{cases}
      u_\delta(x) + |Du_\delta(x)|^p-f(x) \leq 0 &\qquad\text{in}\;\Omega_\delta,\\
      u_\delta(x) + |Du_\delta(x)|^p - f(x) \geq 0 &\qquad\text{on}\;\overline{\Omega}_\delta.
    \end{cases}
\end{equation}
Since $u_\delta\rightarrow w$ locally uniformly as $\delta\rightarrow 0^+$ (see \cite{kim_state-constraint_2020}) and $w$ is bounded, $\{u_\delta\}_{\delta>0}$ is uniformly bounded. Let $v^\varepsilon_\delta\in \rmC^2(\Omega_\delta)\cap \rmC(\overline{\Omega}_\delta)$ be the unique solution to the Dirichlet problem
\begin{equation}\label{eq:vv_eps}
\begin{cases}
      v_\delta^\varepsilon(x) + |Dv_\delta^\varepsilon(x)|^p - f(x) = \varepsilon \Delta v_\delta^\varepsilon(x) &\qquad\text{in}\;\Omega_\delta,\\
    \;\quad\qquad\qquad\qquad\qquad v_\delta^\varepsilon = u_\delta &\qquad \text{on}\;\partial\Omega_\delta.
\end{cases}
\end{equation}
It is well-known that $v^\varepsilon_\delta\to u_\delta$ uniformly on $\overline{\Omega}_\delta$ as $\varepsilon\to 0$.

For $\delta$ small enough, $u_\delta\leq u^\varepsilon$ on $\partial \Omega_\delta$. Hence by the maximum principle, $v^\varepsilon_\delta \leq u^\varepsilon$ on $\overline{\Omega}_\delta$. Now we first let $\varepsilon\to 0$ to obtain $u_\delta \leq u$ on $\overline{\Omega}_\delta$.
Then let $\delta\rightarrow 0$ to get $w\leq u$ in $\Omega$, which implies $w\leq u$ on $\overline{\Omega}$ since both $w,u$ belong to $\rmC(\overline{\Omega})$.
\end{proof}

\section{Rate of convergence}
In this section, we focus on the rate of convergence for the case where $f\in \mathrm{W}^{1,\infty}(\Omega)\cap\mathrm{C}(\overline{\Omega})$ with an extra assumption to be described. By adding a constant to the solution $u^\varepsilon$ of \eqref{eq:PDEeps}, without loss of generality, we can assume the \emph{nonnegativity} of $f$, i.e.,
\begin{equation}\label{e:minf}
    \min_{\overline{\Omega}} f(x) = 0.
\end{equation}
As a consequence, $u^\varepsilon(x),u(x)\geq 0$ for $x\in \Omega$ by Lemma \ref{lem:lower-bound}. Before we show any result about the rate of convergence, we would like to mention a lower bound of $u^\varepsilon - u$ and some properties of $u$ from its optimal control formulation.
\begin{thm} Let $u^\varepsilon$ be the unique solution to \eqref{eq:PDEeps} and $u$ be the unique solution to \eqref{eq:PDE0}. Then there exists a constant $C$ independent of $\varepsilon$ such that
\begin{equation}\label{e:lower1}
    -C\sqrt{\varepsilon} \leq u^\varepsilon(x) - u(x) \qquad\text{for all}\;x\in \Omega.
\end{equation} 
\end{thm}
\begin{proof} The proof relies on a well-known rate of convergence for vanishing viscosity of the viscous Hamilton--Jacobi equation with the Dirichlet boundary condition (see \cite{crandall1984,evans_adjoint_2010,fleming_convergence_1961,Tran2011}). Let $g(x) = u(x)$ for $x\in \partial\Omega$. Let $v^\varepsilon\in \mathrm{C}^2(\Omega)\cap \mathrm{C}(\overline{\Omega})$ be the unique viscosity solution to
\begin{equation*}
    \begin{cases}
          v^\varepsilon(x) + |Dv^\varepsilon(x)|^p - f(x) - \varepsilon \Delta v^\varepsilon(x) = 0 &\qquad\text{in}\;\Omega,\\
        \qquad\qquad\qquad\qquad\qquad \qquad\quad v^\varepsilon(x) = g(x) &\qquad\text{on}\;\partial\Omega.
    \end{cases}
\end{equation*}
It is well-known that $v^\varepsilon \to u$. Furthermore, there exists a positive constant $C$ independent of $\varepsilon\in (0,1)$ such that 
\begin{equation}\label{e:cp1}
     |v^\varepsilon(x)  - u(x)| \leq C\sqrt{\varepsilon} \qquad\text{for}\;x\in \overline{\Omega}.
\end{equation}
By the comparison principle for \eqref{eq:PDEeps}, we have
\begin{equation}\label{e:cp2}
    v^\varepsilon(x)\leq u^\varepsilon(x) \qquad\text{for}\;x\in \Omega.
\end{equation}
From \eqref{e:cp1} and \eqref{e:cp2}, we obtain the lower bound \eqref{e:lower1}.
\end{proof}

%We state some observations  regarding the property of the solution $u$ to \eqref{eq:PDE0}.

\begin{lem}\label{lem:f=0} Assume $f\geq 0$ in $\Omega$. Then $u(x) = 0$ if and only if $f(x) = 0$. In particular, $f \equiv 0$ implies $u \equiv 0$.
\end{lem}
\begin{proof} From Lemma \ref{lem:lower-bound}, we know $u\geq 0$ in $\overline{\Omega}$. The optimal control formula for the solution $u$ to \eqref{eq:PDE0} is (see \cite{Bardi1997,tran_hamilton-jacobi_2021})
\begin{equation}\label{e:foru}
    u(x) = \inf \left\lbrace \int_0^\infty e^{-  s}\Big(c|\dot{\eta}(s)|^{q} +f(\eta(s))\Big)ds: \eta\in \mathrm{AC}([0,\infty);\overline{\Omega}), \eta(0) = x\right\rbrace
\end{equation}
where 
\begin{equation*}
    c = \left(qp^\frac{1}{p-1}\right)^{-1} \qquad\text{and}\qquad \frac{1}{p} + \frac{1}{q} = 1.
\end{equation*}
%\textcolor{red}{(Here, the constant $c=\frac{1}{qp^\frac{1}{p-1}}$? )}

Let $x\in \overline{\Omega}$ such that $f(x) = 0$. We can choose $\eta(s) = x$ for all $s\in [0,\infty)$ as an admissible path in \eqref{e:foru} to obtain that
\begin{equation*}
    u(x) \leq  \int_0^\infty e^{-  s}\Big(c|\dot{\eta}(s)|^{q} +f(\eta(s))\Big)ds = f(x) =  0.
\end{equation*}
As a consequence, $f\equiv 0$ implies $u\equiv 0$.

It is not hard to prove the converse by contradiction. Suppose $u(x_0) =0$ and $f(x_0)>0$. Then there exists $\varepsilon$, $\delta >0$ such that $f(x) > \varepsilon$ for all $x \in B_{\delta} (x_0)$. Let $\eta \in AC([0,\infty);\overline{\Omega})$ such that $\eta(0)=x_0$ and t be the time that $\eta$ first hits $\partial B_{\delta}(x_0)$. Note that $t$ could be $+\infty$. Then
\begin{equation*}
    \begin{aligned}
        \int_0^\infty e^{-s} \left(|\dot{\eta}(s)|^q +f(s )\right) ds &\geq \int_0^t e^{-s} \left(|\dot{\eta}(s)|^q +f(s )\right) ds\\
        &\geq  \frac{1}{e^t t^{q-1}}\left| \int_0^t \dot{\eta} (s) ds\right|^q+ \varepsilon \left(1-e^{-t} \right)\\
        &\geq \frac{\delta^q}{e^t t^{q-1}}+\varepsilon \left(1-e^{-t} \right),
    \end{aligned}
\end{equation*}
where we used Jensen's inequality in the second line.
This implies $u(x_0)>0$ since $q \geq 2$, which is a contradiction.
\end{proof}
\noindent The following lemma is about a crucial estimate that will be used. It is a refined construction of a supersolution for \eqref{eq:PDEeps}.
\begin{lem}\label{lem:super_refined} Let $\delta_0$ be defined as in \eqref{def:delta_0}. There exist positive constants $\nu = \nu(\delta_0)> 1$ and $C_\nu =\mathcal{O}\left(\delta_0^{-(\alpha+2)}\right)$ such that
\begin{equation}\label{e:superwa}
w(x) = \begin{cases}
    \displaystyle\frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x)^\alpha} + \max f + C_\nu \varepsilon^{\alpha+2}, \qquad\;\;\,  p<2,\vspace{0.2cm}\\
    \displaystyle\nu \varepsilon \log\left(\frac{1}{d(x)}\right) + \max f+ C_\nu \varepsilon^2, \qquad  p=2,
\end{cases}
\end{equation}
is a supersolution of \eqref{eq:PDEeps} in $\Omega$. 
%\item[(ii)] If $x\mapsto \mathrm{dist}(x,\partial\Omega)$ is continuously differentiable with $|D \mathrm{dist}(x,\partial\Omega)| = 1$ everywhere in $\Omega$ then we can take $C_\nu = 0$ in \eqref{e:superwa}.
\end{lem}
%\textcolor{orange}{(Why is $\nu \in (1,2)$? I think when $\alpha$ is large, it can be larger than 2. But $\nu$ is certainly bounded.---Oh, you can choose $\delta_0$ so small that $\nu \in (1,2)$. But then $\delta_0$ depends on $\alpha$.)} -  \textcolor{red}{I think dependence on $\alpha$ is fine, as we fix $p$ to start with.} \textcolor{orange}{- Do we include that in the definition of $\delta_0$?} --  \textcolor{red}{I will just write $\nu$ for now, not requiring $\nu \leq 2$ anymore.}

\begin{proof} Let us first consider $1<p<2$. Recall from Theorem \ref{thm:wellposed1<p<2} that $C_\alpha^p \alpha^p = C_\alpha \alpha (\alpha+1)$ and $p(\alpha+1) = \alpha+2$. Compute 
\begin{equation*}
    |D w(x)|^p = \nu^p\frac{(C_\alpha\alpha)^p\varepsilon^{p(\alpha+1)}}{d(x)^{p(\alpha+1)}} \left| Dd(x) \right|^p= \nu^p\frac{C_\alpha \alpha(\alpha+1)\varepsilon^{\alpha+2}}{d(x)^{\alpha+2}}  \left| Dd(x) \right|^p
\end{equation*}
and
\begin{equation*}
    \varepsilon\Delta w(x) = \nu\frac{C_\alpha\alpha(\alpha+1)\varepsilon^{\alpha+2}}{d(x)^{\alpha+2}} \left| Dd(x)\right|^2- \nu\frac{C_\alpha\alpha\varepsilon^{\alpha+2}\Delta d(x)}{d(x)^{\alpha+1}}.
\end{equation*}
We have
\begin{equation*}
    \begin{aligned}
        \mathcal{L}^\varepsilon\left[ w \right] =  &\frac{\nu C_\alpha\varepsilon^{\alpha+1}}{d(x)^\alpha} + \max f - f(x) + C_\nu \varepsilon^{\alpha+2} \\&+ \frac{C_\alpha \alpha (\alpha+1)\varepsilon^{\alpha+2}}{d(x)^{\alpha+2}}\left[\nu^p\left| Dd(x) \right|^p-\nu \left| Dd(x)\right|^2+\nu\frac{d(x)\Delta d(x)}{\alpha+1}\right] .
    \end{aligned}
\end{equation*}
\paragraph{\textbf{Case 1.}} If $\delta< d(x)\leq \delta_0$, we have $|D d(x)| = 1$. Recall that $K_2= \Vert \Delta d\Vert_{L^\infty}$ and observe
\begin{equation*}
    \left|\frac{d(x)\Delta d(x)}{\alpha+1}\right| \leq \frac{\delta_0\Vert \Delta d\Vert_{L^\infty}}{\alpha+1} \leq \frac{K_2\delta_0}{\alpha+1}\leq K_2\delta_0.
\end{equation*}
Therefore,
\begin{equation}\label{e:choose_nu}
    \begin{split}
        \nu^p-\nu +\nu\frac{d(x) \Delta d(x)}{(\alpha+1)} \geq \nu^p - \nu - \nu K_2\delta_0 = \nu\Big(\nu^{p-1} - (1+K_2\delta_0)\Big).
    \end{split}
\end{equation}
We will choose $\nu$ as follows. For $\gamma>1$, we have the inequality
\begin{equation}\label{e:ineq}
    \Big||x+y|^\gamma - |x|^\gamma\Big|\leq \gamma\Big(|x|+|y|\Big)^{\gamma-1}|y|
\end{equation}
for $x,y\in \mathbb{R}$, which implies that
\begin{equation*}
     0 \leq (1+K_2\delta_0)^{\alpha+1} - 1 \leq\underbrace{(\alpha+1)\left(1+K_2\delta_0\right)^\alpha K_2}_{C_2}\delta_0.
\end{equation*}
Hence $(1+K_2\delta_0)^{\alpha+1} \leq 1 + C_2\delta_0$. Since $\alpha+1 = \frac{1}{p-1}$,
\begin{equation}\label{e:cru1}
    (1+K_2\delta_0) \leq (1+C_2\delta_0)^{\frac{1}{\alpha+1}} = (1+C_2\delta_0)^{p-1}. 
\end{equation}
Choose $\nu = 1+C_2\delta_0$ in \eqref{e:choose_nu} and we obtain $\mathcal{L}[w]\geq 0$ in $\{x\in \Omega_\delta: \delta <d(x)\leq \delta_0\}$. \\

\paragraph{\textbf{Case 2.}} If $d(x)\geq \delta_0$, recall that $K_0 = \Vert d\Vert_{L^\infty}$ and $K_1 = \Vert D d\Vert_{L^\infty}$, and we have
\begin{align*}
    \mathcal{L}[w] &= \frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x)^\alpha} + \max_{\Omega} f - f(x)\\
    &+  \nu^p\frac{C_\alpha\alpha(\alpha+1)\varepsilon^{\alpha+2}}{d(x)^{\alpha+2}}|D d(x)|^p - \nu \frac{C_\alpha \alpha(\alpha+1)\varepsilon^{\alpha+2}}{d(x)^{\alpha+2}}|D d(x)|^2 
    + \nu \frac{C_\alpha \alpha \varepsilon^{\alpha+2}\Delta d(x)}{d(x)^{\alpha+1}} + C_\nu \varepsilon^{\alpha+2}\\
    &\geq \frac{C_\alpha \alpha(\alpha+1)\varepsilon^{\alpha+2}}{d(x)^{\alpha+2}}\left(\nu^p|D d(x)|^p - \nu |D d(x)|^2 + \nu \frac{d(x) \Delta d(x)}{\alpha+1}\right) + C_\nu \varepsilon^{\alpha+2}\\
    &\geq \left[C_\nu - C_3\left(\frac{1}{\delta_0}\right)^{\alpha+2}\right]\varepsilon^{\alpha+2}
\end{align*}
where 
\begin{equation*}
    C_3 = C_\alpha\alpha(\alpha+1) \left(\nu^pK_1^p + \nu K_1^2 + \nu \frac{K_0K_2}{\alpha+1}\right).
\end{equation*}
We can choose $C_\nu = C_3\delta_0^{-(\alpha+2)}$ to obtain $\mathcal{L}[w]\geq 0$ in $\{x\in \Omega_\delta:d(x)\geq \delta_0\}$. 
\smallskip

\noindent
If $p=2$, then $\alpha = 0$. We can easily see that the similar calculation holds true with $\nu := 1+K_2\delta_0$ and $C_\nu := \delta_0^{-2}\nu (\nu K_1^2 + K_1^2+ K_0K_2)$.
\end{proof}

Now we begin to present the rate of convergence for the special case where $f = C_f$ in $\Omega$ for some constant $C_f$.

\begin{thm}[Constant data]\label{thm:rate_doubling0} Assume $f\equiv C_f$ in $\Omega$. Let $u^\varepsilon$ be the unique solution to \eqref{eq:PDEeps} and $u \equiv C_f$ be the unique solution to \eqref{eq:PDE0}. Then there exists a constant $C$ independent of $\varepsilon\in(0,1)$ such that 
    \begin{equation*}
    \begin{split}
    &0\leq u^\varepsilon(x) - u(x)\leq C \left(\frac{ \varepsilon^{\alpha+1}}{d(x)^\alpha} + \frac{\varepsilon^{\alpha+2}}{\delta_{0,\Omega}^{\alpha+2}}\right),  \qquad\qquad \;\;\; \text{if}\; 1<p<2,\\
    &0\leq u^\varepsilon(x) - u(x)\leq C \left(\varepsilon \mathrm{log}\left(\frac{1}{d(x)}\right) + \frac{\varepsilon^{2}}{\delta_{0,\Omega}^{2}}\right),  \qquad \text{if}\; p=2,
    \end{split}
\end{equation*}
for $x\in \Omega$, where $\delta_{0,\Omega}$ is defined as in \eqref{def:delta_0}. In particular, 
\begin{itemize}
    \item[(i)] if $1<p<2$, we have $C_f\leq u^\varepsilon(x)\leq C_f + C\varepsilon$ for $x\in \Omega_\varepsilon = \{x\in \Omega: \mathrm{dist}(x,\partial\Omega) \geq \varepsilon\}$, and
    \item[(ii)] for any $K\subset\subset \Omega$ there hols $\Vert u^\varepsilon - u\Vert_{L^\infty(K)} \leq C\varepsilon^{\alpha+1}$ .
\end{itemize}
\end{thm}

\begin{proof} Lemma \ref{lem:f=0} implies $u \equiv C_f$ in $\Omega$. And Lemma \ref{lem:lower-bound} tells us $u^\varepsilon-u=u^\varepsilon-C_f \geq 0$.By the comparison principle of \eqref{eq:PDEeps} and Lemma \ref{lem:super_refined}, the conclusion follows.
\end{proof}
Even this special case (Theorem \ref{thm:rate_doubling0}) is new in the literate. As an immediate consequence, we obtain the rate of convergence on any compact subset that is disjoint from the support of $f$.

\begin{cor} Assume $f$ is Lipschitz with compact support and $K$ is a connected compact subset of $\Omega$ that is disjoint from $\mathrm{supp}(f)$. Then there exists a constant $C = C(K)$ independent of $\varepsilon\in (0,1)$ such that 
\begin{equation*}
     \Vert u^\varepsilon - u\Vert_{L^\infty(K)} \leq C\varepsilon^{\alpha+1}.
\end{equation*}
\end{cor}
\begin{proof} We can choose an open, bounded and connected set $U$ such that $\partial U$ is $\rmC^2$ and 
\begin{equation*}
    K\subset\subset U \subset\subset \Omega.    
\end{equation*}
Let $w^\varepsilon$ be the solution to \eqref{eq:PDEeps} with $\Omega$ replaced by $U$. Then by Theorem \ref{thm:rate_doubling0}, we have
\begin{equation*}
    0\leq w^\varepsilon(x)\leq C\left(\varepsilon^{\alpha+1} + \varepsilon^{\alpha+2}\right), \qquad x\in K
\end{equation*}
where $C$ depends on $\mathrm{dist}(K, \partial U)$ and $U$. Recall that $u = 0$ outside the support of $f$. By the comparison principle in $U$, we see that $u^\varepsilon \leq w^\varepsilon$ and thus the conclusion follows.
\end{proof}

For the general case, we have the following theorem.

\begin{thm}[Compactly supported data]\label{thm:rate_doubling1} Assume that $f$ is Lipschitz and is compactly supported in $\Omega_{\kappa}$. Let $u^\varepsilon$ be the unique solution to \eqref{eq:PDEeps} and $u$ be the unique solution to \eqref{eq:PDE0}. Then there exists a constant $C$ independent of $\varepsilon\in(0,1)$ and $\kappa$ such that 
\begin{align}
   &-C \sqrt{\varepsilon}\leq  u^\varepsilon(x) - u(x) \leq C\left(\sqrt{\varepsilon}+\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+2}\right) + \frac{\nu C_\alpha  \varepsilon^{\alpha+1}}{d(x)^\alpha}, \quad\; \qquad p<2,\label{eq:cp1}\\
   &-C\sqrt{\varepsilon}\leq u^\varepsilon(x) - u(x) \leq C \left(\sqrt{\varepsilon}+ \left(\frac{\varepsilon}{\kappa}\right)^2\right) +  \nu \varepsilon \log\left(\frac{1}{d(x)}\right), \qquad p=2, \label{eq:cp2}
\end{align}
for any $x\in \Omega$. As a consequence $|u^\varepsilon(x)-u(x)|\leq C\sqrt{\varepsilon}$ for all $x\in \Omega$ with $d(x)\geq \varepsilon$.
%Furthermore, on any compact subset $K\subset\Omega\backslash \mathrm{supp}(f)$, $|u^\varepsilon - u| = \mathcal{O}(\varepsilon)$.
\end{thm}

We state the following Lemma as a preparation.

\begin{lem} Let $0<\kappa < \delta_0$ and $U_\kappa = \big\{x\in \Omega: 0<\mathrm{dist}(x,\partial\Omega) < \kappa\big\} = \Omega\backslash \overline{\Omega}_\kappa$. There holds
\begin{equation*}
    \mathrm{dist}(x,\partial\Omega_\kappa) = \kappa - \mathrm{dist}(x,\partial\Omega) \qquad\text{for all}\;x\in U_\kappa.
\end{equation*}
As a consequence, $x\mapsto \mathrm{dist}(x,\partial U_\kappa) = \min\big\{\mathrm{dist}(x,\partial \Omega_k),\mathrm{dist}(x,\partial \Omega)\big\}$ is twice continuously differentiable for $x\in \Omega\backslash \overline{\Omega}_{\frac{\kappa}{2}}$. Hence, we can choose 
\begin{equation}\label{e:delta_kappa}
    \delta_{0,U_\kappa} \geq \frac{\kappa}{4}
\end{equation}
where $\delta_{0,\Omega}$ is defined as in \eqref{def:delta_0}.
\end{lem}

\begin{proof} By the definition of $\delta_0 = \delta_{0,\Omega}$, we have $d(x) = \mathrm{dist}(x,\partial\Omega)$ is twice continuously differentiable in the region $U_{\delta_0} = \Omega\backslash \overline{\Omega}_{\delta_0}$. The proof follows from \cite[p. 355]{gilbarg_elliptic_2001}. 
\end{proof}

%\textcolor{red}{"Furthermore, on any compact subset $K\subset\Omega\backslash \mathrm{supp}(f)$, $|u^\varepsilon - u| = \mathcal{O}(\varepsilon)$."--It sounds like the constant inside $O(\varepsilon)$ is uniform in $K$, and do we require anything about the support of $f$?...(I will send a picture.)}

\begin{proof}[Proof of Theorem \ref{thm:rate_doubling1}] Without loss of generality, assume that $f$ is supported in $\Omega_\kappa$ where $0<\kappa < \delta_{0}$.
Let $g_\kappa = u^\varepsilon$ on $\partial\Omega_{\kappa}$. Then the solution $u^\varepsilon$ of \eqref{eq:PDEeps} also solves
\begin{equation*}
    \left\{
  \begin{aligned}
  u^\varepsilon(x) + |Du^\varepsilon(x)|^p-\varepsilon \Delta u^\varepsilon(x) &=0 \;\qquad \text{in } U_\kappa ,\\
  u^\varepsilon(x) &= +\infty \quad \text{on } \partial \Omega,\\
  u^\varepsilon(x) &= g_\kappa \;\;\quad \text{on } \partial \Omega_{\kappa},
    \end{aligned}
\right.
\end{equation*}
in the annulus $U_\kappa= \Omega \setminus \overline{\Omega}_{\kappa} = \{x\in \Omega: 0< d(x) < \kappa\}$. Let $\tilde{u}^\varepsilon\in \mathrm{C}^2(U_\kappa)$ be the solution to the following problem
\begin{equation*}
    \left\{
        \begin{aligned}
            \tilde{u}^\varepsilon(x) + |D\tilde{u}^\varepsilon(x)|^p-\varepsilon \Delta \tilde{u}^\varepsilon(x) &=0 \;\qquad \text{in } U_\kappa ,\\
            \tilde{u}^\varepsilon(x) &= +\infty \quad \text{on } \partial U = \partial \Omega\cup \partial \Omega_{\kappa},
        \end{aligned}
    \right.
\end{equation*}
whose existence is guaranteed by Theorem \ref{thm:wellposed1<p<2}.
Here the boundary condition is understood in the sense that $\tilde{u}^\varepsilon(x)\to \infty$ as $d_\kappa(x)\to 0$, where $d_\kappa(\cdot)$ is the distance function from the boundary of the annulus $U_\kappa$, i.e.,
\begin{equation*}
    d_\kappa(x) = \min \big\lbrace \mathrm{dist}(x,\partial \Omega_\kappa),\mathrm{dist}(x,\partial\Omega)  \big\rbrace \leq d(x) \qquad\text{for}\;x\in U_\kappa.
\end{equation*}
Since $f = 0$ in $\overline{U}_\kappa$, by Lemma \ref{lem:f=0} $u=0$ in $\overline{U}_\kappa$. Hence, $u$ is also the unique state-constraint solution to
\begin{equation*}
    \left\{
        \begin{aligned}
            u(x)+ |Du(x)|^p &=0 \quad \text{in } U_\kappa ,\\
            u(x)+ |Du(x)|^p &\geq 0 \quad \text{on } \partial U_\kappa = \partial \Omega \cup\partial \Omega_{\kappa}.
        \end{aligned}
    \right.
\end{equation*}
The vanishing viscosity of $\tilde{u}^\varepsilon \to 0$ in $U_\kappa$ can be quantified by Theorem \ref{thm:rate_doubling0}, which gives us
\begin{equation*}
\begin{split}
    &0\leq \tilde{u}^\varepsilon(x) \leq \frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d_\kappa(x)^\alpha}+C_3\left(\frac{\varepsilon}{\delta_{0,U_\kappa}}\right)^{\alpha+2}\qquad\quad\;\text{for}\;p<2,\\
    &0\leq \tilde{u}^\varepsilon(x) \leq \nu \varepsilon \log\left(\frac{1}{d_\kappa(x)}\right)+C\left(\frac{\varepsilon}{\delta_{0,U_\kappa}}\right)^{2}\qquad\text{for}\;p=2,
\end{split}    
\end{equation*}
for $x\in U_\kappa$. From \eqref{e:delta_kappa} and the comparison principle in $U_\kappa$, we have
\begin{align}
    &0\leq u^\varepsilon(x) \leq \tilde{u}^\varepsilon(x)  \leq \frac{\nu C_\alpha\varepsilon^{\alpha+1}}{d_\kappa(x)^{\alpha}} + C_3\left(\frac{4\varepsilon}{\kappa}\right)^{\alpha+2} \qquad\quad\;\text{for}\;p<2, \qquad \label{annulus2}\\
    &0\leq u^\varepsilon(x)\leq \tilde{u}^\varepsilon(x) \leq \nu \varepsilon \log\left(\frac{1}{d_\kappa(x)}\right)+C\left(\frac{4\varepsilon}{\kappa}\right)^{2}\qquad\text{for}\;p=2,\label{annulus2p=2}
\end{align}
for $x\in U_\kappa$. We proceed with the doubling variable method. For $p<2$, consider the auxiliary functional 
\begin{equation*}
    \Phi(x,y)= u^\varepsilon(x) - u(y) -\frac{C_0|x-y|^2}{\sigma} - \frac{\nu C_\alpha \varepsilon^{\alpha +1}}{d(x)^\alpha}, \qquad (x,y)\in \overline{\Omega}\times \overline{\Omega},
\end{equation*}
where $C_0$ is the Lipschitz constant of $u$ from \eqref{e:C0}, $\sigma\in (0,1)$. The fact that $\displaystyle d(x)^\alpha u^\varepsilon(x) \to C_\alpha \varepsilon^{\alpha+1}$ as $d(x) \to 0$ implies
\begin{equation*}
    \max_{(x,y) \in \overline{\Omega} \times \overline{\Omega}} \Phi(x,y) = \Phi(x_\sigma, y_\sigma) \qquad\text{for some}\;(x_\sigma,y_\sigma) \in \Omega \times \overline{\Omega}.
\end{equation*}
From $\Phi(x_\sigma, y_\sigma) \geq \Phi(x_\sigma, x_\sigma)$, we can deduce that
\begin{equation}\label{e:sigma}
    \left| x_\sigma - y_\sigma \right| \leq \sigma.  
\end{equation}
If $ d(x_\sigma) \geq \frac{1}{2}\kappa$, since $x\mapsto \Phi(x,y_\sigma)$ has a maximum over $\Omega$ at $x=x_\sigma$, the subsolution test for $u^\varepsilon(x)$ gives us
\begin{align}\label{e:subsln}
    &u^\varepsilon(x_\sigma) + \left|\frac{2C_0(x_\sigma - y_\sigma)}{\sigma} -  \frac{\nu C_\alpha\alpha \varepsilon^{\alpha+1} D d(x_\sigma)}{d(x_\sigma)^{\alpha+1}}\right|^p - f(x_\sigma)\nonumber\\
    &\qquad -\varepsilon\left(\frac{2nC_0}{\sigma}+ \frac{\nu C_\alpha\alpha(\alpha+1) \varepsilon^{\alpha+1}|D d(x_\sigma)|^2}{d(x_\sigma)^{\alpha+2}} - \frac{\nu C_\alpha\alpha \varepsilon^{\alpha+1}\Delta d(x_\sigma)}{d(x_\sigma)^{\alpha+1}}\right) \leq 0.
\end{align}
Since $y\mapsto \Phi(x_\sigma,y)$ has a maximum over $\overline{\Omega}$ at $y = y_\sigma$, the supersolution test for $u(y)$ gives us
\begin{align}\label{e:supersln}
    u(y_\sigma) + \left|\frac{2C_0(x_\sigma - y_\sigma)}{\sigma}\right|^p - f(y_\sigma) \geq 0.
\end{align}
For simplicity, define
\begin{equation*}
    \xi_\sigma := \frac{2C_0(x_\sigma - y_\sigma)}{\sigma} \qquad\text{and}\qquad \zeta_\sigma :=- \frac{\nu C_\alpha\alpha \varepsilon^{\alpha+1} D d(x_\sigma)}{d(x_\sigma)^{\alpha+1}}.
\end{equation*}
From \eqref{e:sigma} and $d(x_\sigma) \geq \frac{1}{2}\kappa$, 
\begin{equation*}
    |\xi_\sigma|\leq 2C_0 \qquad\text{and}\qquad |\zeta_\sigma| \leq \nu K_1 C_\alpha\alpha \left(\frac{\varepsilon}{d(x_\sigma)}\right)^{\alpha+1} \leq \nu K_1 C_\alpha \alpha  \left(\frac{2\varepsilon}{\kappa}\right)^{\alpha+1}.
\end{equation*}
Using the inequality \eqref{e:ineq} with $\gamma = p > 1$, we deduce that
\begin{align}\label{e:estia}
    \Big||\xi_\sigma +\zeta_\sigma|^p - |\xi_\sigma|^p \Big| &\leq p\Big(|\xi_\sigma|+|\zeta_\sigma|\Big)^{p-1}|\zeta_\sigma|\nonumber\\
    &\leq p\left[2C_0+\nu K_1 C_\alpha\alpha \left(\frac{2\varepsilon}{\kappa}\right)^{\alpha+1}\right]^{p-1}\nu K_1 C_\alpha\alpha \left(\frac{2\varepsilon}{\kappa}\right)^{\alpha+1}.
\end{align}
Combine \eqref{e:estia} together with \eqref{e:subsln}, \eqref{e:supersln} and $|f(x_\sigma) - f(y_\sigma)|\leq C|x_\sigma - y_\sigma|\leq C\sigma$ to obtain
\begin{align*}
     u^\varepsilon(x_\sigma) - u(y_\sigma) \leq & p\left(2C_0+\nu K_1 C_\alpha \alpha\left( \frac{2\varepsilon}{\kappa}\right)^{\alpha+1}\right)^{p-1}\nu K_1 C_\alpha \alpha \left(\frac{2\varepsilon}{\kappa}\right)^{\alpha+1} + C\sigma\nonumber\\
    &+2nC_0\left(\frac{\varepsilon}{\sigma}\right) + \nu K_1^2C_\alpha \alpha(\alpha+1)\left(\frac{2\varepsilon}{\kappa}\right)^{\alpha+2} + \nu K_2 C_\alpha \alpha  \left(\frac{2\varepsilon}{\kappa}\right)^{\alpha+1}\varepsilon\nonumber\\
    &\leq C\left[\sigma + \frac{\varepsilon}{\sigma} + \left(1+\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1}\right)^{p-1}\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1} + \left(\frac{\varepsilon}{\kappa}\right)^{\alpha+2} \right].
    %&\leq C\sqrt{\varepsilon}
\end{align*}
By the fact that $(1+x)^\gamma \leq 1+x^\gamma$ for $x\in [0,1]$ and $\gamma\in [0,1]$, we know
\begin{equation*}
    \left(1+\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1}\right)^{p-1} \leq 1+ \left(\frac{\varepsilon}{\kappa}\right),
\end{equation*}
as $0<p-1\leq 1$.
Therefore,
\begin{equation*}
     u^\varepsilon(x_\sigma) - u(y_\sigma) \leq C\left[\sigma + \frac{\varepsilon}{\sigma} + \left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1} + \left(\frac{\varepsilon}{\kappa}\right)^{\alpha+2}\right]
\end{equation*}
where $C$ is independent of $\kappa$ and $\varepsilon$. Now choose $\sigma = \sqrt{\varepsilon}$ to get (with $\kappa$ fixed)
\begin{equation}\label{e:final1}
    \Phi(x_\sigma,y_\sigma) \leq u^\varepsilon(x_\sigma) - u(y_\sigma) \leq C\sqrt{\varepsilon}.
\end{equation}
If $d(x_\sigma) < \frac{1}{2}\kappa$, then $x_\sigma \in U_\kappa$ and furthermore $\mathrm{dist}(x_\sigma,\partial \Omega_\kappa) > \frac{1}{2}\kappa$. Indeed, for any $y\in\partial\Omega$ and $z\in \partial\Omega_k$, we have $|x_\sigma - z| + |x_\sigma - y| \geq |y-z|$. Taking the infimum over all $y\in \partial\Omega$, we deduce that
\begin{equation*}
    |x_\sigma - z| + d(x_\sigma) \geq \inf_{y\in \partial \Omega}|y-z| = d(z) = \kappa
\end{equation*}
since $z\in \partial\Omega_k = \{x \in \Omega: d(x) = \kappa\}$. Thus, $|x_\sigma - z| \geq \kappa - d(x_\sigma) > \frac{1}{2}\kappa$ for all $z\in \partial\Omega_k$, which implies that $\mathrm{dist}(x_\sigma,\partial\Omega_k)>\frac{1}{2}\kappa$ and hence $d_\kappa(x_\sigma) = d(x_\sigma)$. By \eqref{annulus2} and the fact that $u \geq 0$, we have
\begin{align}
    \Phi(x_\sigma, y_\sigma)\leq u^\varepsilon(x_\sigma) - \frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x_\sigma)^\alpha} \leq C_3\left(\frac{4\varepsilon}{\kappa}\right)^{\alpha+2} \label{e:final2}.
\end{align}
%\textcolor{red}{(I just read this one more time and feel confused here as the RHS should be $\frac{\nu C_\alpha \epsilon^{\alpha + 1}}{d(x_\sigma)}+C_3(\frac{4\epsilon}{\kappa})^{\alpha+2}$. But then the following arguement doesn't make sense. I think the inequality should be something like 
%\begin{equation*}
%\begin{aligned}
%    \Phi(x_\sigma, y_\sigma) &\leq u^\varepsilon (x_\sigma)- \frac{\theta  C_\alpha  \varepsilon^{\alpha+1}}{d(x_\sigma)^\alpha}\nonumber\\
 %   &= u^\varepsilon (x_\sigma)-\frac{\nu C_\alpha  \varepsilon^{\alpha+1}}{d_\kappa(x_\sigma)^\alpha}
 %   \leq C_3\left(\frac{4\varepsilon}{\kappa}\right)^{\alpha+2}
%\end{aligned}   
%\end{equation*} 
%where we choose $\theta = \nu$, similar to the proof in draft01. The conclusion isn't much more different.)}
\noindent
Since $\Phi(x,x) \leq \Phi(x_\sigma,y_\sigma)$ for all $x\in \Omega$, we obtain from \eqref{e:final1} and \eqref{e:final2} that
\begin{equation*}
    u^\varepsilon(x)-u(x)-\frac{\nu C_\alpha  \varepsilon^{\alpha+1}}{d(x)^\alpha} \leq C\sqrt{\varepsilon} +C_3\left(\frac{4\varepsilon}{\kappa}\right)^{\alpha+2}
\end{equation*}
and thus \eqref{eq:cp1} follows.

\noindent For $p=2$, we consider instead the functional 
\begin{equation*}
    \Phi(x,y) = u^\varepsilon(x) - u(y) - \frac{C_0|x-y|^2}{\sigma} - \nu \varepsilon \mathrm{log}\left(\frac{1}{d(x)}\right), \qquad (x,y)\in \overline{\Omega}\times \overline{\Omega}.
\end{equation*}
Similar to the previous case where $1<p<2$, the maximum of $\Phi$ must occur at $(x_\sigma,y_\sigma)\in \Omega\times\overline{\Omega}$ and $|x_\sigma - y_\sigma|\leq \sigma$. If $d(x_\sigma)\geq \frac{1}{2}\kappa$, by the subsolution test for $u^\varepsilon(x)$, we have
\begin{align}\label{e:subslnp=2}
    u^\varepsilon(x_\sigma)&+ \left|\frac{2C_0(x_\sigma - y_\sigma)}{\sigma} - \nu \varepsilon \frac{D d(x_\sigma)}{d(x_\sigma)}\right|^2  -f(x_\sigma) \nonumber\\
    &- 2nC_0\left(\frac{\varepsilon}{\sigma}\right) - \nu |D d(x_\sigma)|^2 \left(\frac{\varepsilon}{d(x_\sigma)}\right)^2  + \nu \Delta d(x_\sigma)\left(\frac{\varepsilon^2}{d(x_\sigma)}\right) \leq 0.
\end{align}
By the supersolution test for $u(y)$, we have
\begin{equation}\label{e:superslnp=2}
    u(y_\sigma) + \left|\frac{2C_0(x_\sigma - y_\sigma)}{\sigma}\right|^2 - f(y_\sigma) \geq 0.
\end{equation}
Subtract \eqref{e:superslnp=2} from \eqref{e:subslnp=2} to get
\begin{align*}
    u^\varepsilon(x_\sigma) - u(y_\sigma) &\leq \left(4C_0+ \nu\varepsilon\frac{D d(x_\sigma)}{d(x_\sigma)}\right)\left(\nu \varepsilon \frac{ D d(x_\sigma)}{d(x_\sigma)}\right) \\
    &+ C\sigma + 2nC_0 \left(\frac{\varepsilon}{\sigma}\right)+ \nu |D d(x_\sigma)|^2 \left(\frac{\varepsilon}{d(x_\sigma)}\right)^2 + \nu|\Delta d(x_\sigma)| \frac{\varepsilon^2}{d(x_\sigma)}.
\end{align*}
Using $d(x_\sigma)\geq \frac{1}{2}\kappa$ and bounds on $d(x)$ in \eqref{boundond}, we see that
\begin{align}\label{p=2a}
    \Phi(x_\sigma,y_\sigma)&\leq u^\varepsilon(x_\sigma) - u(y_\sigma)\nonumber \\
    &\leq 4K_1^2\nu(1+\nu)\left(\frac{\varepsilon}{\kappa}\right)^2 + C\sigma + 2nC_0\left( \frac{\varepsilon}{\sigma}\right) + 2\nu(K_2\varepsilon+4C_0K_1)\left(\frac{\varepsilon}{\kappa}\right)\nonumber\\
    &\leq C\left(\sigma+\frac{\varepsilon}{\sigma} + \frac{\varepsilon}{\kappa} + \left(\frac{\varepsilon}{\kappa}\right)^2\right) \leq C\sqrt{\varepsilon}
\end{align}
if we choose $\sigma = \sqrt{\varepsilon}$.

\noindent
If $d(x_\sigma)<\frac{1}{2}\kappa$, then $x_\sigma\in U_\kappa$. Again, we have $d_\kappa(x_\sigma) = d(x_\sigma)$ and from \eqref{annulus2p=2}
\begin{equation}\label{p=2b}
    \Phi(x_\sigma,y_\sigma) \leq u^\varepsilon(x_\sigma) - \nu \varepsilon \log\left(\frac{1}{d(x_\sigma)}\right) \leq  C\left(\frac{4\varepsilon}{\kappa}\right)^2. %\leq C\left(\varepsilon+\varepsilon\,|\log(\kappa)| + \left(\frac{\varepsilon}{\kappa}\right)^2\right).
\end{equation}
Since $\Phi(x,x)\leq \Phi(x_\sigma,y_\sigma)$ for $x\in \Omega$, we obtain from \eqref{p=2a} and \eqref{p=2b} that
\begin{equation*}
    u^\varepsilon(x) - u(x) - \nu\varepsilon\log\left(\frac{1}{d(x)}\right) \leq C\sqrt{\varepsilon} +C\left(\frac{4\varepsilon}{\kappa}\right)^2
\end{equation*}
and thus \eqref{eq:cp2} follows.
%Similar to the case where $1<p<2$, \eqref{annulus2p=2} gives us a better rate $\mathcal{O}(\varepsilon)$ in any compact set $K$ that is disjoint from $\mathrm{supp}(f)$. Thus, the conclusion follows.
\end{proof}

%\begin{rem} From \cite{Lasry1989} we know that, if $p\in \left(\frac{3}{2},2\right]$ then $u^\varepsilon(x) - \frac{C_\alpha\varepsilon^{\alpha+1}}{d(x)^\alpha}$ is uniformly bounded in $\Omega$. Theorem \ref{thm:rate_doubling1} says that, if $f$ is compactly supported in $\Omega$ then 
%\begin{equation*}
%    u^\varepsilon(x) - (\nu+1)\frac{C_\alpha\varepsilon^{\alpha+1}}{d(x)^\alpha}
%\end{equation*}
%is uniformly bounded from above in $\Omega$ for all $p\in (1,2)$. Recall that $\nu = 1+C_2\delta_{0,\Omega}$ from Lemma \ref{lem:super_refined}, we can always choose $\delta_0$ small so that roughly $\nu\approx 2^+$.
%(\textcolor{red}{This sounds really weird.})
%\end{rem}

%\textcolor{orange}{(I think Theorem \ref{thm:rate_doubling1} only says $u^\epsilon-\frac{3C_\alpha \epsilon^{\alpha+1}}{d(x)^\alpha}$ is uniformly bounded above by some constant $C$. It doesn't say $u^\epsilon-\frac{3C_\alpha \epsilon^{\alpha+1}}{d(x)^\alpha}$ is uniformly bounded below. )}

\begin{rem} For general Lipschitz data $f\in \mathrm{C}(\overline{\Omega})$, it is natural to try a cutoff function argument. Let $\chi_{{\kappa}}\in \mathrm{C}_c^\infty(\Omega)$ such that $0\leq \chi_{\kappa}\leq 1$, $\chi_\kappa = 1$ in $\Omega_{2\kappa}$ and $\mathrm{supp}\;\chi_\kappa\subset\Omega_\kappa$. Let $u^\varepsilon_\kappa\in \rmC^2(\Omega)\cap\rmC(\overline{\Omega})$ solve \eqref{eq:PDEeps} with data $f\chi_{\kappa}$. Then $u^\varepsilon_\kappa\to u^\varepsilon$ as $\kappa\to 0$ (since $f\chi_\kappa\to f$ in the weak$^*$ topology of $L^\infty(\Omega)$ and we have the continuity of the solution to \eqref{eq:PDEeps} with respect to data in this topology \cite[Remark II.1]{Lasry1989}). However, it is not clear at the moment how to quantify this rate of convergence, since $f\chi_\kappa$ does not converge to $f$ in the uniform norm, unless $f = 0$ on $\partial\Omega$. 
\end{rem}

\subsection{A rate for zero boundary data}

We prove the rate of convergence for the case $f=0$ on $\partial \Omega$.

\begin{proof}[Proof of Theorem \ref{main_thm1}] Let $L = \Vert D f\Vert_{L^\infty(\Omega)}$ be the Lipschitz constant of $f$. For $\kappa>0$ small such that $0<\kappa<\delta_0$ and $x\in \Omega\backslash \Omega_\kappa$, let $x_0$ be the projection of $x$ onto $\partial\Omega$. We observe that
\begin{equation}\label{e:bound_aa}
    f(x) = f(x) - f(x_0) \leq L|x-x_0| = L\kappa.
\end{equation}
Define
\begin{equation*}
    g_\kappa(x) = 
    \begin{cases}
        0           &\qquad\text{if}\;0\leq d(x) \leq \frac{\kappa}{2},\\
        2L\left(d(x)-\frac{\kappa}{2}\right)       &\qquad\text{if}\;\frac{\kappa}{2}\leq d(x) \leq \kappa.
    \end{cases}
\end{equation*}

%\textcolor{orange}{(A question is why we can assume $f\in C^1(\overline{\Omega})$? If you smooth $f$ up by mollifier, the support of $f$ might go outside of $\Omega$. Another question is I think you want $g_k$ to be $2L(d(x)-\frac{\kappa}{2})$ if $\frac{\kappa}{2} \leq d(x) \leq \kappa $. )}
\noindent
It is clear that for $x\in \partial\Omega_k$, $g_\kappa(x) = L\kappa \geq f(x)$ since \eqref{e:bound_aa}. Therefore, we can define the following continuous function
\begin{equation}\label{def_f_kappa}
    f_\kappa(x) = 
    \begin{cases}
        0             &\qquad\text{if}\;0\leq d(x) \leq \frac{\kappa}{2},\\
        \min \left\lbrace g_\kappa(x), f(x) \right\rbrace &\qquad\text{if}\;\frac{\kappa}{2}\leq d(x) \leq \kappa,\\
        f(x) &\qquad\text{if}\;\kappa \leq d(x).
    \end{cases}
\end{equation}
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.40]{Drafts and notes/fig1.png}
    \caption{Graph of the function $f_\kappa$.}
    \label{fig:f_kappa}
\end{figure}
A graph of $f_\kappa$ is given in Figure \ref{fig:f_kappa}. The continuity at $x\in \partial\Omega_\kappa$ comes from the fact that when $d(x) =\kappa$, we have $g_k(x) = L\kappa \geq f(x)$ by \eqref{e:bound_aa}. It is clear that $f_\kappa$ is Lipschitz with $\Vert f_\kappa\Vert_{L^\infty(\Omega)}\leq L$ as well and $f_\kappa\to f$ uniformly as $\kappa\to 0$. Indeed, we have $0\leq f_\kappa \leq f$ and
\begin{equation*}
    0\leq \max_{x\in \overline{\Omega}} (f(x) - f_\kappa(x)) \leq \max_{x\in \overline{\Omega}\backslash \overline{\Omega}_\kappa} (f(x) - f_\kappa(x)) = \max_{x\in \overline{\Omega}\backslash \overline{\Omega}_\kappa} f(x) \leq L\kappa.
\end{equation*}
Let $u^\varepsilon_\kappa\in \rmC^2(\Omega)\cap\rmC(\overline{\Omega})$ be the solution to \eqref{eq:PDEeps} with data $f\chi_{\kappa}$ and $u_k\in \mathrm{C}(\overline{\Omega})$ be the corresponding solution to \eqref{eq:PDE0} with data $f{\chi_\kappa}$. By the comparison principle (\cite[Corollary II.1]{Lasry1989}), we have
\begin{equation}\label{e:bound_2aa}
    0\leq u^\varepsilon(x) - u^\varepsilon_\kappa(x) \leq L\kappa \qquad\text{for}\;x\in \Omega.
\end{equation}
By the comparison principle for \eqref{eq:PDE0}, we also have
\begin{equation}\label{e:bound_2ab}
    0\leq u(x) - u_\kappa(x) \leq L\kappa \qquad\text{for}\;x\in \Omega.
\end{equation}
If $1<p<2$, by Theorem \ref{thm:rate_doubling1}, there exists a constant $C$ independent of $\kappa$ such that
\begin{equation}\label{e:bound_2ac}
    -C\sqrt{\varepsilon}\leq u^\varepsilon_\kappa(x) - u_\kappa(x)\leq C\left[\sqrt{\varepsilon} + \left(\frac{\varepsilon}{\kappa}\right)^{\alpha+2} + \frac{\varepsilon^{\alpha+1}}{d(x)^\alpha}\right], \qquad x\in \Omega.
\end{equation}
Combining \eqref{e:bound_2aa}, \eqref{e:bound_2ab} and \eqref{e:bound_2ac}, we obtain
\begin{equation*}
\begin{split}
   -C\sqrt{\varepsilon}\leq u^\varepsilon(x) - u(x) &= \Big(u^\varepsilon(x) - u^\varepsilon_\kappa(x)\Big) + \Big(u^\varepsilon_\kappa(x) - u_\kappa(x)\Big) + \Big(u_\kappa(x) - u(x)\Big) \\
    &\leq L\kappa + C\left[\sqrt{\varepsilon}  + \left(\frac{\varepsilon}{\kappa}\right)^{\alpha+2} + \frac{\varepsilon^{\alpha+1}}{d(x)^\alpha}\right], \qquad x\in \Omega.
\end{split}
\end{equation*}
Choose $\kappa = \sqrt{\varepsilon}$ and we deduce that
\begin{equation*}
    -C\sqrt{\varepsilon}\leq u^\varepsilon(x) - u(x) \leq C\sqrt{\varepsilon} + \frac{C\varepsilon^{\alpha+1}}{d(x)^\alpha}
\end{equation*}
for $x\in \Omega$ and the conclusion follows. \\

\noindent If $p=2$, by Theorem \ref{thm:rate_doubling1}, there exists a constant $C$ independent of $\kappa$ such that
\begin{equation}\label{e:bound_2acp=2}
    -C\sqrt{\varepsilon}\leq u^\varepsilon_\kappa(x) - u_\kappa(x) \leq C\left[\sqrt{\varepsilon} + \left(\frac{\varepsilon}{\kappa}\right)^2 + \varepsilon\log\left(\frac{1}{d(x)}\right)\right], \qquad x\in \Omega.
\end{equation}
Combining \eqref{e:bound_2aa}, \eqref{e:bound_2ab} and \eqref{e:bound_2acp=2}, we obtain
\begin{equation*}
\begin{split}
   -C\sqrt{\varepsilon}\leq u^\varepsilon(x) - u(x) &= \Big(u^\varepsilon(x) - u^\varepsilon_\kappa(x)\Big) + \Big(u^\varepsilon_\kappa(x) - u_\kappa(x)\Big) + \Big(u_\kappa(x) - u(x)\Big) \\
    &\leq L\kappa + C\left[\sqrt{\varepsilon} +  \left(\frac{\varepsilon}{\kappa}\right)^{2} + \varepsilon\log\left(\frac{1}{d(x)}\right) \right], \qquad x\in \Omega.
\end{split}
\end{equation*}
Choose $\kappa = \varepsilon$ and we deduce that
\begin{equation*}
    -C\sqrt{\varepsilon}\leq u^\varepsilon(x) - u(x) \leq C\sqrt{\varepsilon} +\varepsilon\log\left(\frac{1}{d(x)}\right)
\end{equation*}
for $x\in \Omega$ and the conclusion follows. 
\end{proof}

\section{Improved one-sided rate of convergence for compactly supported data}
We still assume $f = 0$ on $\partial\Omega$ and $f\geq 0$ such that $f\in \rmC^2(\overline{\Omega})$. From Lemma \ref{lem:f=0}, we have $u = 0$ on $\partial\Omega$. Extent $u$ to $\R^n$ by setting 
\begin{equation*}
    \tilde{u}(x) = \begin{cases}
      u(x) &\qquad\text{if}\;x\in \overline{\Omega},\\
      0 &\qquad\text{if}\;x\notin \overline{\Omega}.
    \end{cases}
\end{equation*}
By abuse of notion, we still denote the extension of $f$ on $\R^n$ by itself.

\begin{lem} If $f = 0$ on $\partial\Omega$ then the extension $\tilde{u}$ is a viscosity solution to
\begin{equation*}
    \tilde{u}(x)+|D\tilde{u}(x)|^p - f(x) = 0 \qquad\text{in}\;\R^n.
\end{equation*}
As a consequence (Theorem \ref{convex}) $-D^2\tilde{u} \succ -c\;\mathbb{I}_n$ in $\R^n$ if $f\in \mathrm{C}_c^2(\R^n)$ and 
\begin{equation}\label{def_c}
    c = \max \big\lbrace  D_{\xi\xi}f(x): |\xi|=1, x\in \overline{\Omega} \big\rbrace\geq 0.
\end{equation}
\end{lem}
\begin{proof} We only need to check that the viscosity solution test hold for $\tilde{u}$ as $x\in \partial\Omega$. The supersolution test holds trivially since $f = \tilde{u} = 0$ on $\partial\Omega$. If $\varphi\in \mathrm{C}^1(\R^n)$ such that $\tilde{u} - \varphi$ has a local maximum at $x\in \partial\Omega$, then $\tilde{u}(y) - \tilde{u}(y) \leq - \tilde{u}(x)$ for all $y\in \R^n$. Since $u\geq 0$, we deduce that $\varphi$ has a local minimum at $x\in \partial\Omega$, therefore $\tilde{u}(x)+|D\varphi(x)|^p - f(x) = 0$ and thus the viscosity subsolution test is satisfied.
\end{proof}

The following Lemma is a refined version of the local gradient bound in Theorem \ref{thm:grad_1}. We follows \cite[Theorem 3.1]{Armstrong2015a} where the authors use Bernstein's method inside a doubling variables argument and explicitly keep track of all dependencies. We refer the readers to \cite{barles_weak_1991,capuzzo_dolcetta_holder_2010} and the references therein for related versions of the gradient bound. We believe this result is new in the literature since it is uniform in $\varepsilon$. It also indicates that the boundary layer is the strip of size $\mathcal{O}(\varepsilon)$ from the boundary.

\begin{lem}\label{lem:boundDu^eps} For all $\varepsilon$ small enough, there exists a constant $C$ independent of $\varepsilon$ such that
\begin{equation}\label{eq:es_final}
|Du^\varepsilon(x)| \leq C\left( 1 + \left(\frac{\varepsilon}{d(x)}\right)^{\alpha+1}\right) \qquad\text{for}\;x\in \Omega.
\end{equation}
\end{lem}

\begin{proof}[Proof of Lemma \ref{lem:boundDu^eps}] %Let us consider $0<\varepsilon < \min\{4\delta_0,1\}$. 
Let us fix $x_0 \in \Omega\backslash \Omega_{\delta_0}$ and let $\delta = \frac{1}{4}d(x_0)$. Let 
\begin{equation*}
    v(x) = \frac{1}{\delta}u^\varepsilon(x_0+\delta x), \qquad x\in B(0,2).   
\end{equation*}
We have $Dv(x) = Du^\varepsilon(x_0+\delta x)$ and $\Delta v(x) = \delta \Delta u^\varepsilon(x_0+\delta x)$ for $x\in B(0,2)$. The equation for $v$ is given by
\begin{equation*}
    \delta v(x) + |Dv(x)|^p - \tilde{f}(x) - \frac{\varepsilon}{\delta}\Delta v(x) = 0 \qquad\text{in}\;B(0,2),
\end{equation*}
where $\tilde{f}(x) = f(x_0+\delta x)$ defined on $\overline{B(0,2)}$. We note that $\Vert \tilde{f}\Vert_{L^\infty}\leq \Vert f\Vert_{L^\infty}$ and 
\begin{equation*}
    B(x_0,2\delta)\subset \Omega_{2\delta}\subset\subset\Omega.
\end{equation*}
By Lemma \ref{lem:super_refined}, there is a constant $C$ independent of $\delta,\varepsilon$ such that 
\begin{equation*}
     \delta\Vert v\Vert_{L^\infty\left(B\left(0,\frac{3}{2}\right)\right)} \leq \Vert u^\varepsilon\Vert_{L^\infty(\Omega_{2\delta}))} \leq C\left(1 + \frac{\varepsilon^{\alpha+1}}{\delta^\alpha}\right).
\end{equation*}
Note that $p = \frac{\alpha+2}{\alpha+1}$ and $\alpha+1 = \frac{1}{p-1}$, we apply Theorem 3.1 in \cite{Armstrong2015a} to obtain 
\begin{equation*}
\begin{split}
    \sup_{x\in B(0,1)}|Dv(x)| &\leq C\left[\left(\frac{\varepsilon}{\delta}\right)^{\frac{1}{p-1}} + \left(\Vert f\Vert_{L^\infty}+\delta \Vert v\Vert_{L^\infty\left(B\left(0,\frac{3}{2}\right)\right)}\right)^{\frac{1}{p}}\right] \\
    &\leq C\left[\left(\frac{\varepsilon}{\delta}\right)^{\alpha+1} + \left(1+\frac{\varepsilon^{\alpha+1}}{\delta^\alpha}\right)^{\frac{\alpha+1}{\alpha+2}}\right]\leq C \left(1+\left(\frac{\varepsilon}{\delta}\right)^{\alpha+1}\right).
    %&\leq C\left[\left(\frac{\varepsilon}{\delta}\right)^{\alpha+1} + \left(1+\left(\frac{\alpha+1}{\alpha+2}\right)\frac{\varepsilon^{\alpha+1}}{\delta^\alpha}\right)\right] \leq C\left[1+2\left(\frac{\varepsilon}{\delta}\right)^{\alpha+1} \right]
\end{split}
\end{equation*}
%where we use Bernoulli's inequality and the fact that $\delta \leq 1$. 
Replace $\delta = \frac{1}{4}d(x_0)$ we obtain
\begin{equation*}
    |Du^\varepsilon(x_0)| = |Dv(0)| \leq C\left(1+ \left(\frac{\varepsilon}{d(x_0)}\right)^{\alpha+1}\right).
\end{equation*}
In other words, we obtain \eqref{eq:es_final} for all $x\in \Omega\backslash\Omega_{\delta_0}$. From Theorem \ref{thm:grad_1} there exists a constant $C$ independent of $\varepsilon$ such that $|Du^\varepsilon(x)|\leq C$ for all $x\in \Omega_{\delta_0}$ and the proof is complete.
\end{proof}

\begin{proof}[Proof of Theorem \ref{thm:rate_doubling2}] For $1<p<2$, we proceed as in the proof of Theorem \ref{thm:rate_doubling1} to obtain 
\begin{align}
    &0\leq u^\varepsilon(x) \leq \tilde{u}^\varepsilon(x)  \leq \frac{\nu C_\alpha\varepsilon^{\alpha+1}}{d_\kappa(x)^{\alpha}} + C_3\left(\frac{4\varepsilon}{\kappa}\right)^{\alpha+2} \label{annulus2a}
    %&0\leq u^\varepsilon(x)\leq \tilde{u}^\varepsilon(x) \leq \nu \varepsilon \log\left(\frac{1}{d_\kappa(x)}\right)+C_\nu\left(\frac{4\varepsilon}{\kappa}\right)^{2}\qquad\text{for}\;p=2,\label{annulus2p=2a}
\end{align}
for $x\in U_\kappa$. Let
\begin{equation*}
    \psi^\varepsilon(x) := u^\varepsilon(x) - \frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x)^\alpha}, \qquad x\in \Omega
\end{equation*}
where $\nu > 1$ is chosen as in Lemma \ref{lem:super_refined}. It is clear that $u-\psi^\varepsilon$ has a local minimum at some point $x_0\in \Omega$ since $\psi^\varepsilon(x)\to -\infty$ as $x\to \partial\Omega$. The normal derivative test gives us 
\begin{equation*}
    D\psi^\varepsilon(x_0) = Du^\varepsilon(x_0) + \nu C_\alpha\alpha \left(\frac{\varepsilon}{d(x_0)}\right)^{\alpha+1} D d(x_0) \in D^-u(x_0).
\end{equation*}
We note that $x_0$ depends on $\nu$. 
As $u$ is Lipschitz with constant $C_0$, we deduce that
\begin{equation*}
    |D\psi^\varepsilon(x_0)| = \left|Du^\varepsilon(x_0) + \nu C_\alpha\alpha \left(\frac{\varepsilon}{d(x_0)}\right)^{\alpha+1} D d(x_0) \right| \leq C_0.
\end{equation*}
There are two cases to consider:
\begin{itemize}
    \item If $\displaystyle d(x_0)< \frac{1}{2}\kappa$, then as in the proof of Theorem \ref{thm:rate_doubling1}, $x_0\in U_\kappa$ and $d_\kappa(x_0) = d(x_0)$. By the definition of $x_0$, for any $x\in \Omega$, there holds
    \begin{align*}
        u(x) - \left(u^\varepsilon(x) - \frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x)^\alpha}\right) \geq u(x_0) - \left(u^\varepsilon(x_0) - \frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x_0)^\alpha}\right).
    \end{align*}
    Therefore,
    \begin{equation*}
    \begin{split}
        u^\varepsilon(x) - u(x) - \frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x)^\alpha} &\leq  \left(u^\varepsilon(x_0) - \frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x_0)^\alpha}\right) - u(x_0) \leq C_3 \left(\frac{4\varepsilon}{\kappa}\right)^{\alpha+2}
    \end{split}
    \end{equation*}
    thanks to \eqref{annulus2a}. Thus, in this case
    \begin{equation*}
        u^\varepsilon(x) - u(x) \leq \frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x)^\alpha} + C_3 \left(\frac{4\varepsilon}{\kappa}\right)^{\alpha+2}, \qquad x\in \Omega.
    \end{equation*}
    \item If $\displaystyle d(x_0) \geq \frac{1}{2}\kappa$, from the fact that $u$ is semiconcave in $\Omega$, 
\begin{equation*}
    D^2\psi^\varepsilon(x_0) \prec c\;\mathbb{I}_n
\end{equation*}
which implies that
\begin{equation*}
    \Delta \psi^\varepsilon(x_0) \leq nc.
\end{equation*}
In other words, we have
\begin{equation*}
    \varepsilon\Delta u^\varepsilon (x_0) - \frac{\nu C_\alpha \alpha(\alpha+1)\varepsilon^{\alpha+2}}{d(x_0)^{\alpha+2}}| Dd(x_0)|^2 + \frac{\nu C_\alpha \alpha \varepsilon^{\alpha+2}}{d(x_0)^{\alpha+1}}\Delta d(x_0) \leq nc\varepsilon.
\end{equation*}
Since $d(x_0)\geq \frac{1}{2}\kappa$, we can further deduce that
\begin{equation}\label{delta_psi}
    \varepsilon\Delta u^\varepsilon(x_0) \leq nc\varepsilon + \frac{C\varepsilon^{\alpha+2}}{d(x_0)^{\alpha+2}} \leq nc\varepsilon +  C\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+2}
\end{equation}
where $C$ is independent of $\varepsilon$. Since $\psi^\varepsilon\in \rmC^2(\Omega)$, the viscosity supersolution test for $u$ gives us
\begin{equation}\label{convex1}
    u(x_0) + \left|Du^\varepsilon(x_0) + \frac{\nu C_\alpha \alpha \varepsilon^{\alpha+1}}{d(x_0)^{\alpha+1}}D d(x_0)\right|^p - f(x_0) \geq 0.
\end{equation}
On the other hand, $u^\varepsilon$ solves
\begin{equation}\label{convex2}
    u^\varepsilon(x_0) + |Du^\varepsilon(x_0)|^p - f(x_0) - \varepsilon \Delta u^\varepsilon(x_0) = 0.
\end{equation}
Combine \eqref{convex1} and \eqref{convex2} to obtain that
\begin{align}\label{convex3}
    u^\varepsilon(x_0) - u(x_0) &\leq \left|Du^\varepsilon(x_0) + \frac{\nu C_\alpha \alpha \varepsilon^{\alpha+1}}{d(x_0)^{\alpha+1}}D d(x_0)\right|^p - |Du^\varepsilon(x_0)|^p + \varepsilon\Delta u^\varepsilon(x_0).
\end{align}
Now we estimate the gradient terms on the right hand side of \eqref{convex3}, i.e., 
\begin{align}\label{convex4}
    &\left|Du^\varepsilon(x_0) + \frac{\nu C_\alpha \alpha \varepsilon^{\alpha+1}}{d(x_0)^{\alpha+1}}D d(x_0)\right|^p - |Du^\varepsilon(x_0)|^p\nonumber \\
    &\qquad \qquad \leq p\left(|Du^\varepsilon(x_0)| + \frac{\nu C_\alpha \alpha \varepsilon^{\alpha+1}}{d(x_0)^{\alpha+1}}\left|D d(x_0)\right|\right)^{p-1} \frac{\nu C_\alpha \alpha \varepsilon^{\alpha+1}}{d(x_0)^{\alpha+1}}|D d(x_0)|\nonumber\\
    &\qquad \qquad \leq p\left(C_\frac{\kappa}{2} + C\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1}\right)^{p-1} C\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1}\nonumber\\
    &\qquad\qquad\leq CC_\frac{\kappa}{2}\left(1+\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1}\right)^{p-1} \left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1} \leq CC_\frac{\kappa}{2} \left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1}\left(1+ \left(\frac{\varepsilon}{\kappa}\right)\right)
\end{align}
where, as is given in the a priori estimate (Theorem \ref{thm:grad_1}),
\begin{equation}\label{est_Ckappa}
    |Du^\varepsilon(x)|\leq C_{\frac{\kappa}{2}}, \qquad \;x\in \Omega_{\frac{\kappa}{2}},
\end{equation}
and $C$ is a constant that depends only on $\nu$, $\alpha$, and bounds on $d,Dd$ and $\Delta d$. Using \eqref{delta_psi} and \eqref{convex4} in \eqref{convex3}, we get
\begin{equation*}
    u^\varepsilon(x_0) - u(x_0) \leq nc\varepsilon + CC_\frac{\kappa}{2} \left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1}\left(1+ \left(\frac{\varepsilon}{\kappa}\right)\right) + C\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+2}.
\end{equation*}
%\textcolor{orange}{(I feel you missed some terms when estimating $\epsilon \Delta u^\epsilon$, because $\Delta \psi \neq \Delta u^\epsilon$. But it won't affect the result.)}

Therefore, in this case 
\begin{equation*}
    u^\varepsilon(x) - u(x) \leq \frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x)^\alpha} + CC_\frac{\kappa}{2} \left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1}\left(1+ \left(\frac{\varepsilon}{\kappa}\right)\right) + nc\varepsilon, \qquad x\in \Omega.
\end{equation*}
\end{itemize}
%From the two cases, the conclusion follows.

For $p = 2$, the argument is similar. We take $\psi^\varepsilon(x):=u^\varepsilon(x)-\nu \varepsilon \log \left(\frac{1}{d(x)}\right)$ instead and still $u-\psi^\varepsilon$ attains a local minimum at some point $x_0 \in \Omega$. Carrying out the similar computations as in the case of $1 < p < 2$, we have:
\begin{itemize}
    \item If $\displaystyle d(x_0) < \frac{1}{2} \kappa$, then
    \begin{equation*}
        u^\varepsilon(x)-u(x) \leq \nu \varepsilon \log \left( \frac{1}{d(x)} \right) + C\left( \frac{4\varepsilon}{\kappa}\right)^2, \quad x \in \Omega.
    \end{equation*}
    \item If $\displaystyle d(x_0) \geq \frac{1}{2} \kappa$, then
    \begin{equation*}
        u^\varepsilon(x)-u(x) \leq \nu \varepsilon \log\left(\frac{1}{d(x)}\right) + 
        C C_\frac{\kappa}{2} \left( \frac{\varepsilon}{\kappa} \right) + C \left( \frac{\varepsilon}{\kappa} \right)^2 + nc\varepsilon, \quad x \in \Omega.
    \end{equation*}
\end{itemize}
Finally, applying Lemma \ref{lem:boundDu^eps} to get rid of the constant $C_{\frac{\kappa}{2}}$ we obtain the conclusion.
\end{proof}

\begin{proof}[A different proof of Theorem \ref{main_thm1}] Let $L$ be the Lipschitz constant of $f$. For $0 < \kappa <\delta_0$ we define $f_\kappa$ as in \eqref{def_f_kappa}. We have $\mathrm{supp}(f_\kappa)\subset\Omega_{\frac{\kappa}{2}}$ and $0\leq f - f_\kappa \leq L\kappa$. Let $u^\varepsilon_\kappa\in \rmC^2(\Omega)\cap \rmC(\overline{\Omega})$ be the solution to \eqref{eq:PDEeps} and $u_k$ be the solution to \eqref{eq:PDE0} with $f$ being replaced by $f_k$, respectively. It is clear that
\begin{equation*}
  0\leq u^\varepsilon(x) - u^\varepsilon_\kappa(x)\leq L\kappa \qquad\text{for}\;x\in \Omega
\end{equation*}
and
\begin{equation*}
    0\leq u(x) - u_\kappa(x)\leq L\kappa \qquad\text{for}\;x\in \Omega.
\end{equation*}
By Theorem \ref{thm:rate_doubling2} we have 
\begin{equation*}
\begin{aligned}
    u_\kappa^\varepsilon(x) - u_\kappa(x) &\leq\frac{\nu C_\alpha \varepsilon^{\alpha+1}}{d(x)^\alpha} + C\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1} \left(1+ \left(\frac{\varepsilon}{\kappa}\right) +\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+1} +\left(\frac{\varepsilon}{\kappa}\right)^{\alpha+2}\right) + nc\varepsilon,  &\quad p <2, \\
    u_\kappa^\varepsilon(x) - u_\kappa(x) & \leq \nu \varepsilon \log\left( \frac{1}{d(x)}\right)+C\left(\frac{\varepsilon}{\kappa}\right) \left(1+\frac{\varepsilon}{\kappa}\right)+ nc \varepsilon, &\quad p=2.
\end{aligned}
\end{equation*}
Therefore
\begin{equation*}
    u^\varepsilon(x)-u(x)\leq 2L\kappa + \Big(u^\varepsilon_\kappa(x) - u_\kappa(x)\Big).
\end{equation*} 
Choose $\kappa = \sqrt{\varepsilon}$ we obtain the conclusion.
\end{proof}

\section{Something new}
Assume $f\in \mathrm{C}(\overline{\Omega})\cap W^{1,\infty}(\Omega)$ with $f = 0$ in $\overline{\Omega}_{2\kappa}$ for $0<\kappa \ll \delta_0$. By doing so we can assume that
\begin{equation*}
    \delta_{0,\Omega_{2\kappa}} = \frac{1}{2}\delta_{0,\Omega}.
\end{equation*}
Let $1<p<2$ and $d_\kappa(x) = \mathrm{dist}(x,\partial\Omega_\kappa)$ for $x\in \Omega_\kappa$. By Theorem \ref{thm:rate_doubling0} we obtain that
\begin{equation*}
    0\leq u^\varepsilon(x)\leq C \left(\frac{\varepsilon^{\alpha+1}}{d_\kappa(x)^{\alpha}} + \frac{\varepsilon^{\alpha+2}}{\delta_{0,\Omega_{\kappa}}^{\alpha+2}}\right)
\end{equation*}
for some constant $C$ independent of $\varepsilon\in (0,1)$ and $\kappa$.

%\clearpage

%\section{Rate of convergence via nonlinear adjoint method}

\begin{appendices}
%\section*{Appendix}
\subsection*{Estimates on solutions}
We present here a proof for the gradient bound of solution to \eqref{eq:PDEeps} using the Bernstein method (see also \cite{Lasry1989,lions_quelques_1985}). Another proof using a different technique is given in \cite{Armstrong2015a}.

\begin{proof}[Proof of Theorem \ref{thm:grad_1}] Let $\theta\in (0,1)$ be chosen later, $\varphi\in \mathrm{C}_c^\infty(\Omega)$, $0\leq \varphi\leq 1$, $\mathrm{supp}\;\varphi\subset \Omega$ and $\varphi = 1$ on $\Omega_\delta$ such that there exists $C = C(\delta,\theta)$,
\begin{equation}\label{e:ass_power}
    |\Delta \varphi(x)| \leq C\varphi^\theta \qquad\text{and}\qquad |D \varphi(x)|^2 \leq C\varphi^{1+\theta}
\end{equation}
for all $x\in \Omega$.
Define $w(x) := |Du(x)|^2$ for $x \in \Omega$. The equation for $w$ is given by
\begin{equation*}
    -\varepsilon \Delta w + 2 p|D u|^{p-2}D u \cdot D w + 2  w - 2 D f\cdot D u + 2 \varepsilon |D^2u|^2 = 0 \qquad\text{in}\;\Omega.
\end{equation*}
%\textcolor{orange}{(I think the coefficients here are not correct. I put the corrected ones in orange. Originally, all the coefficients were 1 except for the term 2w.)}

Then an equation for $(\varphi w)$ can be derived as follows.
\begin{align*}
    &-\varepsilon \Delta (\varphi w) + 2 p|D u|^{p-2}D u \cdot D (\varphi w) + 2  (\varphi w) + 2 \varepsilon \varphi|D^2u|^2 + 2\varepsilon \frac{D \varphi}{\varphi}\cdot D (\varphi w) \\
    &\qquad\qquad\qquad\qquad = \varphi(D f\cdot D u) + 2 p|D u|^{p-2}(D u\cdot D \varphi)w -\varepsilon w \Delta \varphi + 2\varepsilon \frac{|D \varphi|^2}{\varphi}w
    \qquad\text{in}\;\mathrm{supp}\;\varphi.
\end{align*}
Assume that $\varphi w$ achieves its maximum over $\overline{\Omega}$ at $x_0\in \Omega$. And we can further assume that $x_0\in \mathrm{supp}\;\varphi$, since otherwise the maximum of $\varphi w$ over $\overline{\Omega}$ is zero. By the classical maximum principle, 
\begin{equation*}
    -\varepsilon \Delta(\varphi w)(x_0)\geq 0 \qquad\text{and}\qquad |D(\varphi w)(x_0)| = 0.
\end{equation*}
Use this in the equation of $\varphi w$ above to obtain
\begin{equation*}
    \varepsilon \varphi|D^2u|^2 \leq  \varphi (Df\cdot Du)+ 2 p|Du|^{p-1} |D\varphi|w + \varepsilon w |\Delta\varphi|  + 2\varepsilon  w\frac{|D\varphi|^2}{\varphi},
\end{equation*}
 where all terms are evaluated at $x_0$. From \eqref{e:ass_power}, we have
\begin{equation}\label{e:est_for_D^2u}
    \varepsilon \varphi|D^2u|^2 \leq  \varphi |Df|w^{\frac{1}{2}}+ 2 Cp w^{\frac{p-1}{2}+1} \varphi^{\frac{1+\theta}{2}} + C\varepsilon w \varphi^{\theta} + 2C\varepsilon  w\varphi^\theta.
\end{equation}
By Cauchy-Schwartz inquality, $n|D^2u|^2\geq (\Delta u)^2$. Thus, if $n\varepsilon < 1$, then
\begin{equation}\label{e:est_for_D^2u_2}
    \varepsilon |D^2u|^2 \geq \frac{(\varepsilon \Delta u)^2}{n\varepsilon} \geq (\varepsilon \Delta u)^2 = \left(  u + |Du|^p - f\right)^2 \geq |Du|^{2p} - 2C|Du|^p \geq \frac{|Du|^{2p}}{2} - 2C
\end{equation}
where $C$ depends on $\max_{\overline{\Omega}}f$ only. Using \eqref{e:est_for_D^2u_2} in \eqref{e:est_for_D^2u}, we obtain that
\begin{equation*}
    \varphi\left(\frac{1}{2}w^p - 2C\right) \leq \varphi |Df|w^{\frac{1}{2}}+ 2 Cp w^{\frac{p-1}{2}+1} \varphi^{\frac{1+\theta}{2}} + 3C\varepsilon w \varphi^{\theta}.
\end{equation*}
Multiply both sides by $\varphi^{p-1}$ to deduce that
\begin{align*}
    (\varphi w)^p \leq 4C\varphi^{p-1} + 2\Vert Df\Vert_{L^\infty}\varphi^p w^{\frac{1}{2}} + 4Cp \varphi^{\frac{2p+\theta - 1}{2}}w^{\frac{p+1}{2}} + 6C\varepsilon \varphi^{p+\theta - 1}w.
\end{align*}
Choose $2p+\theta -1 \geq p+1$, i.e., $p+\theta\geq 2$. This is always possible with the requirement $\theta \in (0,1)$, as $1<p <\infty$. Then we get
\begin{equation}\label{e:est_for_D^2u_3}
    (\varphi w)^p \leq C\left(1+ (\varphi w)^\frac{1}{2} + (\varphi w)^\frac{p+1}{2} +(\varphi w)\right).
\end{equation}
As a polynomial in $z = (\varphi w)(x_0)$, this implies that $(\varphi w)(x_0)\leq C$ where $C$ depends on coefficients of the right hand side of \eqref{e:est_for_D^2u_3}, which gives our desired gradient bound since $\overline{\Omega}_\delta\subset \mathrm{supp}\;\varphi$.
\end{proof}

\subsection*{Wellposedness of solutions}
\begin{proof} [Proof of Theorem \ref{thm:wellposed1<p<2}] If $p\in (1,2)$, we use the ansatz $ u(x) = C_\varepsilon d(x)^{-\alpha}$ to find a solution for \eqref{eq:PDEeps}. Plug the ansatz into \eqref{eq:PDEeps} and compute 
\begin{equation*}
\begin{split}
    |Du (x)|^p &= \frac{(\alpha C_\varepsilon)^p }{d(x)^{p(\alpha+1)}}|D d(x)|^p,\\
    \varepsilon\Delta u(x) &= \frac{\varepsilon C_\varepsilon\alpha(\alpha+1)}{d(x)^{\alpha+2}}|D d(x)|^2 - \frac{\varepsilon C_\varepsilon\alpha}{d(x)^{\alpha+1}}\Delta d(x).
\end{split}
\end{equation*}
Since $|D d(x)| = 1$ for $x$ near $\partial\Omega$, as $x\to \partial \Omega$, the explosive terms of the highest order are
\begin{equation*}
         C_\varepsilon^p \alpha^p d^{-(\alpha+1)p}  -\varepsilon C_\varepsilon \alpha(\alpha+1)d^{-(\alpha+2)}.
\end{equation*}
Set the above to be zero to obtain that
\begin{equation}\label{e:relation}
    \displaystyle\alpha = \frac{2-p}{p-1} \qquad\text{and}\qquad C_\varepsilon = \left(\frac{1}{\alpha}(\alpha+1)^\frac{1}{p-1}\right) \varepsilon^{\frac{1}{p-1}} = \frac{1}{\alpha}(\alpha+1)^{\alpha+1}\varepsilon^{\alpha+1}.
\end{equation}
For $0<\delta < \frac{1}{2}\delta_0$ and $\eta$ small, define
\begin{equation*}
\begin{split}
    \overline{w}_{\eta,\delta}(x) &:= \frac{(C_\alpha+\eta)\varepsilon^{\alpha+1}}{(d(x)-\delta)^\alpha} + M_\eta, \qquad x\in \Omega_\delta,\\
    \underline{w}_{\eta,\delta}(x) &:= \frac{(C_\alpha-\eta)\varepsilon^{\alpha+1}}{(d(x)+\delta)^\alpha} - M_\eta, \qquad x\in \Omega^\delta,
\end{split}
\end{equation*}
where $C_\alpha := \frac{1}{\alpha} (\alpha+1)^{\alpha+1} $, $M_\eta$ to be chosen. We will show that $\overline{w}_{\eta,\delta}$ is a supersolution of \eqref{eq:PDEeps} in $\Omega_\delta$, while $\underline{w}_{\eta,\delta}$ is a subsolution of \eqref{eq:PDEeps} in $\Omega^\delta$. Compute
\begin{align*}
    \mathcal{L}^\varepsilon\left[\overline{w}_{\eta,\delta}\right](x) &= \frac{ (C_\alpha + \eta)\varepsilon^{\alpha+1}}{(d(x)-\delta)^\alpha} + M_\eta + \frac{(C_\alpha+\eta)^p \alpha^p\varepsilon^{\alpha+2}}{(d(x)-\delta)^{\alpha+2}}|D d(x)|^p - f(x) \\
    & \qquad\qquad\qquad\qquad\qquad - \frac{(C_\alpha+\eta)\alpha(\alpha+1)\varepsilon^{\alpha+2}}{(d(x)-\delta)^{\alpha+2}}|D d(x)|^2 + \frac{(C_\alpha+\eta)\alpha \varepsilon^{\alpha+2}}{(d(x)-\delta)^{\alpha+1}}\Delta d(x)\\
    &\geq M_\eta - f(x) + \underbrace{\frac{\nu C_\alpha\alpha(\alpha+1)\varepsilon^{\alpha+2}}{(d(x)-\delta)^{\alpha+2}}\left[\nu^{p-1}|Dd(x)|^{p} - |Dd(x)|^2 + \frac{(d(x)-\delta)\Delta d(x)}{\alpha+1}\right]}_{I}
\end{align*}
%\textcolor{orange}{(In the parenthesis of I, I think the first term should be $\nu^{p-1}|Dd(x)|^p$ instead of $\nu^{p-1}|Dd(x)|^{p-1}$.)}
where we use $(C_\alpha\alpha)^p = C_\alpha\alpha(\alpha+1)$ and $\nu = \frac{C_\alpha+\eta}{C_\alpha} \in (1,2)$ for small $\eta$. Let
\begin{equation*}
    \delta_\eta : = \frac{\alpha+1}{K_2}\left[\nu^{p-1}-1\right]
\end{equation*}
and $\delta_\eta \to 0$ as $\eta \to 0$.
Recall the definitions in \eqref{boundond}. To get  $\mathcal{L}^\varepsilon\left[\overline{w}_{\eta,\delta}\right]\geq 0$, there are two cases to consider, depending on how large $d(x)-\delta$ is.
\begin{itemize}
    \item If $0< d(x)-\delta <\delta_\eta < \delta_0$ for $\eta$ small and fixed, then $|Dd(x)| = 1$, and thus $I\geq 0$. Hence $\mathcal{L}^\varepsilon\left[\overline{w}_{\eta,\delta}\right]\geq 0$ if we choose $M_\eta \geq \max_{\overline{\Omega}} f$.
    %\textcolor{orange}{(Question: Here, $\eta$ is fixed right?)}
    \item If $d(x)-\delta\geq \delta_\eta$, then 
    \begin{equation*}
        I \leq \left(\frac{1}{\delta_\eta}\right)^{\alpha+2}\nu C_\alpha \alpha(\alpha+1)\left[\nu^{p-1}K_1^{p}+K_1^2+K_2K_0\right]\varepsilon^{\alpha+2}.
    \end{equation*}
    %\textcolor{orange}{(The first term in the bracket should be $\nu^{p-1}K_1^p$ instead of $\nu^{p-1}K_1^{p-1}$.)}

    Thus, we can choose $M_\eta = \max_{\overline{\Omega}} f + C\varepsilon^{\alpha+2}$ for $C$ large enough (depending on $\eta$) so that  $\mathcal{L}^\varepsilon\left[\overline{w}_{\eta,\delta}\right]\geq 0$.
\end{itemize}
Therefore, $\overline{w}_{\eta,\delta}$ is a supersolution in $\Omega_\delta$.

Similarly, we have 
\begin{align*}
    \mathcal{L}_\varepsilon\left[\underline{w}_{\eta,\delta}\right](x) &= \frac{ (C_\alpha - \eta)\varepsilon^{\alpha+1}}{(d(x)+\delta)^\alpha} - M_\eta + \frac{(C_\alpha-\eta)^p \alpha^p\varepsilon^{\alpha+2}}{(d(x)+\delta)^{\alpha+2}}|D d(x)|^p - f(x) \\
    & \qquad\qquad\qquad\qquad\qquad - \frac{(C_\alpha-\eta)\alpha(\alpha+1)\varepsilon^{\alpha+2}}{(d(x)+\delta)^{\alpha+2}}|D d(x)|^2 + \frac{(C_\alpha-\eta)\alpha \varepsilon^{\alpha+2}}{(d(x)+\delta)^{\alpha+1}}\Delta d(x)\\
    &= - M_\eta - f(x) \\
    &+\underbrace{\frac{\nu C_\alpha\alpha(\alpha+1)\varepsilon^{\alpha+2}}{(d(x)+\delta)^{\alpha+2}}\left[\nu^{p-1}|Dd(x)|^{p} - |Dd(x)|^2 + \frac{(d(x)+\delta)\Delta d(x)}{\alpha+1}+\frac{(d(x)+\delta)^2}{\alpha(\alpha+1)\varepsilon}\right]}_{J}
\end{align*}
%\textcolor{orange}{(In the parenthesis of J, the first term should be $\nu^{p-1}|Dd(x)|^p$ instead of $\nu^{p-1}|Dd(x)|^{p-1}$.)}

where $\nu = \frac{C_\alpha-\eta}{C_\alpha}\in (0,1)$ for small $\eta$. Let
\begin{equation*}
    \delta_\eta := \left(1-\nu^{p-1}\right)\left(\frac{\alpha(\alpha+1)\varepsilon}{1+K_2\alpha\varepsilon}\right) 
\end{equation*}
%\textcolor{orange}{( $\delta_\eta := \left(1-\nu^{p-1}\right)\left(\frac{\alpha(\alpha+1)\varepsilon}{1+K_2\alpha\varepsilon}\right)$)}

and $\delta_\eta \to 0$ as $ \eta \to 0$.
To obtain $\mathcal{L}^\varepsilon\left[\underline{w}_{\eta,\delta}\right]\leq 0$, there are two cases to consider depending on how large $d(x)+\delta$ is.
\begin{itemize}
    \item If $0<d(x)+\delta<\delta_\eta < \delta_0$ for $\eta$ small and fixed, then $|Dd(x)| = 1$ , and thus $J\leq 0$. Hence $\mathcal{L}^\varepsilon\left[\underline{w}_{\eta,\delta}\right]\leq 0$ if we choose $M_\eta \geq -\max_{\Omega}f$.
    \item If $d(x)+\delta\geq \delta_\eta$, then
    \begin{equation*}
        |J|\leq \left(\frac{1}{\delta_\eta}\right)^{\alpha+2} \nu C_\alpha\alpha(\alpha+1)\left[\nu^{p-1}K_1^{p}+K_1^2 + \frac{(K_0+1)K_2}{\alpha+1} + \frac{(K_0+1)^2}{\alpha(\alpha+1)\varepsilon}\right]\varepsilon^{\alpha+2}
    \end{equation*}
    %\textcolor{orange}{($|J|\leq \left(\frac{1}{\delta_\eta}\right)^{\alpha+2} \nu C_\alpha\alpha(\alpha+1)\left[K_1^{p}+K_1^2 + \frac{(K_0+\delta)K_2}{\alpha+1} + \frac{(K_0+\delta)^2}{\alpha(\alpha+1)\varepsilon}\right]\varepsilon^{\alpha+2}$)}

     Thus, we can choose $M_\eta = -\max_{\overline{\Omega}} f - C\varepsilon^{\alpha+2}$ for $C$ large enough (depending on $\eta$) so that $\mathcal{L}^\varepsilon\left[\underline{w}_{\eta,\delta}\right]\leq 0$.
\end{itemize}
Therefore, $\underline{w}_{\eta,\delta}$ is a subsolution in $\Omega^\delta$.

For $p=2$, we use the ansatz $u(x) = -C_\varepsilon \log(d(x))$ instead. Similar to the previous case, one can find $u(x) = -\varepsilon\log(d(x))$. For $0<\delta<\frac{1}{2}\delta_0$,  define 
\begin{equation*}
    \begin{split}
        &\overline{w}_{\eta,\delta}(x) = -(1+\eta)\varepsilon\log(d(x)-\delta) + M_\eta, \qquad x\in \Omega_\delta,\\
        &\underline{w}_{\eta,\delta}(x) = -(1-\eta)\varepsilon\log(d(x)+\delta) - M_\eta, \qquad x\in \Omega^\delta,
    \end{split}
\end{equation*}
where $M_\eta$ is to be chosen so that $\overline{w}_{\eta,\delta}(x)$ is a supersolution in $\Omega_\delta$ and $\underline{w}_{\eta,\delta}$ is a subsolution in $\Omega^\delta$. The computations are omitted here, as they are similar to the previous case.
\smallskip

\noindent We divide the rest of the proof into 3 steps. We first construct a minimal solution, then a maximal solution to \eqref{eq:PDEeps}, and finally show that they are equal to conclude the existence and the uniqueness of the solution to \eqref{eq:PDEeps}.
\smallskip

\paragraph{\textbf{Step 1.}} There exists a minimal solution $\underline{u}\in \mathrm{C}^2(\Omega)$ of \eqref{eq:PDEeps} such that $v\geq \underline{u}$ for any other solution $v\in \mathrm{C}^2(\Omega)$ solving \eqref{eq:PDEeps}.

\begin{proof} Let $w_{\eta,\delta}\in \mathrm{C}^2(\Omega)$ solve
    \begin{equation}\label{e:w_def}
    \begin{cases}
        \mathcal{L}^\varepsilon\left[w_{\eta,\delta}\right] = 0 &\qquad\text{in}\;\Omega,\\
        \qquad w_{\eta,\delta} = \underline{w}_{\eta,\delta} &\qquad\text{on}\;\partial\Omega.
    \end{cases}
    \end{equation}
    \begin{itemize}
        \item Fix $\eta>0$. As $\delta\to 0$, the value of $\underline{w}_{\eta,\delta}$ blows up on the boundary. Therefore, by the standard comparison principle for the second-order elliptic equation with Dirichlet boundary, $\delta_1 \leq  \delta_2$ implies $w_{\eta,\delta_1}\geq  w_{\eta,\delta_2}$ on $\overline{\Omega}$. 
        \item For $\delta'>0$, since $\underline{w}_{\eta,\delta'}$ is a subsolution in $\overline{\Omega}$ with finite boundary, 
            \begin{equation}\label{e:cp_delta1}
                0<\delta \leq \delta'\qquad\Longrightarrow\qquad \underline{w}_{\eta,\delta'} \leq w_{\eta_,\delta'}\leq w_{\eta,\delta} \qquad\text{on}\;\overline{\Omega}.
            \end{equation}
        \item Similarly, since $\overline{w}_{\eta,\delta'}$ is a supersolution on $\Omega_{\delta'}$ with infinity value on the boundary $\partial\Omega_{\delta'}$, by comparison principle,
            \begin{equation}\label{e:cp_delta2}
                w_{\eta,\delta} \leq \overline{w}_{\eta, \delta'} \qquad\text{in}\;\Omega_{\delta'} \qquad\Longrightarrow\qquad w_{\eta,\delta} \leq \overline{w}_{\eta,0} \qquad\text{in}\;\Omega.
            \end{equation}
    \end{itemize}
    \noindent From \eqref{e:cp_delta1} and \eqref{e:cp_delta2}, we have
    \begin{equation}\label{e:cp_delta3}
        0<\delta \leq \delta'\qquad\Longrightarrow\qquad \underline{w}_{\eta,\delta'} \leq w_{\eta_,\delta'}\leq w_{\eta,\delta} \leq \overline{w}_{\eta,0} \qquad\text{in}\;\Omega.
    \end{equation}
    Thus, $\{w_{\eta,\delta}\}_{\delta>0}$ is locally bounded in $L^{\infty}_{\mathrm{loc}}(\Omega)$ ($\{w_{\eta,\delta}\}_{\delta>0}$ is uniformly bounded from below). Using the local gradient estimate for $w_{\eta,\delta}$ solving \eqref{e:w_def}, we deduce that $\{w_{\eta,\delta}\}_{\delta>0}$ is locally bounded in $W^{1,\infty}_{\mathrm{loc}}(\Omega)$. Since $w_{\eta,\delta}$ solves \eqref{e:w_def}, we further have that $\{w_{\eta,\delta}\}_{\delta>0}$ is locally bounded in $W^{2,r}_{\mathrm{loc}}(\Omega)$ for all $r<\infty$ by Calderon-Zygmund estimates.

    \noindent Local boundedness of $\{w_{\eta,\delta}\}_{\delta>0}$ in $W^{2,r}_{\mathrm{loc}}(\Omega)$ implies weak$^*$ compactness, that is, there exists a function $u\in W^{2,r}_{\mathrm{loc}}(\Omega)$ such that (via subsequence and monotonicity)
    \begin{equation*}
        w_{\eta,\delta} \rup u \qquad\text{weakly in}\;W^{2,r}_{\mathrm{loc}}(\Omega),\qquad \text{and}\qquad
        w_{\eta,\delta} \to u \qquad\text{strongly in}\;W^{1,r}_{\mathrm{loc}}(\Omega).
    \end{equation*}
    In particular, $w_{\eta,\delta}\to u$ in $\mathrm{C}^1_{\mathrm{loc}}(\Omega)$ thanks to Sobolev compact embedding. Let us rewrite the equation $\mathcal{L}^\varepsilon\left[w_{\eta,\delta}\right] = 0$ as $\varepsilon\Delta w_{\eta,\delta}(x) = F[w_{\eta,\delta}](x)$ for $x \in U\subset\subset \Omega$, where
    \begin{equation*}
        F[w_{\eta,\delta}](x) =    w_{\eta,\delta}(x) + H(x,Dw_{\eta,\delta}(x)).
    \end{equation*}
    Since $w_{\eta,\delta}\to u$ in $\mathrm{C}^1(U)$ as $\delta \to 0$, we have $F[w_{\eta,\delta}](x) \to F(x)$ uniformly in $U$ as $\delta \to 0$,  where 
    \begin{equation*}
        F(x) =   u(x) + H(x,Du(x)).
    \end{equation*}
    In the limit, we obtain that $u\in L^2(U)$ is a weak solution of $\varepsilon\Delta u = F$ in $U$ where $F$ is continuous. Thus, $u\in \mathrm{C}^2(\Omega)$ and by stability, $u$ solves $\mathcal{L}^\varepsilon[u] = 0$ in $\Omega$. From \eqref{e:cp_delta3}, we also have
    \begin{equation*}
        \underline{w}_{\eta,0} \leq u \leq \overline{w}_{\eta,0} \qquad\text{in}\;\Omega.
    \end{equation*}
    Moreover, $u(x)\to \infty$ as $\mathrm{dist}(x,\partial\Omega)\to 0$ with the precise rate like \eqref{rate_p<2} or \eqref{rate_p=2}. Note that by construction, $u$ may depend on $\eta$. But next, we will show that $u$ is independent of $\eta$, by proving $u$ is the unique minimal solution of $\mathcal{L}^\varepsilon[u] = 0$ in $\Omega$ with $u = +\infty$ on $\partial\Omega$.

    \noindent Let $v\in \rmC^{2}(\Omega)$ be a solution to \eqref{eq:PDEeps}. Fix $\delta>0$. Since $v(x)\to \infty$ as $x\to \partial\Omega$ while $w_{\eta,\delta}$ remains bounded on $\partial \Omega$, the comparison principle yields
    \begin{equation*}
        v\geq w_{\eta,\delta} \qquad\text{in} \; \Omega.
    \end{equation*}
    Let $\delta\to 0$ and we deduce that $v\geq u$ in $\Omega$. This concludes that $u$ is the minimal solution in $\mathrm{C}^2(\Omega)(\forall\,r<\infty)$ and thus $u$ is independent of $\eta$. 
\end{proof}

\paragraph{\textbf{Step 2.}} There exists a maximal solution $\overline{u}\in \mathrm{C}^2(\Omega)$ of \eqref{eq:PDEeps} such that $v\leq \overline{u}$ for any other solution $v\in \mathrm{C}^2(\Omega)$ solving \eqref{eq:PDEeps}.

\begin{proof} For each $\delta>0$, let $u_\delta\in \mathrm{C}^2(\Omega_\delta)$ be the minimal solution to $\mathcal{L}^\varepsilon[u_\delta] = 0$ in $\Omega_\delta$ with $u_\delta = +\infty$ on $\partial\Omega_\delta$. By the comparison principle, for every $\eta>0$, there holds
\begin{equation*}
    \underline{w}_{\eta,\delta} \leq u_\delta \leq \overline{w}_{\eta,\delta} \qquad\text{in}\;\Omega_\delta,
\end{equation*}
and
\begin{equation*}
    0<\delta<\delta' \qquad \Longrightarrow\qquad u_\delta \leq u_\delta' \qquad\text{in}\;\Omega_{\delta'}.
\end{equation*}
The monotoniciy, together with the local boundedness of $\{u_\delta\}_{\delta>0}$ in $W^{2,r}_{\mathrm{loc}}(\Omega)$, implies that there exists $u\in W^{2,r}_{\mathrm{loc}}(\Omega)$ for all $r<\infty$ such that $u_\delta\to u$ strongly in $\rmC^1_{\mathrm{loc}}(\Omega)$. Using the equation $\mathcal{L}^\varepsilon [u_\delta] = 0$ in $\Omega_\delta$ and the regularity of Laplace's equation, we can further deduce that $u\in \mathrm{C}^2(\Omega)$ solves \eqref{eq:PDEeps} and 
\begin{equation*}
    \underline{w}_{\eta,0} \leq u\leq \overline{w}_{\eta,0} \qquad\text{in}\;\Omega
\end{equation*}
for all $\eta>0$. As $u_\delta$ is independent of $\eta$ by the previous argument in Step 1, it is clear that $u$ is also independent of $\eta$. Now we show that $u$ is the maximal solution of \eqref{eq:PDEeps}. Let $v\in\rmC^2(\Omega)$ solve \eqref{eq:PDEeps}. Clearly $v\leq u_\delta$ on $\Omega_\delta$. Therefore, as $\delta \to 0$, we have $v\leq u$.
\end{proof}
\noindent In conclusion, we have found a minimal solution $\underline{u}$ and a maximal solution $\overline{u}$ in $\rmC^2(\Omega)$ such that
\begin{equation}\label{e:chain}
    \underline{w}_{\eta,0} \leq \underline{u}\leq \overline{u}\leq \overline{w}_{\eta,0} \qquad\text{in}\;\Omega
\end{equation}
for any $\eta>0$. This extra parameter $\eta$ now enables us to show that $\overline{u} = \underline{u}$ in $\Omega$. The key ingredient here is the convexity in the gradient slot of the operator.
\smallskip
\paragraph{\textbf{Step 3.}} We have $\overline{u}\equiv \underline{u}$ in $\Omega$. Therefore, the solution to \eqref{eq:PDEeps} in $\mathrm{C}^2(\Omega)$ is unique.

\begin{proof} Let $\theta\in (0,1)$. Define $w_\theta = \theta \overline{u} + (1-\theta) \inf_{\Omega} f$. It can be verified that $w_\theta$ is a subsolution to \eqref{eq:PDEeps}. Then one may argue that by the comparison principle,
\begin{equation*}
    w_\theta = \theta \overline{u} + (1-\theta)\inf_{\Omega} f\leq \underline{u} \qquad\text{in}\;\Omega,
\end{equation*}
and conclude that $\overline{u} \leq \underline{u}$ by letting $\theta\to 1$. But we have to be careful here. As they are both explosive solutions, to use the comparison principle, we need to show that $w_\theta \leq \underline{u}$ in a neighborhood of $\partial\Omega$. From \eqref{e:chain}, we see that
\begin{align*}
    &1\leq \frac{\overline{u}(x)}{\underline{u}(x)} \leq \frac{\overline{w}_{\eta,0}(x)}{\underline{w}_{\eta,0}(x)} = \frac{(C_\alpha+\eta)+ M_\eta d(x)^\alpha}{(C_\alpha-\eta)- M_\eta d(x)^\alpha},& 1<p<2,\\
    &1 \leq \frac{\overline{u}(x)}{\underline{u}(x)} \leq \frac{\overline{w}_{\eta,0}(x)}{\underline{w}_{\eta,0}(x)} = \frac{-(1+\eta)\log(d(x)) + M_\eta}{-(1-\eta)\log(d(x))-M_\eta}, & p=2,
\end{align*}
for $x\in \Omega$. Hence
\begin{align*}
     &1\leq  \lim_{d(x)\to 0} \left(\frac{\overline{u}(x)}{\underline{u}(x)}\right) \leq \frac{C_\alpha+\eta}{C_\alpha-\eta}, & 1< p < 2,\\
     &1\leq  \lim_{d(x)\to 0} \left(\frac{\overline{u}(x)}{\underline{u}(x)}\right) \leq \frac{-(1+\eta)}{-(1-\eta)}, & p = 2.
\end{align*}
Since $\eta>0$ is chosen arbitrary, we obtain
\begin{equation*}
     \lim_{d(x)\to 0} \left(\frac{\overline{u}(x)}{\underline{u}(x)}\right) = 1.
\end{equation*}
 This means for any $\varsigma\in(0,1)$, there exists $\delta(\varsigma)>0$ small such that 
\begin{equation*}
\frac{\overline{u}(x)}{\underline{u}(x)}\leq (1+\varsigma)     \qquad \text{, i.e.,}\qquad \left(\frac{1}{1+\varsigma}\right)\overline{u}(x) \leq \underline{u}(x) \qquad\text{in}\; \Omega\backslash \Omega_\delta.
\end{equation*}
For a given $\theta\in (0,1)$, one can always choose $\varsigma$ small enough such that $(1+\varsigma)^{-1} > \theta$. Now we just choose $\theta$ close to $1$ so that $\underline{u}(x) \geq \theta \overline{u}(x) + (1-\theta)\left(\inf_\Omega f\right)$ for $x \in \Omega \setminus \Omega_\delta$ for some $\delta$ depending on $\theta$.
\end{proof}

\noindent This finishes the proof of the well-posedness of \eqref{eq:PDEeps} for $1<p\leq2$. 
\end{proof}

\begin{proof}[Proof of Lemma \ref{lem:max}] The proof is a variation of Perron's method (see \cite{Capuzzo-Dolcetta1990}) and we proceed by contradiction. Let $\varphi\in \rmC(\overline{\Omega})$ and $x_0\in \overline{\Omega}$ such that $u(x_0) = \varphi(x_0)$ and $u-\varphi$ has a global strict minimum over $\overline{\Omega}$ at $x_0$ with 
\begin{equation}\label{eq:max_a1}
      \varphi(x_0) + H(x_0,D\varphi(x_0)) < 0.
\end{equation}
Let $\varphi^\varepsilon(x) = \varphi(x) - |x-x_0|^2 + \varepsilon$ for $x\in \overline{\Omega}$. Let $\delta > 0$. We see that for $x\in \partial B(x_0,\delta)\cap \overline{\Omega}$,
\begin{equation*}
    \varphi^\varepsilon(x) = \varphi(x) - \delta^2 +\varepsilon \leq \varphi(x) - \varepsilon
\end{equation*}
if $2\varepsilon \leq \delta^2$. We observe that
\begin{equation*}
    \begin{split}
    \varphi^\varepsilon(x) - \varphi(x_0)  &= \varphi(x)-\varphi(x_0) + \varepsilon - |x-x_0|^2 \\
    D\varphi^\varepsilon(x) - D\varphi(x_0) &= D\varphi(x) - D\varphi(x_0) - 2(x-x_0)
    \end{split}
\end{equation*}
for $x\in B(x,\delta)\cap \overline{\Omega}$. By the continuity of $H(x,p)$ near $(x_0,D\varphi(x_0))$ and the fact that $\varphi\in \rmC^1(\overline{\Omega})$, we can deduce from \eqref{eq:max_a1} that if $\delta$ is small enough and $0<2\varepsilon < \delta^2$, then
\begin{equation}\label{eq:max_a2}
      \varphi^\varepsilon(x)+H(x,D\varphi^\varepsilon(x)) < 0 \qquad\text{for}\;x\in B(x_0,\delta)\cap \overline{\Omega}.
\end{equation}
We have found $\varphi^\varepsilon\in \mathrm{C}^1(\overline{\Omega})$ such that $\varphi^\varepsilon(x_0)>u(x_0)$, $\varphi^\varepsilon<u$ on $\partial B(x_0,\delta)\cap \overline{\Omega}$ and \eqref{eq:max_a2}. Let
\begin{equation*}
    \tilde{u}(x) = \begin{cases}
    \max \big\lbrace u(x),\varphi^\varepsilon(x) \big\rbrace &x\in B(x_0,\delta)\cap \overline{\Omega},\\
    u(x)&x\notin B(x_0,\delta)\cap \overline{\Omega}.\\
    \end{cases}
\end{equation*}
We see that $\tilde{u}\in \rmC(\overline{\Omega})$ is a subsolution of \eqref{eq:PDE0} in $\Omega$ with $\tilde{u}(x_0) > u(x_0)$, which is a contradiction. Thus, $u$ is a supersolution of \eqref{eq:PDE0} on $\overline{\Omega}$.
\end{proof}

\begin{thm}\label{convex} Let $H(x,p) = G(p)-f(x)$ where $G\geq 0$ with $G(0) = 0$ is a convex function from $\R^n\to\R^n$ and $f\in \mathrm{C}^2_c(\R^n)$. Let $u\in \mathrm{C}_c(\R^n)$ be a viscosity solution to $u+H(x,Du) = 0$ in $\R^n$. Then $u$ is semiconcave, i.e, $u$ is a viscosity solution of  $-D^2u \succ -c\;\mathbb{I}_n$ in $\R^n$ where 
\begin{equation*}
    c = \max \big\lbrace  D_{\xi\xi}f(x): |\xi|=1, x\in \overline{\Omega} \big\rbrace\geq 0.
\end{equation*}
\end{thm}
\begin{proof} Let us consider the auxiliary functional
\begin{equation*}
    \Phi(x,y,z) = u(x)-2u(y)+u(z) - \frac{\alpha}{2}|x-2y+z|^2 - \frac{c}{2}|y-x|^2-\frac{c}{2}|y-z|^2
\end{equation*}
for $(x,y,z)\in \R^n\times\R^n\times\R^n$. By a priori estimate, $u$ is bounded and Lipschitz, thus we can assume $\Phi$ achieves its maximum over $\R^n\times\R^n\times \R^n$ at $(x_\alpha,y_\alpha,z_\alpha)$. The viscosity solutions tests give us
\begin{align*}
    &u(x_\alpha) + G\big(p_\alpha+c(x_\alpha-y_\alpha)\big) \leq f(x_\alpha)\\
    &u(z_\alpha) + G\big(p_\alpha+c(z_\alpha-y_\alpha)\big) \leq f(z_\alpha)\\
    &u(y_\alpha) + G\left(p_\alpha+\frac{c}{2}(x_\alpha-y_\alpha) + \frac{c}{2}(z_\alpha-y_\alpha)\right) \geq f(y_\alpha),
\end{align*}
where $p_\alpha = \alpha(x_\alpha-2y_\alpha+z_\alpha)$. Using the convexity of $G$ we have
\begin{equation*}
    2G\left(p_\alpha+\frac{c}{2}(x_\alpha-y_\alpha) + \frac{c}{2}(z_\alpha-y_\alpha)\right) \leq G\big(p_\alpha+c(x_\alpha-y_\alpha)\big)  + G\big(p_\alpha+c(z_\alpha-y_\alpha)\big)
\end{equation*}
therefore
\begin{equation*}
    u(x_\alpha) - 2u(y_\alpha) + u(z_\alpha) \leq f(x_\alpha) - 2f(y_\alpha) + f(z_\alpha).
\end{equation*}
\begin{itemize}
    \item $\Phi(x_\alpha,y_\alpha,z_\alpha)\geq \Phi(0,0,0)$ gives us
    \begin{equation*}
        \frac{\alpha}{2}|x-2y+z|^2 + \frac{c}{2}|y-x|^2+\frac{c}{2}|y-z|^2 \leq C.
    \end{equation*}
    Thus we can assume $(x_\alpha-y_\alpha)\to h_0$ and $(y_\alpha-z_\alpha)\to h_0$ as $\alpha\to \infty$.
    \item $\Phi(x_\alpha,y_\alpha,z_\alpha )\geq \Phi(y_\alpha+h_0,y_\alpha,y_\alpha-h_0)$ gives us
    \begin{align*}
        u(x_\alpha)-2u(y_\alpha) + u(z_\alpha) - \frac{\alpha}{2}|x_\alpha - 2y_\alpha+z_\alpha|^2 - \frac{c}{2}|x_\alpha-y_\alpha|^2 - \frac{c}{2}|y_\alpha-z_\alpha|^2 \\
        \geq u(y_\alpha+h_0) - 2u(y_\alpha) + u(y_\alpha-h_0) - c|h_0|^2.
    \end{align*}
    Therefore, using the fact that $u$ is Lipschitz, we have
    \begin{equation*}
    \begin{split}
        \frac{\alpha}{2}|x_\alpha - 2y_\alpha+z_\alpha|^2 
        &\leq c\left(\frac{2|h_0|^2 - |x_\alpha-y_\alpha|^2 - |y_\alpha-z_\alpha|^2}{2}\right) \\
        &+ C\Big(|(x_\alpha-y_\alpha) - h_0| + |(z_\alpha - y_\alpha) + h_0|\Big) \to 0
    \end{split}
    \end{equation*}
    as $\alpha\to \infty$.
\end{itemize}
For any $x\in \R^n$, we have $\Phi(x_\alpha,y_\alpha,z_\alpha) \geq \Phi(x+h,x,x-h)$, thus
\begin{align*}
    u(x+h)-2u(x)+u(x-h)-c|h|^2 &\leq f(x_\alpha)-2f(y_\alpha) + f(z_\alpha) \\
    &- \frac{\alpha}{2}|x_\alpha - 2y_\alpha+z_\alpha|^2-\frac{c}{2}|y_\alpha-x_\alpha|^2-\frac{c}{2}|y_\alpha-z_\alpha|^2.
\end{align*}
If $\{y_\alpha\}$ is unbounded as $\alpha\to \infty$ then since $f\in \mathrm{C}_c^2(\R^n)$, we have $f(y_\alpha)\to 0$ as $\alpha\to \infty$. As a consequence, $x_\alpha,z_\alpha \to \infty$ as well and thus $f(x_\alpha)-2f(y_\alpha) + f(z_\alpha)\to 0$ as $\alpha\to \infty$. Therefore
\begin{equation*}
    u(x+h)-2u(x)+u(x-h)-c|h|^2 \leq 0.
\end{equation*}
If $\{y_\alpha\}$ is bounded, then $y_\alpha\to y_0$ as $\alpha\to \infty$, thus
\begin{equation*}
     u(x+h)-2u(x)+u(x-h)-c|h|^2 \leq f(y_0+h_0) - 2f(y_0) + f(y_0-h_0) -c|h_0|^2.
\end{equation*}
Let $\xi = h_0$, we have
\begin{equation*}
    \begin{cases}
      f(y_0+h_0) - f(y_0) = \int_0^1 D_\xi f(y_0 + t\xi)\cdot \xi dt\\
      f(y_0) - f(y_0-h_0)  = \int_0^1 D_\xi f(y_0 - t\xi)\cdot \xi dt.
    \end{cases}
\end{equation*}
Therefore
\begin{equation*}
\begin{split}
     f(y_0+h_0) - 2f(y_0) + f(y_0-h_0) &= \int_0^1 \Big(D_\xi f(y_0 + t\xi) - D_\xi f(y_0 - t\xi) \Big)\cdot \xi dt\\
     &= \int_0^1 \int_0^1 D^2_{\xi\xi} \xi \cdot f(y_0+st \xi)\cdot \xi\;dtds.
\end{split}
\end{equation*}
Therefore
\begin{equation*}
    |f(y_0+h_0) - 2f(y_0) + f(y_0-h_0)| \leq \left(\max_{|\xi|=1}D_{\xi\xi} f\right)|\xi|^2.
\end{equation*}
Hence
\begin{equation*}
    u(x+h)-2u(x)+u(x-h)-c|h|^2 \leq 0
\end{equation*}
and thus $u$ is semiconcave. It is easy to see that if $\varphi$ is smooth and $u-\varphi$ has a local min at $x$ then $D^2\varphi(x)\prec c\;\mathbb{I}$, thus $-D^2\varphi(x)\geq -c\;\mathbb{I}$.
\end{proof}
%\textcolor{orange}{(I double checked the proof of Jeff and I think his proof is fine. Actually we can choose $y_\alpha$ so that $\{y_\alpha\}$ is bounded, only using the fact that $u$ is bounded. But your proof is good too.)}

\end{appendices}

\stoptocwriting
\section*{Acknowledgement}
The authors would like to express their appreciation to Hung V. Tran for his invaluable guidance.
\resumetocwriting

\bibliography{zzzzlibrary}{}
%\bibliographystyle{ieeetr}
\bibliographystyle{acm}

\end{document}