### 2021-07-15 Drafts and notes/Notes/Notes-May18.tex

\documentclass[english,reqno]{amsart}
%\renewcommand{\baselinestretch}{1.2}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\raggedbottom
\usepackage{subfig}
\usepackage{amssymb,amsbsy,amsmath,amsfonts,amssymb,amscd}
\usepackage{comment}
\usepackage{verbatim}
\usepackage[dvipsnames]{xcolor}
\usepackage{enumitem}
\usepackage[numbers]{natbib} % for alphabetically numbered lists
\usepackage{algorithm,algpseudocode}
\usepackage{setspace}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{assum}{Assumption}[section]
\newtheorem{cor}{Corollary}[section]
\newenvironment{solution}
  {\renewcommand\qedsymbol{$\blacksquare$}\begin{proof}[Solution]}
  {\end{proof}}
\title{Classical mechanics Hamiltonian}

\begin{document}

\maketitle
\date{May 18, 2021}
\section{The first-order equation}\noindent
Work out the details of the following example in $n=1$. Let u be the solution to
    \begin{equation}
    \label{example}
        \left\{
  \begin{aligned}
   u(x) + \frac{1}{2}|u^\prime(x)|^2 -f(x) &\leq 0 \, \quad \qquad \text{in } (-1, 1), \\
              u(x) + \frac{1}{2}|u^\prime (x)|^2 -f(x) &\geq 0 \qquad  \quad \text{on } [-1,1],
  \end{aligned}
\right.
    \end{equation}
with an easier $f$, using deterministic optimal control formula and Euler-Lagrange equation. For instance, $f=0$ in $[-\frac{1}{2}, \frac{1}{2}]$ and $f(x) = x-\frac{1}{2}$ on $[\frac{1}{2},1]$ and $f(x) = -\frac{1}{2} - x$ on $[-1,-\frac{1}{2}]$.
And then try to look into the solution to the equation with $\epsilon \Delta u^\epsilon$.

\paragraph{\textbf{Optimal control formula}}
\noindent Solution is given by
\begin{equation*}
    u(x_0) = \inf_{\gamma(0)=x_0} \int_0^\infty e^{-s}\left(\frac{|\dot{\gamma}(s)|^2}{2} + f(\gamma(s))\right)ds
\end{equation*}
where the admissible set are absolutely continuous curves $\gamma$ such that $\gamma(s)\in \overline{\Omega}$ for all $s\geq 0$ and $\gamma(0)=x_0$.

\paragraph{\textbf{Euler-Lagrange equation}}    
\begin{equation}
\label{EL}
\left\{
\begin{aligned}
     \Ddot{\gamma}(s)&=\dot{\gamma}(s) + f^\prime (\gamma(s)), \qquad s\in (0,\infty)\\
     \gamma(0)&=x_0.
\end{aligned}
\right.
\end{equation}
\paragraph{\textbf{Gradient bound}} Since $u\equiv 1$ is a supersolution and $u\equiv 0$ is a subsolution, we have $0\leq u\leq 1$, thus
\begin{equation*}
    \frac{1}{2}|u'(x)|^2 \leq f(x) - u(x) \leq 1
\end{equation*}
and thus (in viscosity sense)
\begin{equation*}
    |u'(x)| \leq 2.
\end{equation*}
\paragraph{\textbf{An identity (analog of conservation of energy)}} 
\begin{equation*}
    \frac{d}{ds}\left(\frac{|\dot{\gamma}(s)|^2}{2} - f(\gamma(s))\right) = |\dot{\gamma}(s)|^2.
\end{equation*}

\paragraph{\textbf{Observations}}
\begin{itemize}
    \item If $x_0\in [-\frac{1}{2},\frac{1}{2}$ then obviously the minimizer is $\gamma(s) \equiv x_0$ for all $s\geq 0$, which resulted in $u(x_0) = 0$.
    \item We are interested in solving for $u(x_0)$ if $x_0\notin [-\frac{1}{2},\frac{1}{2}]$. For example let us consider $x_0 \in [\frac{1}{2},1]$, then heuristically a minimizing path has to traverses to the left to reach $\frac{1}{2}$ so that the potential term
    \begin{equation*}
        f(\gamma(s))    
    \end{equation*}
    in the cost function decreases, while also try to move not too fast so that the kinetic term
    \begin{equation*}
        \frac{1}{2}|\dot{\gamma}(s)|^2
    \end{equation*} 
    does not contribute too much. So we have a game of balancing between the kinetic energy and potential energy.
    \item Now we see that, it could be the case that it takes an infinite amount of time to reach $\frac{1}{2}$. Let us denote $t_0$ the fist time $\gamma$ reaches $\frac{1}{2}$.
\end{itemize}
\subsection{The case $t_0$ is finite} From the dynamic programming principle we have
\begin{equation*}
    u(x_0) = \int_0^{t_0}e^{-s}\left(\frac{|\dot{\gamma}(s)|^2}{2} + f(\gamma(s))\right)ds + \underbrace{u(\gamma\left(t_0\right))}_{u(1/2) = 0}
\end{equation*}
On $[0,t_0]$, the Euler-Lagrange equation reads 
\begin{equation*}
    \ddot{\gamma}(s) = \dot{\gamma}(s)+1, \qquad s\in (0,t_0)
\end{equation*}
thus we can solve for $\dot{\gamma}$, assuming the initial velocity $\dot{\gamma}(0) = \xi_0$. We have
\begin{equation*}
    \dot{\gamma}(s) = (\xi_0+1)e^{s} - 1, \qquad s\in (0,t_0).
\end{equation*}
Thus
\begin{equation*}
    \gamma(s) = (\xi_0+1)(e^s-1) - s + x_0, \qquad s\in (0,t_0).
\end{equation*}
Set $\gamma(t_0) = \frac{1}{2}$ we have
\begin{equation*}
    (\xi_0+1)(e^{t_0}-1) - t_0 + x_0 = \frac{1}{2}.
\end{equation*}
We compute (should be with a minimizer)
\begin{equation*}
\begin{split}
    u(x_0) &= \int_0^{t_0}e^{-s}\left(\frac{1}{2}\Big((\xi_0+1)e^{s}-1\Big)^2 +\left[ (\xi_0+1)(e^{s}-1)-s+x_0 - \frac{1}{2}\right]\right)ds\\
    &=\int_0^{t_0}\left[\frac{1}{2}(\xi_0+1)^2e^{s}+\frac{e^{-s}}{2} - (\xi_0+1) \right]ds + \int_0^{t_0}\left((\xi_0+1)(1-e^{-s})-se^{-s} + e^{-s}\left(x_0-\frac{1}{2}\right)\right)ds.
\end{split}
\end{equation*}
We have
\begin{align*}
    &\int_0^{t_0}\left[\frac{1}{2}(\xi_0+1)^2e^{s}+\frac{e^{-s}}{2} - (\xi_0+1) \right]ds = \frac{(\xi_0+1)^2(e^{t_0}-1)}{2} + \frac{1-e^{-t_0}}{2} - (\xi_0+1)t_0 \\
    &\int_0^{t_0} (\xi_0+1)(1-e^{-s}) ds = (\xi_0+1)t_0 - (\xi_0+1)(1-e^{-t_0})\\
    &\int_0^{t_0} -se^{-s}ds = t_0e^{-t_0} + (1-e^{-t_0})\\
    &\int_0^{t_0} e^{-s}\left(x_0-\frac{1}{2}\right)ds = \left(x_0-\frac{1}{2}\right)(1-e^{-t_0}).
\end{align*}
Therefore
\begin{equation*}
\begin{split}
    u(x_0) &= \frac{(\xi_0+1)^2(e^{t_0}-1)}{2} +t_0e^{-t_0} + (1-e^{-t_0})\Bigg(\frac{1}{2}-\xi_0-1 +1+x_0-\frac{1}{2} \Bigg)\\
    &= \frac{(\xi_0+1)^2(e^{t_0}-1)}{2} + t_0e^{-t_0} + (1-e^{-t_0})(x_0-\xi_0)
\end{split}
\end{equation*}
Now given a time $t_0\in (0,\infty)$ from the equation
\begin{equation*}
    (\xi_0+1)(e^{t_0}-1) - t_0 + x_0 = \frac{1}{2}
\end{equation*}
we can uniquely solve for $\xi_0 = \xi_0(t_0)$ as a function (smooth, obviously). Thus we can try to optimize the function
\begin{equation*}
    \mathcal{G}(t_0) =\frac{(\xi_0+1)^2(e^{t_0}-1)}{2} + t_0e^{-t_0} + (1-e^{-t_0})(x_0-\xi_0).
\end{equation*}
Taking derivative

\subsection{A different potential}
Now using 
\begin{equation*}
    f(x) = \begin{cases}
    0 & |x|\leq \frac{1}{2}\\
    \left(|x|-\frac{1}{2}\right)^2 & |x|\in \left[\frac{1}{2},1\right].
    \end{cases}
\end{equation*}
We can use the fact that for a minimizer $\gamma$ starting at $x_0\in \left[\frac{1}{2},1\right]$ then $\gamma(t_0) = \frac{1}{2}$ implies $\dot{\gamma}(t_0) = 0$. We have
\begin{equation*}
    \displaystyle \gamma (s) = \frac{1}{2}+\left( x_0 -\frac{1}{2}\right) e^{-s},\qquad s\geq 0.
\end{equation*}
We deduce that for $|x|\geq \frac{1}{2}$ then
\begin{equation*}
    u(x) = \frac{3}{2}\int_0^\infty  \left(x-\frac{1}{2}\right)^2e^{
    -3s}ds = \frac{1}{2}\left(x-\frac{1}{2}\right)^2.
\end{equation*}

\end{document}